---
title: 'OHI 2020 - Land based nutrient and chemical pourpoints prep '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../workflow/templates/ohi_hdr.html' 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---


[REFERENCE RMD FILE](http://ohi-science.org/ohiprep_v2020/globalprep/cw_nutrient_o_chem/v2020/pourpoints_PO4eq_prep.html)


# Summary
This document describes the steps for obtaining and prepping data used for the land based nutrient and chemical data layer for clean waters. In specific, this data prep focuses on prepping the PO4eq pourpoint shapefile, which will be fed into the plume model. 

## The following data are used:

* 


# Updates from previous assessment
This is a new layer for the 2020 assessment year.

## Initial set-up code

```{r setup, eval = FALSE}
library(tidyverse)
library(raster)
library(sf)
library(mapview)
library(janitor)


source('../../workflow/R/common.R')

ww_raw_dir <- "/home/shares/ohi/git-annex/land-based/wastewater/data/raw"
ww_intermediate_dir <- "/home/shares/ohi/git-annex/land-based/wastewater/data/interim"
cw_nutrient_int_dir <- "/home/shares/ohi/git-annex/globalprep/cw_nutrient_o_chem/int"

```

# Methods 

**Read in watersheds files, transform the CRS to 4326, and bind together to make one file.**
These watershed files are taken from the wastewater project at NCEAS. 
```{r, eval = FALSE}
## Watersheds; read in, reproject to EPSG 4326, and combine into one file. 

watersheds_af <- st_read(file.path(ww_raw_dir, "basins_laea/af_bas.shp")) %>%
  clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id) ## select columns of interest
crs(watersheds_af) ## check CRS
plot(watersheds_af$geometry)
watersheds_af <- st_transform(watersheds_af, 4326) ## change CRS to 4326
plot(watersheds_af$geometry)
crs(watersheds_af)
st_write(watersheds_af, file.path(cw_nutrient_int_dir, "watersheds/watersheds_af_4326.shp")) ## write transformed file to mazu. 

watersheds_au <- st_read(file.path(ww_raw_dir, "basins_laea/au_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_au$geometry)
watersheds_au <- st_transform(watersheds_au, 4326)
plot(watersheds_au$geometry)
crs(watersheds_au)
st_write(watersheds_au, file.path(cw_nutrient_int_dir, "watersheds/watersheds_au_4326.shp"))

watersheds_eu <- st_read(file.path(ww_raw_dir, "basins_laea/eu_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_eu$geometry)
watersheds_eu <- st_transform(watersheds_eu, 4326) 
crs(watersheds_eu)
st_write(watersheds_eu, file.path(cw_nutrient_int_dir, "watersheds/watersheds_eu_4326.shp"))

watersheds_na <- st_read(file.path(ww_raw_dir, "basins_laea/na_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_na$geometry)
watersheds_na <- st_transform(watersheds_na, 4326) 
plot(watersheds_na$geometry)
crs(watersheds_na)
st_write(watersheds_na, file.path(cw_nutrient_int_dir, "watersheds/watersheds_na_4326.shp"))

watersheds_pa <- st_read(file.path(ww_raw_dir, "basins_laea/pa_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_pa$geometry)
watersheds_pa <- st_transform(watersheds_pa, 4326) 
plot(watersheds_pa$geometry)
crs(watersheds_pa)
st_write(watersheds_pa, file.path(cw_nutrient_int_dir, "watersheds/watersheds_pa_4326.shp"))

watersheds_sa <- st_read(file.path(ww_raw_dir, "basins_laea/sa_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_sa$geometry)
watersheds_sa <- st_transform(watersheds_sa, 4326) 
crs(watersheds_sa)
st_write(watersheds_sa, file.path(cw_nutrient_int_dir, "watersheds/watersheds_sa_4326.shp"))

watersheds_as <- st_read(file.path(ww_raw_dir, "basins_laea/as_bas.shp")) %>%
    clean_names() %>%
  dplyr::select(id, gridcode, area, basin_id)
plot(watersheds_as$geometry)
watersheds_as <- st_transform(watersheds_as, 4326) 
crs(watersheds_as)
st_write(watersheds_as, file.path(cw_nutrient_int_dir, "watersheds/watersheds_as_4326.shp"))

watersheds_all <- rbind(watersheds_eu, watersheds_au, watersheds_na, watersheds_pa, watersheds_af, watersheds_sa, watersheds_as) ## bind all of these files together to make the global watershed files with CRS 4326. 
plot(watersheds_all$geometry)

st_write(watersheds_all, file.path(cw_nutrient_int_dir, "watersheds/watersheds_all_4326.shp"))

```

**Read in pourpoints file and transform crs to be 4326**
```{r, eval = FALSE}
## Pourpoints

global_plume <- st_read("/home/shares/ohi/git-annex/land-based/wastewater/data/raw/pour_points/global_plume_2007_2010.shp")
plot(global_plume$geometry)
mapview(global_plume)

global_plume <- st_transform(global_plume, 4326) 

```

**Read in PO4eq fertilizer data and disaggregate by 15 to make higher resolution.**
```{r, eval = FALSE}
## N/P fertilizer data

n_p_rast <- raster(file.path("/home/shares/ohi/git-annex/globalprep/_raw_data/NCEAS_food_systems/d2020/crop_fert_PO4eq.tif"))
plot(log(n_p_rast + 1))

crs(n_p_rast) 

n_p_rast
plot(area(n_p_rast))

n_p_dis_rast <- disaggregate(n_p_rast, fact = 15, method = '', progress = TRUE) # this takes awhile
n_p_dis_rast
plot(n_p_dis_rast)
plot(area(n_p_dis_rast))

#writeRaster(x = n_p_dis_rast, filename = file.path(cw_nutrient_int_dir, "rasters/crop_fert_PO4eq_disagg.tif"), overwrite = TRUE)

n_p_dis_log_rast <- raster::calc(n_p_dis_rast, fun = function(x){log(x + 1)}) ## take the log of this and save so that we can better visualize the disaggregated raster. 

#writeRaster(x = n_p_dis_log_rast, filename = file.path(cw_nutrient_int_dir, "rasters/crop_fert_PO4eq_disagg_log.tif"))

plot(n_p_dis_log_rast)
plot(watersheds_all$geometry, add = TRUE)
```

**Calculate zonal stats per watershed and save.**
Take the watersheds shapefile and rasterize it so that we can run zonal stats on it. We will calculate the zonal stats per each watershed to obtain an estimate of the PO4eq concentration per each watershed. 
```{r, eval = FALSE}
## Zonal stats per watershed 
n_p_dis_rast <- raster::raster(file.path(cw_nutrient_int_dir, "rasters/crop_fert_PO4eq_disagg.tif")) ## read in disaggregated raster

watersheds_all <- st_read(file.path(cw_nutrient_int_dir, "watersheds/watersheds_all_4326.shp")) ## read in watersheds file

wat_length <- length(watersheds_all$basin_id)

watersheds_all <- watersheds_all %>%
  arrange(id) %>%
  mutate(id_new = 1:wat_length) # produce a new id for watersheds so that we can fasterize correctly... I don't believe fasterize would work with basin_id since it is a character variable. 


plot(n_p_dis_log_rast)
plot(watersheds_all$geometry, add = TRUE)


raster_watersheds <- fasterize::fasterize(watersheds_all, n_p_dis_rast, field = "id_new") ## make watershed raster
plot(raster_watersheds)
#writeRaster(x = raster_watersheds, filename = file.path(cw_nutrient_int_dir, "rasters/raster_watersheds.tif"), overwrite = TRUE)

raster_watersheds <- raster::raster(file.path(cw_nutrient_int_dir, "rasters/raster_watersheds.tif"))

zs_n_p <- raster::zonal(n_p_dis_rast, raster_watersheds, fun = 'sum', progress = 'text', na.rm = TRUE) ## calculate sum of PO4eq per watershed
zs_n_p_df <- data.frame(zs_n_p)
#write_csv(zs_n_p_df, "int/zonal_n_p_watersheds.csv")
zs_n_p_df <- read_csv("int/zonal_n_p_watersheds.csv")


watersheds_all_zs <- watersheds_all %>%
  left_join(zs_n_p_df, by = c("id_new" = "zone"))

st_write(watersheds_all_zs, file.path(cw_nutrient_int_dir, "watersheds/watersheds_zonal_stats.shp"))

## 43 NAs of sum after joining... This means that 43 watersheds are being left out for some reason... likely because they are too small for fasterizing/zonal statistics. 
```

**Fix the missing 43 watersheds that zonal stats missed and add back into the data**
```{r, eval = FALSE}
watersheds_all <- st_read(file.path(cw_nutrient_int_dir, "watersheds/watersheds_all_4326.shp")) %>%
  arrange(id)  ## read in watersheds file

watersheds_zonal <- st_read(file.path(cw_nutrient_int_dir, "watersheds/watersheds_zonal_stats.shp"))

watersheds_zonal_NA <- watersheds_zonal %>%
  dplyr::filter(is.na(sum)) ## filter for NAs

watersheds_zonal_NA$centroid <- st_centroid(watersheds_zonal_NA) %>%
  st_geometry() ## check the centroids

## take a look to compare and see if it worked correctly
plot(st_geometry(watersheds_zonal_NA))
plot(st_set_geometry(watersheds_zonal_NA, 'centroid')[, 0], add = T, col = 'red', pch = 19) ## it worked..

## now lets extract from our PO4eq raster disaggregated
watersheds_zonal_NA_centroid <- watersheds_zonal_NA %>%
  dplyr::select(-centroid) %>%
  st_centroid() %>%
  dplyr::select(-sum)

## extract fron PO4eq
zonal_NA_sum <- extract(n_p_dis_rast, watersheds_zonal_NA_centroid, method = 'simple') ## extract zonal stats based on centroid point data 

# join extracted values with NA df
watersheds_zonal_NA_fixed <- data.frame(watersheds_zonal_NA, sum_new = zonal_NA_sum) %>%
  mutate(sum = ifelse(is.na(sum), sum_new, sum)) %>%
  dplyr::select(-centroid, -sum_new) %>%
  st_as_sf()
 

##  bind back together with overall zonal sf
watersheds_zonal_all_fixed <- watersheds_zonal %>%
  dplyr::filter(!is.na(sum)) %>%
  rbind(watersheds_zonal_NA_fixed)  %>%
  dplyr::select(basin_id, PO4eq_total = sum) ## now we have a dataset with the columns we want

st_write(watersheds_zonal_all_fixed, file.path(cw_nutrient_int_dir, "watersheds/watersheds_zonal_FIXED.shp"))

```

**Combine watersheds zonal stats with pourpoints data**
```{r, eval = FALSE}
watersheds_zonal_all_fixed <- as.data.frame(st_read(file.path(cw_nutrient_int_dir, "watersheds/watersheds_zonal_FIXED.shp"))) %>%
  dplyr::select(-geometry, basin_id = basin_d, PO4eq_total = PO4q_tt) ## fix the column names, for some reason they were mispelled in the writing process
  

length(unique(global_plume$basin_id)) # 142642
length(unique(watersheds_zonal_all_fixed$basin_id)) # 142641

## there is one more watershed present in the pourpoint data than in the watershed data

## join these datasets together 
ws_pp_zonal_all <- left_join(watersheds_zonal_all_fixed, global_plume, by = "basin_id") %>%
  dplyr::select(-3, -4, -5)

unique(ws_pp_zonal_all$basin_id)

n_occur <- data.frame(table(ws_pp_zonal_all$basin_id)) %>% ## check the duplicates (or triplicates)
      filter(Freq >1) ## filter for the duplicates in the data

dups <- ws_pp_zonal_all %>%
  filter(basin_id %in% n_occur$Var1) %>% ## duplicates... there appear to be multiple geometries for some basin_ids
st_as_sf()
  
mapview(dups)

dups_orig <- watersheds_zonal_all_fixed %>%
  filter(basin_id %in% n_occur$Var1) ## compare to original data

dups_pourpoints <- global_plume %>%
  filter(basin_id %in% n_occur$Var1) ## compare to pourpoints data 

## I've convinced myself this is nothing to worry about. Just select one of the multiple geometries and use that. They are all very close together. 

dups_fix <- dups %>%
  group_by(basin_id) %>%
  filter(row_number() == 1) %>% ## filter for the first row in each basin id group.
  ungroup() %>%
  as.data.frame()

`%notin%` <- Negate(`%in%`)

## remove the dups from joined data 
ws_pp_zonal_all_no_dups <- ws_pp_zonal_all %>%
  filter(basin_id %notin% dups$basin_id)

## add the dups back into the no dups file
ws_pp_zonal_all_fix <- ws_pp_zonal_all_no_dups %>%
  rbind(dups_fix) %>%
  st_as_sf() %>%
  rename("effluent" = "PO4eq_total")
  
str(ws_pp_zonal_all_fix)

summary(ws_pp_zonal_all_fix)

st_write(ws_pp_zonal_all_fix, file.path(cw_nutrient_int_dir, "pourpoints/PO4eq_pourpoints.shp"), overwrite=TRUE, delete_dsn = TRUE) ## this is our final pourpoint data set we will plug into the plume model 

mapview(ws_pp_zonal_all_fix)
```

**Create a shapefile to test the plume model on first**
```{r, eval = FALSE}
## Write out an Australia shapefile to use for testing... 

ws_pp_zonal_all_fix <- st_read(file.path(cw_nutrient_int_dir, "pourpoints/PO4eq_pourpoints.shp")) %>%
  mutate(basin_id = as.character(basin_id))
crs(ws_pp_zonal_all_fix)

au_pp_all <- ws_pp_zonal_all_fix %>%
  dplyr::filter(str_detect(basin_id, "au")) %>%
  arrange(effluent) %>%
  top_n(10, effluent)

mapview(au_pp_all)

st_write(au_pp_all, file.path(cw_nutrient_int_dir, "pourpoints/AU_PO4eq_test.shp"), overwrite = TRUE, delete_dsn = TRUE)
```

