---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages.. you might need to install them first
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(tidyverse)
library(arsenal)
```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Melanie or Gage"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You can ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

*Note* There are scores for 2022, you can ignore these. These are just here as a placeholder until we calculate new ones for 2022 (they are the 2021 scores). 

```{r, include=FALSE}
# load in latest OHI global data
# using read.csv(here()) that requires the data to be in the proj:
#scores <- read.csv(here("Reference/Lessons_in_OHI/explore_scores/scores.csv"))
# reading in directly from github:
scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220? Can you make this data more human readable by including the full goal name (vs. the 2 letter abbreviation) and the country name.

*Hint 1* Here is a link to the official region list (the rgn_id column in this file matches the region_id column in the scores data): https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Here is a table that includes goal names: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv

*Hint 3* Which region has all NA values for scores?

*Answer* Antarctica has all NA values for score. This region is not included in the assessment, as noted in the 'Notes' column, which explains why there is 1 extra values The `regions` data starts at index 1 for the `regions_id` values, and the `scores` data starts at index 0 for the `regions_id` values. This explains why there is 1 extra value for `region_id` than expected.

```{r, include=FALSE}
# read in regions_list.csv data
regions <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>%
  # rgn_id column = region_id column in scores data, so rename column here to match
  rename(region_id = rgn_id)

# read in the ten_goals.csv
goals <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv") %>% 
  rename(goal = Abbreviation)

# join scores data to regions data
scores_regions <- full_join(scores, regions)

# join the goals data to joined data
scores_regions_goals <- full_join(scores_regions, goals)

# determine which country has all NA for "score" column
na_by_rgn <- scores_regions_goals %>% 
  group_by(rgn_name, score) %>%
  summarize() %>% 
  filter(all(is.na(score)))
# Antarctica

# explore why there are 2 extra regions by comparing the vectors of the region_id columns in the regions & scores data
# confirm that 222 unique region_id values are present in the joined data
n_distinct(scores_regions_goals$region_id) # 222

# determine all unique values for region_id
unique(scores_regions_goals$region_id)

# do any of the region_id's have notes?
unique(scores_regions_goals$Notes) # one value for Notes indicates that at least 1 row is a region that is not included in the global assessment

not_inc <- scores_regions_goals %>% 
  filter(Notes == "Not included in global assessment")
head(not_inc) # Antarctica is the country that is not included

# explore unique region_id values in the independent data frames with this variable (regions and scores)
# regions:
unique(regions$region_id) # starts at index 1
n_distinct(regions$region_id) # 221 unique id values

# scores:
unique(scores$region_id) # starts at index 0
n_distinct(scores$region_id) # 222 unique values

diff_id <- setdiff(scores$region_id, regions$region_id)
diff_id # value of 0 is the difference between these vectors
```

*Question* Which region id corresponds to the United States?
*Answer* region id 163 corresponds to the United States

```{r}
# determine the region id for the United States
usa <- scores_regions_goals %>% 
  filter(rgn_name == "United States")

head(usa) # region_id = 163
```


### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/documents/methods/Supplement.html

*Answer* 10 goals of OHI assessments are abbreviated with 2 letters, which indicates that they are the 10 main goals, and the 8 sub-goals are indicated by 3 letters. An example of a main goal is "Biodiversity", which is abbreviated as "BD", and this goal has 2 sub-goals: "Habitat" and "Species", which are abbreviated with 3 letters each as "HAB" and "SPP", respectively. There is also a value of "Index" within the `goals` variable, which makes up the 19th unique goal value. 

```{r, include=FALSE}
# determine the number of distinct goals in the joined data to confirm it is indeed 19
n_distinct(scores_regions_goals$goal) # 19

# determine unique goal values in the joined data
goal_ids <- unique(scores_regions_goals$goal)
goal_ids

# determine which rows have "index" for the goal
index_goal <- scores_regions_goals %>% 
  filter(goal == "Index")
index_goal

```


### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  

*Answer* The difference of 2442 rows for these dimensions is explained by the 2442 rows within the `Index` goal that are associated with the `score` dimension. Since the only dimensions associated with `Index` are `score` and `future`, and the rows associated with each are equal, I infer that `Index` is the result of a calculation that includes these both these dimensions or at least the `score` dimension as an input.

```{r, include=FALSE}
# explore the unique values of the dimension variable
unique(scores_regions_goals$dimension)

# check how many rows of each dimension are present
unique(scores$dimension)
table(scores$dimension)

#table(scores$goal)
#table(scores$goal, scores$dimension)

# check how many rows of each dimension are present (in a different way, in just 1 df)
dimensions_explore <- scores_regions_goals %>% 
  group_by(dimension) %>% 
  summarize(n())

# check number of rows that are different between score and status
46398 - 43956 # 2442

unique(scores_regions_goals$goal) # explore the index goal as I am not familiar with it

index_goal <- scores_regions_goals %>% 
  filter(goal == "Index")
nrow(index_goal) # 4884

unique(index_goal$dimension) # there are only "future" and "score" in the dimension column for the "index" goal, meaning there are no index for the dimension "status", so index might be the result of a calculation that includes status

index_goal_score <- index_goal %>% 
  filter(dimension == "score")
nrow(index_goal_score) # 2442
```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2021.  Which ones have the least? Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2021, and cut region_id = 0.

*Answer* The goal with the most NA values for scores in 2021 is carbon storage (`CS`). The subgoal with the most NA values is mariculture (`MAR`). I guess that carbon storage has the most NA values for score in 2021 because many countries may *not*  have ever had or currently have the habitats that are assessed for this goal: mangroves, tidal marshes and seagrasses. If these countries did historically or currently do have these habitats, they are at least not present in the data we are using to assess this goal.

The goals with the fewest NA values (with Antarctica included) are `BD`, `CW`, `FP`, and `SP`. The subgoals with the fewest NA values are `FIS`, `HAB`, `ICO`, `LSP`, and `SPP`. When you remove Antarctica, these goals are not listed amongst those that have NA values. The biodiversity goal probably has no NA values because the data for the species subgoal mostly came from IUCN Red List of Threatened Species, which I would hope and assume is updated very regularly and is comprehensive for many if not all countries. Furthermore, the methodology for the other subgoal, habitats, explains that "a significant amount of pre-processing of the habitat data was needed to fill data gaps and resolve data quality issues" so the gaps that initially existed were filled to the best of OHI's ability before the score was calculated. The other goals and subgoals without missing values also probably had comprehensive data or had NA values filled in before the score was calculated.

```{r, include=FALSE}
# filter data to only include 2021 and count the number of na values for score, grouping by goal
na_2021 <- scores_regions_goals %>% 
  filter(year == 2021,
         dimension == "score",
         region_id != 0,
         # remove Antarctica from the regions
         region_id != 213) %>%
  # select only certain rows to make the dataframe easier to interpret
  select(goal, dimension, region_id, score, year) %>%
  # add col for TRUE/FALSE depending on if the score is NA
  mutate(na_value = is.na(score)) %>%
  # we only are interested in na values
  filter(na_value == TRUE) %>% 
  group_by(goal) %>% 
  # count the total occurrences of na scores by country
  summarise(total_na_scores = n())

```

### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* The overall score is the equally weighted average of future and status scores.

```{r, include=FALSE}
# unique(scores_regions_goals$dimension)

future_status_score <- scores_regions_goals %>%
  filter(dimension %in% c("future", "status", "score")) %>% 
  group_by(goal, dimension) %>% 
  summarise(avg = mean(score, na.rm = TRUE))
# seems like the average of the future and status scores result in the overall score

# use the first goal, AO, as a use case to cheeck this hypothesis
(74.914138 + 75.388657) / 2 # yep it is the 
```


### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

Define variables to read into table when knitting:
```{r}
goals <- unique(scores_regions_goals$goal)
dimensions <- unique(scores_regions_goals$dimension)
min_region_id <- min(scores_regions_goals$region_id)
max_region_id <- max(scores_regions_goals$region_id)
min_score <- min(scores_regions_goals$score)
max_score <- max(scores_regions_goals$score)
```


Variable   | Description                  | Values
---------- | ---------------------------- | ------------------------------
goal       |  ocean benefits delivered to humans, OHI has 10 goals, 4 of which have 2 subgoals.                            | `r goals`
dimension  |  describes the category of the subscores that contributes to the overall score                             | `r dimensions`
region_id  |  unique numeric code that identifies each geographic region (country or territory)                             | `r min_region_id` - `r max_region_id`
score      |  numeric score for goal, region, year, and dimension                             | `r min_score` - `r max_score`         
year       |  Scenario year               | 2012 to 2022


```{r, include=FALSE}
# 

```


### Plot

*Project* Create a scatterplot that compares 2012 and 2021 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space



```


Create a plot that shows the distribution (e.g., histogram or barplot or ??) of *Index* scores of all countries.

```{r across year plot}

## Code to create a plot showing distribution of country Index scores



```
