---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages
library(dplyr)
library(tidyr)
library(tidyverse)

```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Julie or Melanie or Gage"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You can ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

```{r, include=FALSE}
## Working space

# Read in OHI global data
scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220?

*Hint 1* Here is a link to the official region list: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Which region has all NA values for scores?

*Answer* Region 0 is listed but isn't actually a region. Region 0 is the overall OHI global score; it's the average of all the other scores. Region 213 (Antarctica) has all NA values for scores; this is because OHI doesn't assess Antarctica.

```{r, include=FALSE}
## Working space

length(unique(scores$region_id))

# Find NA values
NA_scores <- scores %>% 
  group_by(region_id) %>%
  summarize(na = length(score[is.na(score)])) %>% 
  arrange(-na)

# See regions with most NAs
head(NA_scores)
  # Region 213 has all NAs (Antarctica)

```

*Question* Which region id corresponds to the United States?

*Answer* Region 163 corresponds to the United States.

### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/documents/methods/Supplement.html

*Answer* There are 10 goals (2-letter abbreviations), but 4 of those goals (BD, LE, FP, SP) have 2 subgoals each; the 8 subgoals have 3-letter abbreviations. The last goal is "Index," which refers to the overall average score of all the goals for each region.

```{r, include=FALSE}
## Working space

unique(scores$goal)

```


### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  

*Answer* There are more score values because the index values don't have status values (or pressures, resilience, and trend values). The index is the overall average global score for that region. Region 0 doesn't have values for pressures, resilience, or trend because those 3 variables go into calculating the future scores, which are calculated on a per-country basis. We don't have overall gobal average scores calculated for this, so there are no values for Region 0.

```{r, include=FALSE}
## Working space

unique(scores$dimension)
table(scores$dimension)

table(scores$goal)
table(scores$goal, scores$dimension)

# ok...but why are the different values for pressures/resilience/trend and future?
# addition of region 0!
ao <- filter(scores, goal=="AO")
table(scores$region_id, scores$dimension)

```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2019.  Which ones have the least?  Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2019, and cut region_id = 0.

*Answer* The Natural Products (NP) goal has the most NA values (96), followed by the Carbon Storage (CS) goal (73 NA values). This could be due to the NP goal having a lack of data, probably because a lot of the regions don't produce the natural products assessed in this goal. In general, sometimes we can't get great global data, which will be reflected in the amount of NA values in our scores. BD, CW, FIS, FP, ICO, LSP, SP, and SPP all have the least (1) NA values.

```{r, include=FALSE}
## Working space

# Find which goals/subgoals have most/least NA score values for 2019 (get rid of region 0)
na_data <- scores %>% 
  filter(dimension == "score") %>%
  filter(region_id != 0) %>%
  filter(year == 2019) %>% 
  group_by(goal) %>% 
  summarise(na = length(score[is.na(score)])) %>% 
  arrange(-na)

na_data

```


### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* The score is just the average of the future and present status. Therefore, the score would be 85.

```{r, include=FALSE}
## Working space

source("http://ohi-science.org/ohiprep_v2020/workflow/R/common.R")
region_data()
# Dominica is 123

# Filter for future, status, and score values for Dominica and AO goal
dominica_ao <- scores %>%
  filter(region_id == 123) %>% 
  filter(goal == "AO") %>% 
  filter(dimension == "future" | dimension == "status" | dimension == "score")
  # Score = average of future and status

```


### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

Variable   | Description                  | Values
---------- | ---------------------------- | ------------------------------
goal       | Category of ocean benefit to people. 2-letter abbreviations represent the 10 goals; 3-letter abbreviations represent the 8 subgoals.                            | AO, BD (HAB, SPP), CP, CS, CW, FP (FIS, MAR), LE (LIV, ECO), NP, SP (ICO, LSP), TR
dimension  | Variables used to calculate goal/subgoal scores                             | future, pressures, resilience, score, status, trend
region_id  | Number associated with the region or country assessed                             | 0 - 250
score      | Score associated with each goal/subgoal and dimension                             | 0 - 100         
year       | Scenario year               | 2012 - 2020 

```{r, include=FALSE}
## Working space

summary(scores)

```


### Plot

*Project* Create a scatterplot that compares 2012 and 2019 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* Scores for the artisanal opportunities goal appear to have increased over time, as the graph below shows the majority of points are positive.

```{r, include=FALSE}
## Working space

# Filter out 2012 and 2019 scores for AO goal
ao_2012_2019 <- scores %>% 
  filter(year == "2012" | year == "2019") %>% 
  filter(dimension == "score") %>% 
  filter(goal == "AO") %>% 
  spread(year, score) %>% 
  mutate(ao_difference = `2019` - `2012`)

# Create scatterplot that compares differences in 2012 and 2019 scores for each region for AO goal
ao_2012_2019_plot <- ggplot(ao_2012_2019, aes(x = region_id, y = ao_difference)) +
  geom_point(color = ifelse(ao_2012_2019$ao_difference > 0, "seagreen3", "palevioletred3")) +
  geom_hline(yintercept = 0) +
  labs(x = "Region", y = "Difference in AO Goal Scores between 2012 and 2019") +
  theme_bw() +
  theme(legend.position = "none")
                            
ao_2012_2019_plot                            
        
```

