---
title: "Learning about OHI scores"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
## loading packages.. you might need to install them first
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
```

## OH NO!

While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Melanie or Gage"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions. You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory. However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.

You can ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: <https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv>

*Note* There are scores for 2023, you can ignore these. These are just here as a placeholder until we calculate new ones for 2023 (they are the 2022 scores).

```{r, include=FALSE}
# read in the score data
scores <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

#read in the region data
region_list <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv")

#read in the goal names
goal_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv")
```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220? Can you make this data more human readable by including the full goal name (vs. the 2 letter abbreviation) and the country name.

*Hint 1* Here is a link to the official region list (the rgn_id column in this file matches the region_id column in the scores data): <https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv>

*Hint 2* Here is a table that includes goal names: <https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv>

*Hint 3* Which region has all NA values for scores?

*Answer*

**Additional Concerns:**

-    **one region is missing from the region list (region_id 0)**

-   **there are only 10 goals listed in the data set with goal names, but there appear to be 19 in the scores data set.**

```{r, include=FALSE}
#check out the scores data
head(scores)

#look at the number of unique region ids
length(unique(scores$region_id))

#check out the region data
head(region_list)

#check that there are the same number of region ids (appear to be one less)
length(unique(region_list$rgn_id))


head(goal_names)

#join them together 
scores_expanded <- scores %>% 
  left_join(region_list, by = c("region_id" = "rgn_id")) %>% 
  left_join(goal_names, by = c("goal" = "Abbreviation")) %>% 
  rename("goal_name" = "Goal", "goal_abbreviation" = "goal") %>% 
  

length(unique(scores_expanded$Goal))
length(unique(scores$goal))

#for each unique region, check for scores that are not NA


```

*Question*

Which region id corresponds to the United States?

**There appears to be both a region id and an admin region id. The admin region id is 163. However within that admin_region id, there are other region ids, presumably areas that are not technically within the US but are territories of the us. These region ids are printed below.**

```{r}
#filter to the us
united_states <- scores_expanded %>% filter(str_detect(admin_country_name, "United States"))

#print the region id
unique(united_states$admin_rgn_id)


all_us_regions <- c(unique(united_states$region_id))
#look at all of the US region ids 
cat("Additional region_ids associated with the United states admin region include", all_us_regions)

```

### Goal explore

*Question* When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?

*Hint 1* Review the methods doc here: <https://raw.githack.com/OHI-Science/ohi-global/published/documents/methods/Supplement.html>

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space


```

### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend). When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space

# unique(scores$dimension)
# table(scores$dimension)
# 
# table(scores$goal)
# table(scores$goal, scores$dimension)


```

### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2021. Which ones have the least? Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2022, and cut region_id = 0.

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space


```

### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space


```

### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable. Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table. NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

| Variable  | Description   | Values       |
|-----------|---------------|--------------|
| goal      |               |              |
| dimension |               |              |
| region_id |               |              |
| score     |               |              |
| year      | Scenario year | 2012 to 2022 |

```{r, include=FALSE}
## Working space

```

### Plot

*Project* Create a scatterplot that compares 2012 and 2022 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time?

*Answer* Write your answer here!

```{r, include=FALSE}
## Working space



```

Create a plot that shows the distribution (e.g., histogram or barplot or ??) of *Index* scores of all countries.

```{r across year plot}

## Code to create a plot showing distribution of country Index scores



```
