---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages.. you might need to install them first
library(dplyr)
library(tidyr)
library(ggplot2)
```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Melanie or Gage"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You can ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

*Note* There are scores for 2022, you can ignore these. These are just here as a placeholder until we calculate new ones for 2022 (they are the 2021 scores). 

```{r, include=FALSE}
## Working space
scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220? Can you make this data more human readable by including the full goal name (vs. the 2 letter abbreviation) and the country name.

*Hint 1* Here is a link to the official region list (the rgn_id column in this file matches the region_id column in the scores data): https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Here is a table that includes goal names: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv

*Hint 3* Which region has all NA values for scores?

```{r, include=FALSE}
## Working space
regions <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv")
goals <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv")

scores_full <- scores %>% 
  left_join(regions, by = c("region_id" = "rgn_id")) %>% 
  left_join(goals, by = c("goal" = "Abbreviation"))

num_rgns <- length(unique(scores_full$region_id))

all_na_rgn <- scores_full %>% 
  group_by(region_id, rgn_name) %>% 
  summarise(score = mean(score, na.rm = T)) %>% 
  filter(is.na(score)) %>% 
  pull(rgn_name)
  
usa <- scores_full %>% 
  filter(rgn_name == "United States") %>% 
  pull(region_id) %>% 
  unique()
```

*Answer* Write your answer here!

The region that has all NA values for scores is `r all_na_rgn`. There is also a note that says Antartica is not included in the global assessment. I also see there is a region 0 in the scores data but it does not have any metadata in the regions data frame.

*Question* Which region id corresponds to the United States?

The region id that corresponds to the United States is `r usa`

### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/documents/methods/Supplement.html

*Answer* Write your answer here!

At first it looked like the file from the hint above lists only the 10 goals. Looking at the github repo, the other file in that folder "Subgoal_Descriptions.csv" has 8 Subgoals listed. Together these two files have all 18, and presumably the 19th would be NA values. I then realized that there is a column `goal` and a column `Goal` and the little g goal had 19 values. These would be the 10 primary goals, 8 subgoals, and the Index itself.

```{r, include=FALSE}
## Working space
sub_goals <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/Subgoal_Descriptions.csv")

length(unique(scores_full$goal))

length(unique(goals$Goal))
length(unique(sub_goals$Subgoal))
```

### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  

*Answer* Write your answer here!

The index goals adds another 2442 values, the exact difference between scores and status. The index score is calculated differently than the other goals.

```{r, include=FALSE}
## Working space

unique(scores$dimension)
table(scores$dimension)

table(scores$goal)
table(scores$goal, scores$dimension)

```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2021.  Which ones have the least? Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2021, and cut region_id = 0.

*Answer* Write your answer here!

The top goal with missing scores in 2021 is carbon storage and the top subgoal is mariculture. The lack of data for mariculture makes sense as many places may not have any mariculture. There are many with no missing scores, likely due to these goals being widespread with available data.

```{r, include=FALSE}
## Working space
na_count_by_rgn <- scores_full %>% 
  filter(dimension == "score",
         region_id > 0, region_id != 213,
         year == 2021) %>% 
  group_by(goal) %>% 
  summarise(na_n = sum(is.na(score))) %>% 
  # filter(na_n >0) %>% 
  arrange(desc(na_n))

na_count_by_rgn %>% 
  head(10)
na_count_by_rgn %>% 
  tail(10)
```


### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* Write your answer here!

The score would be 85 which is found by taking the average of the future status and the status. 

```{r, include=FALSE}
## Working space

scores_full$dimension %>% unique()

scores_summary <- scores_full %>% 
  filter(region_id == 63, 
         goal == "AO",
         dimension %in% c("future", "status", "score"),
         year == 2012) %>% 
  group_by(dimension, year) %>% 
  summarise(score_summary = mean(score)) %>% 
  arrange(year)
scores_summary
round((scores_summary$score_summary[1] + scores_summary$score_summary[3])/2,2) == scores_summary$score_summary[2]
```

### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).


```{r echo = FALSE}
## Working space
dplyr::tibble(
  Variable = c(
    "goal", 
    "dimension", 
    "region_id", 
    "score", 
    "year"),
  Description = c(
    "10 publicly held goals and 8 subgoals, along with the index itself",
    "Dimensions used to calculate OHI goal scores",
    "Global regions defined by each contries EEZ",
    "index scores for each goal",
    "Scenario year"),
  Values = c(
    list(unique(scores_full$goal)),
    list(unique(scores_full$dimension)),
    paste(min(scores_full$region_id, na.rm = T), "to", max(scores_full$region_id, na.rm = T)),
    paste(min(scores_full$score, na.rm = T), "to", max(scores_full$score, na.rm = T)),
    paste(min(scores_full$year, na.rm = T), "to", max(scores_full$year, na.rm = T)))
) %>% DT::datatable()
```


### Plot

*Project* Create a scatterplot that compares 2012 and 2021 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* Write your answer here!

My plot is hard to interpret but it looks generally like some areas have increased artisinal opportunities over time. The mean values are very close and  

```{r, echo=FALSE}
## Working space
scores_comp <- scores_full %>% 
  filter(year %in% c(2012, 2021),
         goal == "AO", 
         dimension == "score",
         region_id > 0) %>% 
  group_by(year) %>% 
  mutate(mean_score = mean(score, na.rm = T))
idx_scores <- scores_full %>% 
  filter(
    goal == "Index" 
    )
mean(filter(scores_comp, year == 2012)$score, na.rm = T)
mean(filter(scores_comp, year == 2021)$score, na.rm = T)
```


Create a plot that shows the distribution (e.g., histogram or barplot or ??) of *Index* scores of all countries.

```{r across year plot}
## Code to create a plot showing distribution of country Index scores
ggplot(data = scores_comp, aes(x = region_id, y = score, color = region_id)) +
  geom_point() +
  facet_wrap(~year)+
  geom_hline(aes(yintercept = mean_score))

ggplot(data = idx_scores, aes(x = score, fill = region_id)) +
  geom_histogram(bins = 100)
```

