---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages.. you might need to install them first
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Melanie or Gage"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You can ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

*Note* There are scores for 2022, you can ignore these. These are just here as a placeholder until we calculate new ones for 2022 (they are the 2021 scores). 

```{r, include=FALSE}
## Working space
scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")
```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220? Can you make this data more human readable by including the full goal name (vs. the 2 letter abbreviation) and the country name.

*Hint 1* Here is a link to the official region list (the rgn_id column in this file matches the region_id column in the scores data): https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Here is a table that includes goal names: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv

*Hint 3* Which region has all NA values for scores?

*Answer* Antarctica (region 213) has all NA values for scores



```{r, include=FALSE}
## Working space
length(unique(scores$region_id))

regions <-  read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv")
goals <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/metadata_documentation/ohi_model/tables/ten_goals.csv")

scores_regions_goals <- scores %>% 
  left_join(regions, by = c("region_id" = "rgn_id")) %>% 
  left_join(goals, by = c("goal" = "Abbreviation"))

scores_regions_goals %>% 
  group_by(region_id) %>% 
  summarize(mean_score = mean(score, na.rm = TRUE)) %>% 
  filter(is.na(mean_score))
```

*Question* Which region id corresponds to the United States?

```{r}
regions %>% filter(rgn_name == "United States")
```
Region id 163 corresponds to the United States.


### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/documents/methods/Supplement.html

*Answer* This includes all goals, subgoals, and the Index score. The abbreviations with 2 letters correspond to the 10 main goals and the three letter abbreviations correspond to subgoals.



```{r, include=FALSE}
## Working space

length(unique(scores$goal))
unique(scores$goal)
```


### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  

*Answer* The difference between the number of `score` and `status` values is explained by the `Index` values which are `score` values calculated based on all the goals and thus don't have a status associated with them.

```{r, include=FALSE}
## Working space

# unique(scores$dimension)
# table(scores$dimension)
# 
# table(scores$goal)
table(scores$goal, scores$dimension)

# Calculating difference between "score" instances and "status" instances
length(scores$dimension[scores$dimension == "score"]) - length(scores$dimension[scores$dimension == "status"])
```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2021. Which ones have the least? Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2021, and cut region_id = 0.

*Answer* The table printed below shows the goals and subgoals with the most missing score values in 2021 - the top culprtis being CS and NP for goals, and MAR for subgoals. The half not shown below have zero missing values. The goals and subgoals with the most missing values seems like they might be more difficult to get reliable data for in particular regions. 


```{r, include=FALSE}
## Working space


most_na <- scores_regions_goals %>% 
  filter(dimension == "score" & year == 2021 & region_id != 0 & region_id != 213) %>% 
  filter(is.na(score)) %>% 
  group_by(goal) %>% 
  count() %>% 
  ungroup()

slice_max(most_na, order_by = n, n = nrow(most_na))
```


### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* Based on isolating one region and one goal, it looks like current status and future status are averaged to get the score, so the score for the goal in question would be 85.

```{r, include=FALSE}
## Working space

colombia <- scores_regions_goals %>% 
  filter(rgn_name == "Colombia" & goal == "BD" & dimension %in% c("status", "score", "future")) %>% 
  select(dimension, score, year, goal)
```


### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

Variable   | Description                  | Values
---------- | ---------------------------- | ------------------------------
goal       | Goal or subgoal abbreviation | AO, BD, CP, CS, CW, ECO, FIS, FP, HAB, ICO, Index, LE, LIV, LSP, MAR, NP, SP, SPP, TR
dimension  | Component of score           | future, pressures, resilience, score, status, trend
region_id  | Unique region identifier     | 0 to 250
score      | Numerical value of dimension | -1 to 100         
year       | Scenario year                | 2012 to 2022


```{r, include=FALSE}
## Working space

scores$goal %>% 
  unique()

scores$dimension %>% 
  unique()

range(scores$region_id)

range(scores$score, na.rm = TRUE)
```


### Plot

*Project* Create a scatterplot that compares 2012 and 2021 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* It's hard to tell exactly from this plot, but it looks like there is a higher concentration of regions that scored at or above 90 in 2021 compared to 2012.

```{r, include=FALSE}
## Working space

AO <- scores_regions_goals %>% 
  filter(goal == "AO" & dimension == "score" & year %in% c(2012, 2021))

ggplot(AO, aes(x = region_id, y = score, color = score)) + 
  geom_point() +
  facet_wrap("year")
```


Create a plot that shows the distribution (e.g., histogram or barplot or ??) of *Index* scores of all countries.

```{r across year plot}

## Code to create a plot showing distribution of country Index scores


index <- scores %>% 
  filter(goal == "Index" & dimension == "score" & year == 2022)

ggplot(index, aes(x = score)) + 
  geom_histogram()
```




