---
title: "Livelihoods and Economies"
output: html_document
date: "2023-08-18"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script was created in 2023 to progress towards updating the livelihoods and economies goal.

It is used to read in and wrangle all available livelihoods and economies data as of August 2023. When newer updated data is not available, the original data used in the livelihoods and goal in 2013 is included. For more details on what data is available, please refer to the spreadsheet linked in the livelihoods and economies issue.

The second half of the document focuses on developing the methods to create the data layers. This script can be developed and completed in future years.

# Setup

```{r}
# load libraries, set directories
library(ohicore)  #devtools::install_github('ohi-science/ohicore@dev')
library(tidyverse)
library(plotly)
library(here)
library(janitor)
library(pdftools)
source(here('workflow/R/common.R')) # directory locations

source(here('workflow/R/fao_fxn.R')) # function for cleaning FAO data

version_year <- 2023
```

# Read in Data

```{r}
# OECD Data 
oecd_raw <- read_csv(file.path(dir_M, "/git-annex/globalprep/_raw_data/OECD/d2023/ocean_economy.csv")) %>% 
  select(Country, Year, VARIABLE, Variable, Unit, Value, PowerCode) 

#ilostat data
ilo_raw <- read_csv(file.path(dir_M, "/git-annex/globalprep/_raw_data/ILOSTAT/d2023/ILO_earnings_economic.csv"))

#fao commodities data 
commodities_value <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/FAO_commodities/d2023/FAO_raw_commodities_value_1976_2021.csv")) %>% 
  janitor::clean_names()
#we likely only want exports?
#re-download updated version of this. 

#FAO aquaculture value
aquaculture_value <- read.csv(file.path(dir_M, "git-annex/globalprep/_raw_data/FAO_mariculture/d2023/FAO_GlobalAquacultureProduction_Value_1950_2021.csv"))

#fao yearbook pdf tables
fao_pdf <- pdf_text(file.path(dir_M, "/git-annex/globalprep/_raw_data/FAO/d2023/fao_fishers_table.pdf"))

#oww
oww <- read_csv(file.path(dir_M,"/git-annex/globalprep/_raw_data/OWW/d2023/oww3.csv"))

#whale watching pdf
whale_pdf <- pdf_text(file.path(dir_M, "/git-annex/globalprep/_raw_data/IFAW_MarineMammalWatching/whale_watching_worldwide.pdf"))

#unwto -come back to tourism later when we know which data want to use

```

# Data Wrangling

## Jobs

### Turn number of fishers table from FAO yearbook into data frame and clean.

```{r}
#turn the fao pdf into a format we can use
fishers_table <- fao_pdf[[3]] %>%
  as.data.frame()

#specify columns 
columns <- c("country", "1995", "2000", "2005", "2010", "2014", "2015", "2016", "2017", "2018", "2019")

fao_fisher_jobs <- fishers_table %>%
  separate_rows('.', sep = "\n") %>% #separate rows based on line breaks
  rename(main_column = ".") %>%
  mutate(main_column = str_remove_all(main_column, "\\*")) %>% #remove the * 
  mutate(main_column = str_replace_all(main_column, "F(?!\\p{L})", " ")) %>% #replace all the Fs not followed by letters (used to indicate estimate)
  slice(7:69) %>% #remove all rows not in the table
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% #separate into columns based on double spaces
  pivot_longer(cols = -c("country"), names_to = "year") %>% 
  mutate(value = str_squish(value)) %>% #remove extra spaces
  mutate(value = str_remove_all(value, " ")) %>% 
  mutate(value = if_else(value == "…" , NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(Unit = "persons (1)") %>% 
  mutate(sector = "commerical fishing") %>% 
  mutate(variable = "FAO Number of Fishers") %>% 
  clean_names() %>% 
  select(country, year, variable, unit, value, sector)
```

### Turn number of fish farmers table from FAO yearbook into data frame and clean.

```{r}
#turn the fao pdf into a format we can use
fish_farm_table <- fao_pdf[[5]] %>%
  as.data.frame()

#specify columns
columns <- c("country", "1995", "2000", "2005", "2010", "2014", "2015", "2016", "2017", "2018", "2019")

fao_farm_jobs <- fish_farm_table %>%
  separate_rows('.', sep = "\n") %>%
  rename(main_column = ".") %>%
  mutate(main_column = str_remove_all(main_column, "\\*")) %>% #remove the * 
  mutate(main_column = str_replace_all(main_column, "F(?!\\p{L})", " "))  %>% 
  slice(7:36) %>%
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% 
  pivot_longer(cols = -c("country"), names_to = "year") %>% 
  mutate(value = str_squish(value)) %>% 
  mutate(value = str_remove_all(value, " ")) %>% 
  mutate(value = if_else(value == "…" , NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(Unit = "persons (1)") %>% 
  mutate(sector = "mariculture") %>% 
  mutate(variable = "FAO Number of Fish Farmers") %>% 
  clean_names() %>% 
  select(country, year, variable, unit, value, sector)
```

### OECD Jobs Data

```{r}
#number of people employed in th fishing sector 
#note that this is data for the same category as the FAO fishers data, one could potentially be used for gapfilling whichever is used as the main data set
oecd_fisher_jobs <- oecd_raw %>% 
  filter(VARIABLE == "FISH_EMP_MARINE_FISH_TOT") %>% 
  mutate(Unit = ("persons (1000)")) %>% 
  select(- c(PowerCode, VARIABLE))

#number of people employed in aquaculture
#note that this is not specifically marine
#note that this is data for the same category as the FAO fish farmers data, one could potentially be used for gapfilling the main data set
oecd_fishproc_jobs <- oecd_raw %>% 
  filter(VARIABLE == "FISH_EMP_PROCESSING_TOT") %>% 
  mutate(Unit = ("persons (1000)")) %>% 
  select(- c(PowerCode, VARIABLE))
```

### Marine Mammal Watching Jobs

```{r}
# from the methods we know: Jobs based on number of whale watchers in a country and a regional average number of whale watchers per employee. Includes all marine mammal watching. 

#get table with regional averages out of pdf

columns <- c("region", "job_number", "watch_per_employee")

regional_whale_employee <- whale_pdf[26] %>% 
  as.data.frame() %>% 
  separate_rows('.', sep = "\n") %>% 
  rename(main_column = ".") %>% 
  slice(23:31) %>% 
  slice(1,2,3,4, 6,7,9) %>% 
  mutate(main_column = str_trim(main_column)) %>% 
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% 
  mutate(region = case_when(str_detect(region, "Oceania") ~ "Oceania and the Pacific
Islands",
str_detect(region, "Central America") ~  "Central America and Caribbean",
T ~ region)) %>% 
  mutate(watch_per_employee = as.numeric(str_remove(watch_per_employee, ","))) %>% 
  select(-job_number)

#table with number of whale watchers

#each region has its own table in report so making a function to extract each one 
whale_pdf_convert <- function(page, slice_set, region, new_name){
  
  columns <- c("country", "1998", "2008", "growth") #create columns
  
  final_data <- whale_pdf[page] %>%  #call the entire page
  as.data.frame() %>% 
  separate_rows('.', sep = "\n") %>% 
  rename(main_column = ".") %>% 
  slice(slice_set) %>% 
  mutate(main_column = str_trim(main_column)) %>% 
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% 
    select(-c("growth")) %>% 
     pivot_longer(cols = c("1998", "2008"), names_to = "year") %>% 
    mutate(value = case_when(value == "None" ~ "0",
                           value == "Minimal"~ "250",
                           T ~ value)) %>% 
  mutate(value = str_remove_all(value, ",")) %>% 
  mutate(value = str_remove(value, " ")) %>% 
  mutate(value = ifelse(value == "N/A", NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(region = region)
  
  assign(new_name, final_data, envir = .GlobalEnv)
}

#africa and middle east
whale_pdf_convert(42, 5:26, "Africa and Middle East", "africa_data")

#europe
whale_pdf_convert(82, 5:26, "Europe", "europe_data")

#asia
whale_pdf_convert(121, 7:25, "Asia", "asia_data")

#oceania
whale_pdf_convert(157, 5:22, "Oceania, Pacific Islands and Antarctica", "oceania_data")
  
oceania_data <- oceania_data %>% filter(country != "Micronesia") %>% 
  mutate(country = str_replace(country, "Federated States of", "Federated States of Micronesia"))

#north america
whale_pdf_convert(199, 7:10, "North America", "na_data")

#north america
whale_pdf_convert(238, 5:27, "Central America and Caribbean", "ca_data")

#north america
whale_pdf_convert(269, 17:27, "South America", "sa_data")

#now that we have all of the available countries, combine into one data frame
whale_watching_numbers <- rbind(africa_data, europe_data, asia_data, oceania_data, na_data, ca_data, sa_data)


#convert whale watching numbers into number of jobs
#divide the total number of whale watchers by the number of whale watchers per employee
whale_jobs <- whale_watching_numbers %>% 
  left_join(regional_whale_employee, by = c("region")) %>% 
  mutate(jobs = value/watch_per_employee) %>% 
  mutate(variable = "Whale Watching Jobs", unit = "Persons(1)", sector = "Marine Mammal Watching") %>% 
  select(country, year, variable, unit, value = jobs, sector)
```

## Revenue

### OECD Revenue Data Wrangling

```{r}
#get the ocean energy revenue 
offshore <- oecd_raw %>% filter(VARIABLE == "RD_TOTAL_OCEAN_OFFSHORE") %>% 
  select(-VARIABLE) %>% 
  mutate(sector = "Ocean Energy") %>% 
  mutate(Unit = "USD (1000000)") %>% 
  select(-c(PowerCode)) %>% 
  clean_names()
```

### FAO Raw commodities data wrangling

```{r}
#ornamental fishing revenue 
ornamental_value <- commodities_value %>%
  filter(str_detect(commodity_name, "Ornamental")) # we only want ornamental fish

#commercial fishing revenue: 
#since it's revenue do we only want exports? 
#would all other fish besides aquarium fish count for this? 
commercial_value <- commodities_value %>% filter(str_detect(commodity_name, "Ornamental"))
```

### FAO aquaculture data wrangling

```{r}

#mariculture revenue 
maricutlure_revenue <- aquaculture_value %>% 
   dplyr::rename(country = "Country (Name)",
                species = "ASFIS species (Name)",
                area = "FAO major fishing area (Name)",
                environment = "Environment (Name)") %>%
  dplyr::select(-c("Unit (Name)")) %>% #added removing Unit also
  dplyr::rename_with(~ base::gsub("\\[", "", .)) %>% 
  dplyr::rename_with(~ base::gsub("\\]", "", .)) %>% 
  filter(environment %in% c("Brackishwater", "Marine")) %>% 
  pivot_longer(cols = "1984":"2021", names_to = "year", values_to = "value") %>% 
  fao_clean_data_new() #do we keep all non-freshwater mariculture for revenue?


#apply the new split function when done, to split up:netherlands antilles, bonaire, and channel islands 

mar_rgn <- name_2_rgn(df_in = maricutlure_revenue, fld_name='country')

#sum by region and year (we need total revenue)

total_mar_rev <- mar_rgn %>%
  group_by(rgn_id, rgn_name, year) %>% 
  summarize(value = sum(value, na.rm = T), Unit = first(Unit))
```

### Whale Watching revenue

The whale watching revenue is stored in a separate table for each country throughout the paper. We will create function to extract each of the tables and then combine them.

Key takeaways from the OHI methods (7.19.0.4 Marine mammal
watching) to guide wrangling.

-    total expenditures are used as a close proxy for total revenue.

-   We used total expenditure data (direct and indirect expenditures) to avoid using a literature derived multiplier effect.

-   When IFAW reported "minimal" revenue from whale watching, we converted this description to a 0 for lack of additional information.

-   For countries with both marine and freshwater cetacean viewing, we adjusted by the proportion of marine revenue as described for the jobs dataset.

```{r}
#country <- country_list[74] # for testing

#test <- whale_extract(country)

#create a function to find all of the country tables with revenue and extract
whale_extract <- function(country){

#the tables don't match the way all of the countries are written 
#this replaces them with the way they are written in the pdf when needed
replacements <- c("The Gambia" = "Gambia, The",
                  "Portugal – Madeira Archipegalo" = "Portugal ‐ Madeira Archipelago", "China ‐ Mainland" = "China (Mainland)", "Myanmar" = "Myanmar (Burma)", "Guadeloupe and islands" = "Guadeloupe and islands (including St. Martin and St. Barthélemy)", "Falkland Islands" = "Falkland Islands (Las Malvinas)", 
"Georgia, Ukraine – Black Sea" = "Georgia, Ukraine and Russia – Black Sea", "St. Vincent and the Grenadines" = "St Vincent and the Grenadines", "Netherlands Antilles" = "Netherlands Antilles – Aruba, Bonaire, Curaçao and St. Maarten")

country_new <- str_replace_all(country, replacements)

# Escape special characters in the country name
country_new <- str_replace_all(country_new, "[\\(\\)]", "\\\\\\0") #remove special characters


match_character <- paste0(country_new, "\n") # add /n to help find the locations 

matches <- gregexpr(match_character, whale_pdf) #search the pdf for the country

#set each item to true or false depending on if it was a match
matches_good <- sapply(matches, function(match) any(match != -1))

#get the page number with the table (where match is true)
page_number_original <- which(matches_good)

#if there is more than one match use another method to narrow it down
if(length(page_number_original) >1) {
 updated_match <- paste(country_new, "Year")
  #search again
 page_search <- whale_pdf[page_number_original]
 
 squish <- str_squish(page_search)
 
  matches <- gregexpr(updated_match, squish)
  matches_good <- sapply(matches, function(match) any(match != -1))

#get the page number with the table
page_number_new <- which(matches_good)

#update the page number to just the correct one
page_number_original <- page_number_original[page_number_new]

#if it is still greater than one then hand select which one we want
if (length(page_number_original) > 1){
  
  page_number_original <- subset(page_number_original, page_number_original %in% c("118", "114", "160", "162", "194"))
  
}

}

#subset to the page with the table 
page <- whale_pdf[page_number_original]

#set columns 
column_jobs <- c("Year", "Number of whale watchers", "AAGR", "Number of operators",   "Direct expenditure", "Indirect expenditure", "Total Expenditure")

#turn the page into a table 
data <- page %>% 
  as.data.frame() %>% 
  separate_rows('.', sep = "\n") %>% 
  rename(main_column = ".") %>% 
  mutate(main_column = str_trim(main_column)) %>% 
  separate(main_column, into = column_jobs, sep = " {2,}", extra = "merge")

# Find the index of the row with value of country
index <- which(data$Year == country_new)
if(length(index) == 0){
  index <- 1
}

#filter the page to just the table 
data_clean <- data %>%
  slice((index + 1):n()) %>% #remove any other tables above 
  filter(Year %in% c("1991", "1994", "1998", "2004", "2005", "2006", "2007", "2008")) %>% 
  mutate(first_2008_index = min(which(Year == 2008, arr.ind = TRUE))) %>%
  filter(ifelse(is.na(first_2008_index), TRUE, row_number() <= first_2008_index)) %>%
  select(-first_2008_index) %>% 
  clean_names() %>% 
  mutate_all(~ str_replace_all(., ",", "")) %>% 
  mutate_all(~ str_replace_all(., "\\$", "")) %>% 
  mutate_all(~ str_replace_all(., "None", "0")) %>% 
  mutate_all(~ str_replace_all(., "Minimal", "0")) %>% 
  mutate(country = country) #add a column with the country

return(data_clean)
}

#identify the countries based on data we created earlier
country_list <- unique(whale_jobs$country)


whale_list <- list()
for (i in seq_along(country_list)) {
  
  country <- country_list[i]
  data <- whale_extract(country)
  
  whale_list[[i]] <- data 
  
}

whale_revenue_data <- bind_rows(whale_list) %>% 
  mutate_all(~ str_replace_all(., "N/A", NA_character_)) %>% 
  mutate(direct_expenditure = as.numeric(direct_expenditure),
         indirect_expenditure = as.numeric(indirect_expenditure),
         total_expenditure = as.numeric(total_expenditure)) %>% 
  mutate(unit =, value = , sector = ) %>%
```

### Tourism Revenue Data Wrangling

(Placeholder)

## Wage Data

```{r}

```
