---
title: "OHI `r format(Sys.Date(), '%Y')` - saltmarsh trend"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script generates the pressure incurred from invasive species for each OHI region. 

## Updates from previous assessment

2022 - Updating the data with new dataset!

***

## Data Source 

**Reference**: Pagad, S., Genovesi, P., Carnevali, L. et al. Introducing the Global Register of Introduced and Invasive Species. Sci Data 5, 170202 (2018). https://doi.org/10.1038/sdata.2017.202

**Downloaded**: 2022-08-04

**Description**:  
Harmonised, representative data on the state of biological invasions remain inadequate at country and global scales, particularly for taxa that affect biodiversity and ecosystems. Information is not readily available in a form suitable for policy and reporting. The Global Register of Introduced and Invasive Species (GRIIS) provides the first country-wise checklists of introduced (naturalised) and invasive species. GRIIS was conceived to provide a sustainable platform for information delivery to support national governments. We outline the rationale and methods underpinning GRIIS, to facilitate transparent, repeatable analysis and reporting. Twenty country checklists are presented as exemplars; GRIIS Checklists for close to all countries globally will be submitted through the same process shortly. Over 11000 species records are currently in the 20 country exemplars alone, with environmental impact evidence for just over 20% of these. GRIIS provides significant support for countries to identify and prioritise invasive alien species, and establishes national and global baselines. In future this will enable a global system for sustainable monitoring of trends in biological invasions that affect the environment.

**Time range**: 1999-2019

**Download link**: 
????

**Variables**:
- ???

***

# Methods

## Overview

1. saaaaa

## Setup

``` {r setup, eval = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE)

if (!require(librarian)){install.packages("librarian")}
if (!require(ohicore)){devtools::install_github('ohi-science/ohicore@dev')}

librarian::shelf(
  tidyverse,
  here,
  janitor,
  sf,
  plotly,
  countrycode,
  ohicore
) 
### directory paths and relevant files
current_year <- 2022
version_year <- paste0("v", current_year)
data_year <- paste0("d", current_year)
# source(here::here('workflow', 'R', 'common.R'))

### Mazu
dir_here  <- here::here('globalprep', 'prs_alien', version_year)
dir_data <- file.path(dir_M, 'git-annex', 'globalprep', '_raw_data', 'griis', data_year)

# rgns <- ohicore::georegion_labels # Kiribati is broke v2022
rgns <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep/master/globalprep/spatial/v2017/output/georegion_labels.csv")

rgns_eez <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2022/gh-pages/globalprep/spatial/v2017/output/rgn_labels.csv") %>% 
  dplyr::filter(type == "eez") %>% 
  dplyr::select(-type, rgn_name = label)
```

## Download the data

```{bash}
mkdir /home/shares/ohi/git-annex/globalprep/_raw_data/griis/d2022 && cd $_

wget --recursive --no-parent --level 1 --accept zip http://griis.org/download

mv griis.org/assets/checklists/*.zip .

rm -r griis.org

for FILE in *; do 
  echo $FILE 
  unzip $FILE 
  rm *.txt
done 

rm *.zip
```

## Clean raw data

Read in country codes from the `countrycode` package.

```{r}
codes <- countrycode::codelist %>% 
  janitor::clean_names() %>% 
  dplyr::select(country = country_name_en, country_code = iso2c) %>% 
  tidyr::drop_na()
```

Loop through the raw data files and join them into one large file. 

```{r}
files <- list.files(dir_data, full.names = T)

output <- tibble("id" = character())

for (file in files) {
  tmp <- read_csv(file, col_types = cols(.default = "c")) 
  output <- full_join(output, tmp)
}
```

Make the data consistent and tidy. 

```{r}
df <- output %>%
  janitor::clean_names() %>% 
  dplyr::select(
    id, taxon_id, scientific_name,kingdom, phylum, class, order, 
    family, taxon_rank, taxonomic_status, is_invasive, habitat,
    location_id, country_code, occurrence_status, establishment_means) %>% 
  dplyr::left_join(codes) %>% 
  dplyr::mutate(
    habitat = tolower(habitat),
    habitat = str_replace_all(habitat, "/", "|"),
    habitat = str_replace_all(habitat, "terrestriali", "terrestrial\\|"),
    habitat = str_replace_all(habitat, "terrestrial,", "terrestrial"),
    habitat = str_replace_all(habitat, "freshwatetr", "freshwater"),
    habitat = str_replace_all(habitat, "terrestrial \\|freshwater \\|brackish", "terrestrial\\|freshwater\\|brackish"),
    habitat = str_replace_all(habitat, "freshwater\\|marine", "marine\\|freshwater"),
    habitat = str_replace_all(habitat, "brackish\\|marine\\|freshwater", "marine\\|freshwater\\|brackish"),
    habitat = str_replace_all(habitat, "freshhwater\\|brackish\\|marine", "marine\\|freshwater\\|brackish"),
    habitat = str_replace_all(habitat, "freshwater\\|brackish\\|marine", "marine\\|freshwater\\|brackish"),
    habitat = str_replace_all(habitat, "freshwater\\|brackish", "brackish\\|freshwater"),
    establishment_means = case_when(establishment_means == "Present" ~ occurrence_status, T ~ establishment_means),
    establishment_means = tolower(establishment_means),
    establishment_means = str_replace_all(establishment_means, "/", "|"),
    establishment_means = str_replace_all(establishment_means, "uncerain", "uncertain"),
    establishment_means = str_replace_all(establishment_means, "cryptogenic\\|", ""),
    establishment_means = str_replace_all(establishment_means, "native\\|", ""),
    establishment_means = replace_na(establishment_means, "uncertain"),
    is_invasive = tolower(is_invasive),
    is_invasive = str_replace_all(is_invasive, "null", "0"),
    is_invasive = str_replace_all(is_invasive, "false", "0"),
    is_invasive = str_replace_all(is_invasive, "to be evaluated", "0"),
    is_invasive = replace_na(is_invasive, "0"),
    is_invasive = str_replace_all(is_invasive, "yes", "1"),
    is_invasive = str_replace_all(is_invasive, "invasive", "1"),
    is_invasive = as.numeric(is_invasive),
    country_code = toupper(country_code),
    country_code = case_when(location_id == "Namibia" ~ "NA", T ~ country_code),
    country = str_replace_all(country, "St. Lucia", "Saint Lucia"),
    country = str_replace_all(country, "&", "and"),
    country = str_replace_all(country, "Congo - Kinshasa", "Democratic Republic of the Congo"),
    country = str_replace_all(country, "Côte d’Ivoire", "	Ivory Coast"),
    country = str_replace_all(country, "Curaçao", "Curacao"),
    country = str_replace_all(country, "Myanmar \\(Burma\\)", "Myanmar"),
    country = str_replace_all(country, "Saint Martin \\(French part\\)", "Northern Saint-Martin"),
    country = str_replace_all(country, "São Tomé and Príncipe", "Sao Tome and Principe"),
    country = str_replace_all(country, "St. Vincent and Grenadines", "Saint Vincent and the Grenadines"),
    country = str_replace_all(country, "Svalbard & Jan Mayen", "Jan Mayen"),
    country = case_when(
      country_code == "EC-W" ~ "Ecuador", # NEEDS to be first
      is.na(country) & !is.na(location_id) ~ location_id,
      country_code == "BQ-BO" ~ "Bonaire",
      country_code == "BQ-SE" ~ "Sint Eustatius",
      country_code == "BQ-SA" ~ "Saba",
      country_code == "FM-PNI" ~ "Micronesia",
      country_code == "FM-TRK" ~ "Micronesia",
      country_code == "YE-SU" ~ "Yemen",
      T ~ country)) %>% 
  dplyr::filter(!habitat %in% c("terrestrial", "freshwater", "terrestrial|freshwater", "freshwater|terrestrial", "host", NA)) %>% 
  dplyr::select(-c(location_id)) %>% 
  dplyr::left_join(rgns, by = c('country' = 'rgn_label')) %>% 
  dplyr::select(-rgn_id)
```

## Match countries to OHI regions


```{r}
df_ohi <- ohicore::name_2_rgn(df_in = df)

# v2022
# These data were removed for not having any match in the lookup tables:
# French Southern Territories (will be split up later)   North Macedonia (landlocked)   St. Barthélemy (IDK lol)
```

Break up French Southern Territories into 8 regions and Kiribati into 2, each repeating the data of the larger region. 

```{r}
### Report these regions at higher spatial resolution:
french_territories <- c(
    'Glorioso Islands', 'Juan de Nova Island', 'Bassas da India', 
    'Ile Europa', 'Ile Tromelin', 'Crozet Islands',
    'Amsterdam Island and Saint Paul Island', 'Kerguelen Islands')

kiribati <- c(
  "Line Islands (Kiribati)", "Phoenix Islands (Kiribati)", "Gilbert Islands (Kiribati)")

country_split_data <- dplyr::tibble(
  country = c(rep("French Southern Territories", 8),
              rep("Kiribati", 3)), 
  region = c(french_territories, kiribati)) %>%
  dplyr::left_join(df) %>%
  dplyr::select(-c(country, r0_label, r1_label, r2_label)) %>%
  dplyr::rename(country = region) %>% 
  dplyr::left_join(rgns, by = c('country' = 'rgn_label')) %>% 
  dplyr::mutate(rgn_name = country) 
```

Bind French Southern Territories and kiribati back to the full dataset. 

```{r}
alien_sp_df <- rbind(df_ohi, country_split_data) %>%
  dplyr::select(-country) %>% 
  dplyr::filter(rgn_name != 'Kiribati') %>% 
  dplyr::distinct()
```

Summarize the number of alien/invasive species to each region and fill missing regions with NA values. 

```{r}
alien_sp_summary <- alien_sp_df %>% 
  dplyr::group_by(rgn_id, rgn_name) %>%
  dplyr::mutate() %>% 
  dplyr::summarise(spp_count = n() - sum(is_invasive),
                   invasive_count = sum(is_invasive)) %>% 
  dplyr::full_join(rgns, by = c("rgn_id", 'rgn_name' = 'rgn_label')) %>% 
  dplyr::ungroup() 
```

## Gapfilling

Gapfill values from regional averages.

### Step 1

First try "r2 regions" (i.e. Caribbean countries filled with Caribbean islands mean).

```{r, eval = F}
gf_step_1 <- alien_sp_summary %>% 
  dplyr::mutate(
    r0_spp = mean(spp_count, na.rm = T),
    r0_inv = mean(invasive_count, na.rm = T)) %>% 
  dplyr::group_by(r1_label) %>% 
  dplyr::mutate(
    r1_spp = mean(spp_count, na.rm = T),
    r1_inv = mean(invasive_count, na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(r2_label) %>% 
  dplyr::mutate(
    r2_spp = mean(spp_count, na.rm = T),
    r2_inv = mean(invasive_count, na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    has_data       = ifelse(is.na(spp_count), 0, 1),
    gapfilled      = ifelse(is.na(spp_count), NA, "no"),
    method         = ifelse(is.na(spp_count), NA, "not gapfilled"),
    spp_count      = ifelse(is.na(spp_count), r2_spp, spp_count),
    invasive_count = ifelse(is.na(invasive_count), r2_inv, invasive_count),
    gapfilled      = ifelse(has_data == 0 & !is.na(spp_count), "yes", gapfilled),
    method         = ifelse(has_data == 0 & !is.na(spp_count), "used r2 average", method))
```

In v2022, everything was filled with step 1. Step 2 and 3 were written just in case step 1 stops filling everything. 

### Step 2

Then try "r1 regions" (i.e. "Oceania" or "Asia").

```{r eval = F}
gf_step_2 <- gf_step_1 %>% 
  dplyr::mutate(
    spp_count      = ifelse(is.na(spp_count), r1_spp, spp_count),
    invasive_count = ifelse(is.na(invasive_count), r1_inv, invasive_count),
    gapfilled      = ifelse(has_data == 0 & !is.na(spp_count) & is.na(gapfilled), "yes", gapfilled),
    method         = ifelse(has_data == 0 & !is.na(spp_count) & is.na(method), "used r1 average", method))
```

### Step 3

If there remain any that can't be gapfilled in one of the regions, gapfill with the world score.

__Note:__ This is an aggressive strategy but also an unlikely scenario that will ensure complete data

```{r eval = F}
gf_step_3 <- gf_step_2 %>% 
  dplyr::mutate(
    spp_count      = ifelse(is.na(spp_count), r0_spp, spp_count),
    invasive_count = ifelse(is.na(invasive_count), r0_inv, invasive_count),
    gapfilled      = ifelse(has_data == 0 & !is.na(spp_count) & is.na(gapfilled), "yes", gapfilled),
    method         = ifelse(has_data == 0 & !is.na(spp_count) & is.na(method), "used r0 average", method),
    year = 2022)
```

## Calculate the pressure

```{r}
min_spp_count <- min(gf_step_3$spp_count)
min_inv_count <- min(gf_step_3$invasive_count)

max_spp_count <- max(gf_step_3$spp_count)
max_inv_count <- max(gf_step_3$invasive_count)

spp_weight <- 0.1
inv_weight <- 0.9

pressure <- gf_step_3 %>% 
  dplyr::group_by(rgn_id, rgn_name, year) %>% 
  dplyr::summarise(
    value = (((spp_count-min_spp_count)/(max_spp_count-min_spp_count)) * spp_weight) + 
      (((invasive_count-min_inv_count)/(max_inv_count-min_inv_count)) * inv_weight)) %>% 
  dplyr::ungroup()
```

## Save the prepped data

```{r, eval = F}
## save gapfilling flags 
gf_step_3 %>%
  dplyr::select(rgn_id, rgn_name, year, gapfilled, method) %>%
  readr::write_csv("prs_alien_gf.csv")

## save value dataset
pressure %>%
  dplyr::select(rgn_id, year, value) %>% 
  # readr::write_csv(here::here("globalprep", "prs_alien", version_year, "output", "prs_alien.csv"))
  readr::write_csv(here::here(dir_here, "output", "prs_alien.csv"))
```

## Datacheck

Lets compare to 2013 data. 

```{r, eval = F}
new_data <- here::here(dir_here, "output", "prs_alien.csv") %>% 
  readr::read_csv() %>%
  dplyr::left_join(rgns_eez)

old_data <- here::here('globalprep', 'prs_alien', 'v2013', 'data', "p_sp_alien_2013a.csv") %>% 
  readr::read_csv() %>%
  dplyr::rename(value = pressures.score) %>% 
  dplyr::left_join(rgns_eez)

compare <- new_data %>%
  dplyr::bind_rows(old_data) %>% 
  tidyr::pivot_wider(names_from = year, values_from = value)

plot_diff <- 
  ggplot2::ggplot(
    compare, ggplot2::aes(x = `2022`, y = `2008`, text = rgn_name, label = rgn_id), color = "black") +
  ggplot2::geom_jitter(width = 0.025, height = .025) +
  ggplot2::geom_abline() +
  ggplot2::labs(title = "Invasive species pressure") +
  ggplot2::theme_bw() 

plotly::ggplotly(plot_diff, tooltip = c("rgn_id", "rgn_name", "x", "y"))
```
