---
title: 'OHI 2021: Mangrove trend'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

Old methods: 
We used Hamilton and Casey’s (2016) mangrove cover data to estimate the proportional yearly change in mangrove area using a linear regression model of the most recent five years of data (i.e., slope divided by data from the earliest year included in the regression model). The slope was then multiplied by five to get the predicted change in 5 years. The original mangrove data are provided yearly (2000-2012) at 30m raster cell resolution (with the estimated area of mangrove cover in each cell).

## Updates from previous assessment
Using a new dataset from global mangrove watch. 

***
## Data Source 

**Downloaded**: 03/09/2021

**Description**:  
Global Mangrove Watch (1996 - 2016)	
https://data.unep-wcmc.org/datasets/45
Reported at spatial cell scale. 

The GMW aims to provide geospatial information about mangrove extent and
changes to the Ramsar Convention, national wetland practitioners, decision makers
and NGOs. It is part of the Ramsar Science and Technical Review Panel (STRP) work
plan for 2016-2018 and a Pilot Project to the Ramsar Global Wetlands Observation
System (GWOS), which is implemented under the GEO-Wetlands Initiative. The
primary objective of the GMW has been to provide countries lacking a national
mangrove monitoring system with first cut mangrove extent and change maps, to
help safeguard against further mangrove forest loss and degradation.
The GMW has generated a global baseline map of mangroves for 2010 using ALOS
PALSAR and Landsat (optical) data, and changes from this baseline for seven epochs
between 1996 and 2017 derived from JERS-1, ALOS and ALOS-2. Annual maps are
planned from 2018 and onwards

**Time range**: 1996-2016


***
# Methods


## Setup
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(readr)      # for read_csv()
library(raster)
library(here)
library(sf)
library(fasterize)
library(tidyverse)
library(mapview)
library(sp)
library(rgeos)

source(file.path('~/github/ohiprep_v2021/workflow/R/common.R'))

goal     <- 'globalprep/hab_mangrove/v2021'
dir_git  <- file.path('~/github/ohiprep_v2021', goal)
dir_wcmc <- file.path(file.path(dir_M, 'git-annex/globalprep/_raw_data/wcmc_mangrove'))
ohi_rasters() # call the region zones raster
regions_shape()
region_data()

```

```{r}
gmw_ohi_2016 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2016.csv")))
sum(gmw_ohi_2016$km2) # 136804.3

gmw_ohi_1996 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_1996.csv")))
sum(gmw_ohi_1996$km2) # 142919.7

gmw_ohi_2007 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2007.csv")))
sum(gmw_ohi_2007$km2) # 139059.1

gmw_ohi_2008 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2008.csv")))
sum(gmw_ohi_2008$km2) # 139264.3

gmw_ohi_2009 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2009.csv")))
sum(gmw_ohi_2009$km2) # 139033.7

gmw_ohi_2010 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2010.csv")))
sum(gmw_ohi_2010$km2) # 137775.4

gmw_ohi_2015 <- read.csv(file.path(here("globalprep/hab_mangrove/v2021/int/habitat_extent_mangrove_2015.csv")))
sum(gmw_ohi_2015$km2) # 136855.9


all_mangrove_extent <- rbind(gmw_ohi_1996, gmw_ohi_2007, gmw_ohi_2008, gmw_ohi_2009, gmw_ohi_2010, gmw_ohi_2015, gmw_ohi_2016) %>%
  filter(rgn_id <= 250) %>%
  filter(year >1996) %>% ## going to exclude 1996 from trend
  filter(km2 != 0) # filter out the regions that have no extent and never have (all of them that have at least 1 zero have never had mangrove extent)

data_region_trend <-  all_mangrove_extent %>%
  left_join(rgns_eez) %>%
  dplyr::select(rgn_id, rgn_name, year, habitat, km2) %>%
  group_by(rgn_id, rgn_name) %>%
  mutate(total_area = sum(km2)) %>% 
  mutate(km2_rel = km2/km2[year==2007]) %>%
  mutate(km2_rel = ifelse(total_area==0, 0, km2_rel)) %>% 
  do(mdl = lm(km2_rel ~ year, data=.)) %>%
  summarize(rgn_name = rgn_name,
            rgn_id = rgn_id,
            habitat = "mangrove",
            trend = coef(mdl)['year']*5) %>%
  mutate(trend = round(trend, 6)) %>%
  ungroup()

mean(data_region_trend$trend)

test <- data_region_trend %>%
  filter(rgn_id != 99)
mean(test$trend) # -0.006990163

  
## save it 
write.csv(data_region_trend, 'globalprep/hab_mangrove/v2021/data/habitat_trend_mangrove.csv', row.names=FALSE)

## look at Benin since it had a large increase
benin <- all_mangrove_extent %>%
  filter(rgn_id == 99)

## compare to old 
old <- read.csv('globalprep/hab_mangrove/v2015/data/habitat_trend_mangrove_updated.csv')

## note compare these when we obtain the new extents...
# 
setdiff(old$rgn_id, data_region_trend$rgn_id) 
setdiff(data_region_trend$rgn_id, old$rgn_id)

old <- old %>%
  left_join(data_region_trend, by='rgn_id')

plot(old$trend.x, old$trend.y, ylab="new trend", xlab="old trend") ## trend is wildly different 


```





