---
title: "OHI `r format(Sys.Date(), '%Y')` - tidal flat extent"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script generates the extent of tidal flat for each OHI region. 


## Updates from previous assessment

2022 - Completely new dataset and layer!

***
## Data Source 

**Reference**: Murray, N.J., Phinn, S.R., DeWitt, M., Ferrari, R., Johnston, R., Lyons, M.B., Clinton, N., Thau, D. & Fuller, R.A. (2019) The global distribution and trajectory of tidal flats. Nature, 565, 222-225.

**Downloaded**: 2022-07-26

**Description**:  
The Murray Global Intertidal Change Dataset contains global maps of tidal flat ecosystems produced via a supervised classification of 707,528 Landsat Archive images. Each pixel was classified into tidal flat, permanent water or other with reference to a globally distributed set of training data.

The classification was implemented along the entire global coastline between 60° North and 60° South from 1 January 1984 to 31 December 2016. The image collection consists consists of a time-series of 11 global maps of tidal flats at 30m pixel resolution for set time-periods (1984−1986; 1987−1989; 1990−1992; 1993−1995; 1996−1998; 1999−2001; 2002−2004; 2005−2007; 2008−2010; 2011−2013; 2014−2016)

This product depicts tidal flat ecosystems around the global coastline.

Pixels classified as tidal flat in the analysis represent several types of tidal flat ecosystems, including unconsolidated fine-grain sediments (tidal mudflats), unconsolidated coarse-grain sediments (tidal sand flats), and consolidated sediments, organic material or rocks (wide tidal rock-platforms), while excluding spectral signatures indicating the presence of vegetation dominated intertidal ecosystems such as mangroves and vegetated marshes. The analysis aimed to identify pixels that are subject to regular tidal inundation, and therefore may also include other intertidal systems where intertidal dynamics are observable.

**Time range**: 1984-2016

**Download link**: https://www.intertidal.app/download/direct-download (use the provided shell script in this file)

**Variables**:

- classification	occurrence: intertertidal area classification for the interval.

  - 0 is not tidal flat
  - 1 is tidal flat

- raster shards use geodetic extent in the file names

***

# Methods

## Overview

The goal is to find tidal flat area in each raster

1. Down sample rasters from 30 m$^2$ ground sample distance (GSD) to ~ 1 km$^2$ GSD

    - Group 30 pixels into 1 to get 900 m$^2$ resolution
      - $\frac{(900 * 900)}{(1,000 * 1,000)}$ = 0.81 km$^2$
      - take the sum of the pixels in the group
  
2. Convert to area

    - multiply sum of grouped pixels by 0.0009
    - If the native raster cells are 30x30 m$^2$ area, then a raster cell with a value of 1 would be equivalent to 0.0009 km$^2$ habitat area
      - $\frac{(30*30)}{(1000*1000)}$ = 0.0009 km$^2$
    
3. Summarize the grid cells within each region of the raster shard

    - Take the sum of the cells in each country for each year
    - Output a `.csv` file for each raster

4. Summarize tidal flat inside each region

    - read in the tidal flat files
    - concatenate files
    - summarize to OHI regions
    - output final tidal flat extent summary

## Setup

``` {r setup, eval = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE)

if (!require(librarian)){install.packages("librarian")}
if (!require(ohicore)){devtools::install_github('ohi-science/ohicore@dev')}

librarian::shelf(
  tidyverse,
  here,
  janitor,
  plotly,
  sf,
  terra,
  exactextractr,
  # fasterize,
  # raster,
  tictoc,
  foreach,
  doParallel
) 

### directory paths and relevant files
current_year <- 2022
version_year <- paste0("v", current_year)
data_year <- paste0("d", current_year)

# source(here::here('workflow', 'R', 'common.R'))

### Mazu
dir_here  <- here::here('globalprep', 'hab_tidal_flat', version_year)
dir_data <- file.path(dir_M, 'git-annex', 'globalprep', '_raw_data', 'global_tidal_wetland_change', data_year)

### Aurora
# dir_here <- file.path('/home/shares/food-systems/Global_datasets/global_tidal_wetland_change')
# dir_data <- file.path('/home/shares/food-systems/Global_datasets/global_tidal_wetland_change/d2022')
```

### Download the data

The following bash script will download the raw data for the tidal wetland change dataset. 

```{bash eval = FALSE}
### Make the new directory and move into it

### MAZU
# mkdir /home/shares/ohi/git-annex/globalprep/_raw_data/global_tidal_wetland_change/d2022/raw/tidal_flats && cd $_

### Aurora
mkdir /home/shares/food-systems/Global_datasets/global_tidal_wetland_change/d2022/raw/tidal_flats && cd $_

### Declare an array of strings
declare -a StringArray=(
  "1984-1986" "1987-1989" "1990-1992" "1993-1995" "1996-1998" 
  "1999-2001" "2002-2004" "2005-2007" "2008-2010" "2011-2013" "2014-2016"
)

### Iterate the string array using for loop
for val in ${StringArray[@]}; do
  echo $val
  wget https://storage.googleapis.com/uq-intertidal/v1_1/global_intertidal/$val.zip
  unzip $val.zip -d ./$val
  rm $val.zip
done
```

### Load OHI region data

```{r eval = FALSE}
regions_shape()

regions <- read_sf(here::here(dir_here, 'regions_2017_update.shp'))

regions_eez_and_land <- regions %>%
  dplyr::filter(rgn_type %in% c("eez", 'land')) %>% 
  sf::st_transform(crs = 4326) 

dummy_df <- regions_eez_and_land %>% 
  as_tibble() %>% 
  dplyr::select(rgn_id, rgn_name) 

rgns <- ohicore::rgn_master %>% 
  dplyr::filter(rgn_typ == "eez") %>% 
  dplyr::select(rgn_id = rgn_id_2013, rgn_name = rgn_nam_2013) %>% 
  dplyr::distinct() %>% 
  dplyr::filter(!rgn_id %in% c(213, 255)) # Antarctica and DISPUTED
```

## Find tidal flat area in each raster

Down sample rasters from 30 m$^2$ ground sample distance (GSD) to ~ 1 km$^2$ GSD

- Group 30 pixels into 1 to get 900 m$^2$ resolution
  - $\frac{(900 * 900)}{(1,000 * 1,000)}$ = 0.81 km$^2$
  - take the sum of the pixels in the group
  
Convert to area

- multiply sum of grouped pixels by 0.0009
- If the native raster cells are 30x30 m$^2$ area, then a raster cell with a value of 1 would be equivalent to 0.0009 km$^2$ habitat area
  - $\frac{(30*30)}{(1000*1000)}$ = 0.0009 km$^2$
    
Summarize the grid cells within each country

- Take the sum of the cells in each country for each year
- Output a `.csv` file for each raster

```{r eval = FALSE}
tictoc::tic()

folders <- here::here(dir_data, "raw", "tidal_flats") %>% 
  list.files()

j <- 1

doParallel::registerDoParallel(11)

foreach::foreach (dir = folders) %dopar% {

# for (dir in folders) { 
  
  files <- here::here(dir_data, "raw", "tidal_flats", dir) %>% 
    list.files()
  
  year = stringr::str_split(dir, pattern = "-", n = 2)[[1]][2]
  
  for (file in files){
    
    cat(j, "/", length(files), "\n")
    
    j <- j + 1
    
    fn <- file %>% 
      stringr::str_replace(
        pattern     = ".tif",
        replacement = paste0("_", year, ".csv"))
    
    file_name <- here::here(dir_data, "int", "tidal_flat", fn)
    
    if(!file.exists(file_name)){
      
      cat(fn, " is being created!\n")
      
      cat("    down sampling...\n")
      
      image <- here::here(dir_data, "raw", "tidal_flats", dir, file) %>% 
        terra::rast() 
       
      down_sampled_image <- terra::aggregate(
        x = image, 
        fact = 30, 
        fun = sum,
        na.rm = T
      )

      cat("    converting to area (km2)...\n\n")
      
      down_sampled_image <- down_sampled_image * 0.0009
      
      cat("    extracting...\n")
      
      extracted_values <- exactextractr::exact_extract(
        x = down_sampled_image, y = regions_eez_and_land, fun = 'sum'
      )
      
      cbind("km2" = extracted_values, dummy_df) %>% 
        dplyr::group_by(rgn_id, rgn_name) %>% 
        dplyr::summarise(km2 = sum(km2)) %>% 
        dplyr::mutate(year = year, habitat = "tidal flat") %>% 
        readr::write_csv(file_name)
      
    } else {cat(paste0(fn, " already exists!\n"))}
  }
}
tictoc::toc()
```

## Summarise tidal flat inside each region

Read in the tidal flat files, concatenate files, summarize to OHI regions, and output final tidal flat extent summary

```{r eval = FALSE}
files <- here::here(dir_data, "int", "tidal_flat") %>% 
  list.files(full.names = T)

output <- dplyr::tibble()
  
for (file in files) {
  
  tmp <- readr::read_csv(file, col_types = cols()) 

  output <- rbind(output, tmp) 
  
}

final <- output %>%
  dplyr::group_by(rgn_id, year) %>%
  dplyr::summarise(km2 = sum(km2, na.rm = T)) %>% 
  readr::write_csv(here::here(dir_here, 'output', paste0("tidal_flat_extent.csv")))
```




