---
title: "GFW_API"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Global Fishing Watch API](https://github.com/GlobalFishingWatch/gfwr)

```{r}
library(gfwr)
library(here)
library(raster)
library(tidyverse)
library(foreach)
library(doParallel)
library(tictoc)
library(readr)
library(dplyr)
library(doSNOW)

# Save API token information to an object every time you need to extract the token and pass it to `gfwr` functions:
key <- gfw_auth()

source('http://ohi-science.org/ohiprep_v2022/workflow/R/common.R')

#options(scipen = 999)

# raster template
#r_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
```

# Quick run through of methods that will be applied to every country and year
## Use a combination of the Vessels & Events APIs to extract fishing events from all trawlers between the dates of interest

```{r}
# extract unique trawler ID's for fishing vessels
# when we eventually execute this for the final script, we might need/want to edit the next function to loop through this function for each country, so subsetting the query for example "flag = 'USA' AND geartype = 'trawlers'"
# trawlers <- get_vessel_info(query = "geartype = 'trawlers'",
#                             search_type = "advanced",
#                             dataset = "fishing_vessel") # omit carrier vessels and support vessels

# combine all unique trawler ID's into a comma separated list that we will use for the next query
# trawler_ids <- paste0(trawlers$id[1:100], collapse = ",")
# # change this to unique(trawler$id)?
# # query all fishing events for just trawlers 
# fishing_events_trawler <- get_event(event_type = "fishing",
#                                     vessel = trawler_ids,
#                                     include_region = TRUE,
#                                     start_date = "2012-01-01",
#                                     end_date = "2013-01-01",
#                                     key = key)

# seems that we NEED to subset the list of trawler$id's in order for the next function get_event() to work. consider opening issue about that if it is not fixed?
# subsetting 1:100 worked but 1:1000 took ages
```

```{r}
# check that trawlers$id values are the vessel ID's and not the fishing event ID's
#id_check <- trawlers %>% 
#  filter(id == "a6be75fec-cab1-83bc-6e10-17687e4814e1") # that is the case!
```

## get fishing hours from diff in time stamps

```{r}
# wrangle data for total fishing hours per vessel
# fishing_events_trawler <- fishing_events_trawler %>%
#   # calculate time for each fishing event
#   mutate(fishing_event_time = end - start) %>% 
#   # change name of column to be able to have unique colnames in next step when we unnest th elist column with the vessel ID
#   rename(fishing_event_id = id) %>%
#   # unlist column that contains the vessel id so we can sum total hours by vessel
#   unnest_wider(vessel) %>% 
#   select("fishing_event_id", lat, lon, vessel_id = id, fishing_event_time) %>% 
#   group_by(vessel_id) %>% 
#   # sum total fishing hours by vessel!
#   summarize(total_vessel_fishing_time = sum(fishing_event_time))

```

## Start over process to repeat these steps for each country individually by year, then create a dataframe of all total sums with column names (rgn_id, year, total_fishing_hours)

```{r}
# load all OHI regions df:
# region_data()
# 
# #rgn = # convert OHI region to GFW 3-letter region code
#   
# #foreach(r = rgns_all$rgn_id)
# 
# trawlers <- get_vessel_info(query = paste0("flag = ", rgn, " AND geartype = 'trawlers'"),
#                             search_type = "advanced",
#                             dataset = "fishing_vessel") # omit carrier vessels and support vessels
# 
# # combine all unique trawler ID's into a comma separated list that we will use for the next query
# trawler_ids <- paste0(trawlers$id[1:100], collapse = ",")
# 
# year <- 2012:2020
# 
# # query all fishing events for just trawlers 
# fishing_events_trawler <- get_event(event_type = "fishing",
#                                     vessel = trawler_ids,
#                                     include_region = TRUE,
#                                     start_date = paste0(year, "-01-01"),
#                                     end_date = paste(year, "-12-31"),
#                                     key = key)
```

### scratch code:

```{r}
# usa_fishing <- get_vessel_info(query = "flag = 'USA'",
#                                search_type = "advanced",
#                                dataset = "fishing_vessel")
# 
# usa_ids <- paste0(usa_fishing$id[100], collapse = ',')
# usa_ids
# 
# usa_fishing$id
# 
# fishing_events_trawler <- get_event(event_type = "fishing",
#                                     vessel = usa_ids,
#                                     include_region = TRUE,
#                                     start_date = "2017-01-01",
#                                     end_date = "2017-02-01",
#                                     key = key)
# 
# fishing_events_trawler$time_diff <- fishing_events_trawler$end - fishing_events_trawler$start
```

## Switch to using the Map Visualization API

```{r}
# load all OHI regions df:
# region_data()

# try one country's eez first using the ISO3 code, with only 1 eez:
# code_eez <- get_region_id(region_name = 'FJI', region_source = 'eez', key = key)
# 
# fishing_hours_16_17 <- gfwr::get_raster(spatial_resolution = 'low',
#                  temporal_resolution = 'yearly',
#                  group_by = 'flag',
#                  date_range = '2016-01-01,2017-12-31',
#                  region = code_eez$id,
#                  region_source = 'eez',
#                  key = key) %>% 
#   rename(year = "Time Range",
#          fishing_rgn = flag)


# this df represents all countries' fishing hours in that EEZ, but rows separate the lat and lon, so group by country to get the total per country by year
# in order to expand out to all eez's, i wonder if i can make a list with c() for the region argument
```

```{r}
# try with cocos islands:
# code_eez <- get_region_id(region_name = 'CCK', region_source = 'eez', key = key)
# 
# # aus_eez_ids <- paste0(code_eez$id, collapse = ",")
# 
# fishing_hours_16_17 <- gfwr::get_raster(spatial_resolution = 'low',
#                  temporal_resolution = 'yearly',
#                  group_by = 'flag',
#                  date_range = '2016-01-01,2017-12-31',
#                  region = code_eez$id[1],
#                  region_source = 'eez',
#                  key = key)

# did not work
```

```{r}
# load all OHI regions df:
region_data()

# iterate through all eez id values in `id` column (inner loop) for each country (outer loop):
foreach(r = rgns_eez$eez_iso3) %do% {
  foreach(e = code_eez$id) %do% {
    fishing_hours_16_17 <- gfwr::get_raster(spatial_resolution = 'low',
                                            temporal_resolution = 'yearly',
                                            group_by = 'flag',
                                            date_range = '2016-01-01,2017-12-31',
                                            region = e, 
                                            region_source = 'eez',
                                            key = key)
    
    write_csv(fishing_hours_16_17, paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", r, "_effort_16_17.csv"))
  }
}

# expand range to 2012-2020 
# add parallelization
```

## Loop development: smaller subset of countries

```{r}
region_data()

# test on subset of countries: choose large countries for which we know there is at least some fishing data 
rgns_eez_subset <- rgns_eez %>%
  filter(eez_iso3 %in% c("USA", "CHN", "THA", "RUS"))

foreach(r = rgns_eez_subset$eez_iso3) %do% {
  # extract all eez codes for countries
  code_eez <- get_region_id(region_name = r, region_source = 'eez', key = key)
  # iterate through all eez codes, e, for countries, r, to extract apparent fishing hours
  print(paste0("Processing apparent fishing hours for ", r, " EEZ ", code_eez$id))
  
  foreach(e = code_eez$id, .combine = rbind) %do% {
    fishing_hours_16_17 <- gfwr::get_raster(spatial_resolution = 'low',
                                            temporal_resolution = 'yearly',
                                            group_by = 'flag',
                                            date_range = '2015-01-01,2018-12-31',
                                            region = e, 
                                            region_source = 'eez',
                                            key = key) %>% 
      # rename columns for clarity
      rename(year = "Time Range",
             fishing_rgn = flag,
             apparent_fishing_hours = "Apparent Fishing hours") %>% 
      mutate(eez_admin_rgn = r)
    
    # add country name to each row so we can keep track of the eez'a admin country group later in the analysis, even after we combine all data
    #fishing_hours_16_17$eez_admin_rgn <- r # might be able to take this code out - we might not do anything with the admin country cause we just care about the fishing rgn
    
    print(paste0("Extracted all apparent fishing hours for ", r, " EEZ ", e))
    
    write_csv(fishing_hours_16_17, paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", r, "_effort_16_17.csv"))
  }
}

# check out the output
usa <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/USA_effort_16_17.csv"))
chn <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/CHN_effort_16_17.csv"))
tha <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/THA_effort_16_17.csv"))
rus <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/RUS_effort_16_17.csv"))
```

## group output CSV's by latitude, longitude, and year

```{r}
# read in each csv and concatenate into one list
fish_effort_df <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api"), pattern = ".csv", full = TRUE) %>%
  lapply(data.table::fread) %>%
  bind_rows()

# group by lat, long, and year
fish_effort_annual <- fish_effort_df %>%
  group_by(cell_ll_lat, cell_ll_lon, year) %>%
   summarize(fishing_hours = sum(fishing_hours)) %>%
    data.frame()

# add region id to each column


# save final dataframe
#write_csv(, ".csv")
```






Notes:
- can we assume that the time difference between stamps is ALL fishing hours? there is also a variable called hours or something in the files downloaded from GFW manually, perhaps look again on website for this difference in variables or contact GFW to inquire
- consider open issue about needing to subset the id list in order to plug it all into get_event()?



