---
title: "OHI 2022 - Soft bottom pressure data prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

This habitat pressure layer is created from apparent fishing effort data from Global Fishing Watch. The apparent fishing effort contains trawling and dredging data within EEZ's because these types of fishing damage the seabed. GFW apparent fishing effort data does not distinguish between bottom and mid-water trawling, so the trawling data is corrected (subset) to just represent the trawling that is likely to be bottom trawling based on catch data from Watson & Tidd (2018). For this catch data, the total tonnes are calculated for each OHI region and then standardized by area (km^2) soft-bottom habitat in each region.

# Updates from previous assessment

New data source: [Global Fishing Watch API](https://github.com/GlobalFishingWatch/gfwr)
Processing this new data requires a completely different data prep. Some methods were based on the `OHI-Science/food_systems` approach [here.](https://github.com/OHI-Science/food_systems/tree/master/fisheries/marine/disturbance). Methods were upscaled appropriately and switched from `raster` to `terra`.

***

# Data Sources 

## Global Fishing Watch Apparent Fishing Effort

**Reference**:
1. Global Fishing Watch. [2022]. www.globalfishingwatch.org
2. [`gfwr` API](https://github.com/GlobalFishingWatch/gfwr)

**Downloaded**: July 21, 2022

**Description**: API to extract apparent fishing effort within global EEZ's, labeled with geartype and date.

**Native data resolution**: 0.01 degree

**Time range**: 2012 - 2020

**Format**:  API

## Catch data:

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://www.sciencedirect.com/science/article/pii/S0308597X18300605?via%3Dihub)

**Downloaded**: July 29, 2022 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0)

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**: 0.5 degree

**Time range**: 2012 - 2017

**Format**:  csv format

***
  
# Methods 
1. Pull apparent fishing effort data for all EEZ regions for 2012-2020 from the GFW API, combine into one dataframe, and filter for trawling and dredging geartypes
2. Correct trawling data by subsetting for only bottom trawling using catch data from Watson & Tidd (2018) and combine with dreding data
3. Layer the fishing effort data with OHI regional polygons and standardize by soft-bottom habitat of each region calculate fishing effort for each OHI region **(insert transformation approach here, like ln(x + 1) transformation if density data are extremely skewed)**
4. Rescale the transformed data y dividing the maximum value across all years. **also consider using a second rescaling using the median values across years**

This script provides pressure scores, health indicators, and trend values for soft-bottom habitats.  

***

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gfwr)
library(terra)
library(raster)
library(tidyverse)
library(foreach)
library(doParallel)
library(tictoc)
library(readr)
library(sf)
library(exactextractr)
library(here)

# Save API token information to an object every time you need to extract the token and pass it to `gfwr` functions
# jcohen's key is saved to jcohen's .Renviron file, which is read by this function
key <- gfw_auth()
source('http://ohi-science.org/ohiprep_v2022/workflow/R/common.R')
options(scipen = 999)
```

## Iterate through all ISO3 regions' EEZ codes to extract all apparent fishing effort from 2012-2020 

```{r}
# load all OHI region ISO codes
region_data()

# convert regional codes into characters first:
regions_all <- sort(unique(rgns_eez$eez_iso3)) # 201
regions_all

# remove China from list because there is so much fishing effort over the entire time period that the API cannot process their first EEZ code, it needs to be processed 
regions <- regions_all[regions_all != "CHN"]

# remove CW since that is not an iso3 code it is an iso2 code that is not compatible with the GFW API
regions <- regions[regions != "CW"]
# add in the correct iso3 code for Curacao
regions <- append(regions, "CUW") # even tho the GFW database does not currently have EEZ codes listed for this iso3 region, perhaps they will in the future when this script is run again

# remove AW since that is not an iso3 code it is an iso2 code that is not compatible with the GFW API
regions <- regions[regions != "AW"]
# add in the correct iso3 code for Aruba
regions <- append(regions, "ABW") # even tho the GFW database does not currently have EEZ codes listed for this iso3 region, perhaps they will in the future when this script is run again

# remove BQ since that is not an iso3 code it is an iso2 code that is not compatible with the GFW API
regions <- regions[regions != "BQ"]
# add in the correct iso3 code for Bonaire
regions <- append(regions, "BQ-BO") # even tho the GFW database does not currently have EEZ codes listed for this iso3 region, perhaps they will in the future when this script is run again

# View list of regions that will be fed into the GFW API loop
sort(regions)
# after the loop extracts data for all other countries, we will manually extarct and save a csv for CHN only 
```

```{r}
tic()

# iterate through all EEZ codes for all regions to extract apparent fishing hours:
for(i in regions) {
  
  # create dataframe that contains the column `id` that is list of all EEZ codes for one region
  eez_code_df <- get_region_id(region_name = i, region_source = 'eez', key = key)
  # convert that column into a numeric list of EEZ codes to feed into the next loop:
  eez_codes <- eez_code_df$id
  
  print(paste0("Processing apparent fishing hours for ", i, " EEZ code ", eez_codes))
  
  for(j in eez_codes) { 
    fishing_hours <- gfwr::get_raster(spatial_resolution = 'high', # high = 0.01 degree resolution which we think is close to 30 m resolution
                                      temporal_resolution = 'yearly',
                                      group_by = 'flagAndGearType', # maybe change to just geartype
                                      date_range = '2012-01-01,2020-12-31', 
                                      region = j, 
                                      region_source = 'eez',
                                      key = key) %>%
      # rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = i) %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    fishing_hours$year <- as.numeric(fishing_hours$year)
    fishing_hours$apparent_fishing_hours <- as.numeric(fishing_hours$apparent_fishing_hours)
    fishing_hours$y <- as.numeric(fishing_hours$y)
    fishing_hours$x <- as.numeric(fishing_hours$x)
    fishing_hours$eez_admin_rgn <- as.character(fishing_hours$eez_admin_rgn)
    fishing_hours$geartype <- as.character(fishing_hours$geartype)
    
    print(paste0("Extracted all apparent fishing hours for ", i, " EEZ code ", j))
    
    write_csv(fishing_hours, paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", i, "_", j, "_effort.csv")) 
  }
}

toc()
```

### Loop through apparent fishing effort for China separately due to the large quantity of data over all years 

```{r}
# Running CHN's first listed EEZ for all years, 2012-2020, breaks the API, so we need to pull data in 3 time chunks

chn_code_eez <- get_region_id(region_name = "CHN", region_source = 'eez', key = key)

# 2012-2015 for first code
chn_eez1_2012_2015 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2012-01-01,2015-12-31',
                             region = chn_code_eez$id[1],
                             region_source = 'eez',
                             key = key) %>% 
  # rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = "CHN") %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    chn_eez1_2012_2015$year <- as.numeric(chn_eez1_2012_2015$year)
    chn_eez1_2012_2015$apparent_fishing_hours <- as.numeric(chn_eez1_2012_2015$apparent_fishing_hours)
    chn_eez1_2012_2015$y <- as.numeric(chn_eez1_2012_2015$y)
    chn_eez1_2012_2015$x <- as.numeric(chn_eez1_2012_2015$x)
    chn_eez1_2012_2015$eez_admin_rgn <- as.character(chn_eez1_2012_2015$eez_admin_rgn)
    chn_eez1_2012_2015$geartype <- as.character(chn_eez1_2012_2015$geartype)

#  2016-2017 for first code
chn_eez1_2016_2017 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2016-01-01,2017-12-31',
                             region = chn_code_eez$id[1],
                             region_source = 'eez',
                             key = key) %>% 
  # rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = "CHN") %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    chn_eez1_2016_2017$year <- as.numeric(chn_eez1_2016_2017$year)
    chn_eez1_2016_2017$apparent_fishing_hours <- as.numeric(chn_eez1_2016_2017$apparent_fishing_hours)
    chn_eez1_2016_2017$y <- as.numeric(chn_eez1_2016_2017$y)
    chn_eez1_2016_2017$x <- as.numeric(chn_eez1_2016_2017$x)
    chn_eez1_2016_2017$eez_admin_rgn <- as.character(chn_eez1_2016_2017$eez_admin_rgn)
    chn_eez1_2016_2017$geartype <- as.character(chn_eez1_2016_2017$geartype)

# 2018-2020 for first code
chn_eez1_2018_2020 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2018-01-01,2020-12-31',
                             region = chn_code_eez$id[1],
                             region_source = 'eez',
                             key = key) %>% 
  # rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = "CHN") %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    chn_eez1_2018_2020$year <- as.numeric(chn_eez1_2018_2020$year)
    chn_eez1_2018_2020$apparent_fishing_hours <- as.numeric(chn_eez1_2018_2020$apparent_fishing_hours)
    chn_eez1_2018_2020$y <- as.numeric(chn_eez1_2018_2020$y)
    chn_eez1_2018_2020$x <- as.numeric(chn_eez1_2018_2020$x)
    chn_eez1_2018_2020$eez_admin_rgn <- as.character(chn_eez1_2018_2020$eez_admin_rgn)
    chn_eez1_2018_2020$geartype <- as.character(chn_eez1_2018_2020$geartype)
    
# combine data for all 3 time periods into 1 dataframe for this one EEZ code for CHN
chn_eez1_all_years <- bind_rows(chn_eez1_2012_2015, chn_eez1_2016_2017, chn_eez1_2018_2020)

write_csv(chn_eez1_all_years, file = paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/CHN_", chn_code_eez$id[1], "_effort.csv"))



# the second eez code for China is fine to process over all years
chn_eez2 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2012-01-01,2020-12-31',
                             region = chn_code_eez$id[2],
                             region_source = 'eez',
                             key = key) %>% 
# rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = "CHN") %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    chn_eez2$year <- as.numeric(chn_eez2$year)
    chn_eez2$apparent_fishing_hours <- as.numeric(chn_eez2$apparent_fishing_hours)
    chn_eez2$y <- as.numeric(chn_eez2$y)
    chn_eez2$x <- as.numeric(chn_eez2$x)
    chn_eez2$eez_admin_rgn <- as.character(chn_eez2$eez_admin_rgn)
    chn_eez2$geartype <- as.character(chn_eez2$geartype)

write_csv(chn_eez2, path = paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/CHN_", chn_code_eez$id[2], "_effort.csv"))
```

## Dataset cleaning: remove files with 0 rows in order to combine all files row-wise into one object in next steps

Document which files (EEZ regions) had no fishing detected by GFW data in 2012-2020 in order to notice trends over years.

```{r}
fish_effort_files <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/"), pattern = ".csv", full = TRUE)

# first check the length to see how many files we start with
num_files_start <- length(fish_effort_files)
print(paste0("Starting with ", num_files_start, " files."))

# delete files that do not have any rows (because no fishing has been recorded in that EEZ)
for (i in seq_along(fish_effort_files)) {
  # read each file and check number of rows
  filename <- fish_effort_files[i]
  print(paste0("Counting rows (fishing effort observations) for ", substr(filename, -29, -21)))
  # save the numbe of rows for that file to an object
  rows <- nrow(data.table::fread(filename))
  print(paste0(rows, " rows in this file."))
  # if there are 0 rows, delete the file
  if (rows == "0") {
      unlink(filename) 
    }
}

# redefine variable after deleting some files
fish_effort_files <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/"), pattern = ".csv", full = TRUE)

# check length again to see how many files we end with
num_files_end <- length(fish_effort_files)
print(paste0("Ending with ", num_files_end, " files because ", (num_files_start - num_files_end), " files were deleted for containing no data."))
# 2022 assessment: 27 files were removed because had 0 rows, indicating no EEZ codes in the GFW database for that region, or no fishing activity at all within those countries' EEZs in 2012-2020
```

## Data Wrangling: Combine regional csv's into one dataframe and filter by geartype

We do this because our goal is to summarize fishing effort spatially to produce annual rasters of fishing effort for all regions. After these fishing effort rasters have been processed, we will overlay them with the OHI regional polygons to attribute fishing activity to certain OHI regions and produce CSV's.

```{r}
# concatenate csv's into one list for trawling data, separate from dredging so we can correct the trawling data and then sum the corrected trawling data with the dredging data later on
fish_effort_trawl <- fish_effort_files %>%
  lapply(data.table::fread) %>% # read in each file that represents data for 1 EEZ for 1 country
  bind_rows() %>% # combine all files into one dataframe
  filter(geartype == "trawlers") %>% # trawl fishing damages the seafloor
  dplyr::select(-geartype) # now that all observations are trawlers, we can drop this variable

# delete:
fish_effort_trawl_nga <- fish_effort_trawl %>% filter(eez_admin_rgn == "NGA")
trawl_nga_2020 <- fish_effort_trawl_nga %>% filter(year == 2020)
sum_nga_2020_trawl <- sum(trawl_nga_2020$apparent_fishing_hours, na.rm = TRUE)

# concatenate csv's into one list for dredging data, we will later sum this with the corrected trawling data
fish_effort_dredge <- fish_effort_files %>%
  lapply(data.table::fread) %>% # read in each file that represents data for 1 EEZ for 1 country
  bind_rows() %>% # combine all files into one dataframe
  filter(geartype == "dredge_fishing") %>% # dredge fishing damages the seafloor
  dplyr::select(-geartype) # now that all observations are from dredging, we can drop this variable

# delete:
fish_effort_dredge_nga <- fish_effort_dredge %>% filter(eez_admin_rgn == "NGA")
dredge_nga_2020 <- fish_effort_dredge_nga %>% filter(year == 2020)
sum_nga_2020_dredge <- sum(dredge_nga_2020$apparent_fishing_hours, na.rm = TRUE)


# take a look at the data
head(fish_effort_dredge) # still contains admin rgn
```

## Group by coordinate and year for apparent fishing effort for both geartypes of interest

We do not group by the administrative region for each EEZ because the coordinates of the fishing activity will allow us to attribute fishing activity to OHI regions. 

```{r}
# trawling
fish_effort_trawl_annual <- fish_effort_trawl %>% 
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 
# convert year column from integer to factor in preparation for next steps:
fish_effort_trawl_annual$year <- as.factor(fish_effort_trawl_annual$year)

# dredging
fish_effort_dredge_annual <- fish_effort_dredge %>% # all dredge data will be used, we will not be subsetting it like we will for the trawling data
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 
# convert year column from integer to factor in preparation for next steps:
fish_effort_dredge_annual$year <- as.factor(fish_effort_dredge_annual$year)

# take a look
head(fish_effort_dredge_annual) # no longer contains EEZ admin rgn
```

### Save the trawling and dredging data as annual raster files before we can adjust them for mid-water versus bottom trawling.

```{r}
tic()
# this loop sometimes randomly errors with: Error in x$.self$finalize() : attempt to apply non-function
# but just run again and it should work! Takes ~27 minutes
years = as.factor(2012:2020) # make it factor in order to match the class of year column in the annual trawling & dredging dataframes

for(i in years){
  # trawling data:
  trawl_annual_raster <- fish_effort_trawl_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast(type = "xyz", crs = "EPSG:4326", digits = 6, extent = NULL)
  
  # save annual raster file for trawling:
  terra::writeRaster(trawl_annual_raster, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i,"/fish_effort_trawl_", i, ".tif"), overwrite = TRUE)
}

for(i in years){
  # dredging data:
  dredge_annual_raster <- fish_effort_dredge_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast(type = "xyz", crs = "EPSG:4326", digits = 6, extent = NULL)

  # save annual raster file for dredging:
  terra::writeRaster(dredge_annual_raster, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_dredge_", i , ".tif"), overwrite = TRUE) # put in folder that will be the destination of the trawling data as well, after it is corrected, so the dredging and trawling by year can be summed easily in a stack
}

toc()
```

### Correct Trawling Data: Subset trawling fishing effort by catch data for the corresponding year to distinguish between mid-water trawling and bottom trawling

We use fisheries catch data from the paper by Watson, R. A. and Tidd, A. 2018 to distinguish between the two types of trawling, because the GFW data groups all trawling together. We only maintain bottom trawling because that is the type that destroys soft bottom habitat. We multiply a raster that represents a proxy for the proportion of mid-water trawling to bottom trawling by the Global Fishing Watch raster for that respective year. We then sum that data with all the dredging data in the next steps. We only have catch data (the trawling correction data) for 2012-2017, so we use 2017 data for 2018-2020 GFW data.

### Iterate through 2012-2017 data years to subset trawling data for just bottom trawling

```{r}
tic() # expect chunk to run for ~18 minutes

# use this subset because we only have trawling proportion data for these years from Watson and Tidd 2018
years_subset = as.factor(2012:2017)

# for years with trawling depth proportion data:
for (j in years_subset){
  
  print(paste0("Processing trawling fishing effort for ", j))
  # read in the GFW apparent fishing effort raster for trawling
  # we need this read in to resample the Watson catch data to the same resolution
  fish_effort_trawl <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/fish_effort_trawl_", j, ".tif"))
  
  # read in the annual file that represents the proportion of bottom trawling to midwater trawling, based on catch data, from Watson et al.
  trawl_depth_proportion <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_", j, ".tif"))
  
  print(paste0("Interpolating trawling depth proportion data for ", j))
  # interpolate many of the missing values in the trawling depth proportion raster by taking the local average of cells around it (using a window of 9 cells) that represent the proportion of bottom trawling to midwater trawling that occurred at that coordinate. This provides a more comprehensive raster which we will later multiply by the GFW trawling raster, cell by cell
  trawl_depth_proportion_interpolated <- terra::focal(x = trawl_depth_proportion, 
                                                      w = 5, # sets a 5 cell window around NA pixel
                                                      fun = "mean", 
                                                      na.policy = "only", # only interpolate NA values
                                                      na.rm = T,
                                                      overwrite = TRUE)
  
  # document how many NA values were filled by `terra::focal()`
  print(paste0("Started with ", terra::global(trawl_depth_proportion, fun = "isNA")[1,1], " NA values, and ended with ", terra::global(trawl_depth_proportion_interpolated, fun = "isNA")[1,1], " so filled ", terra::global(trawl_depth_proportion, fun = "isNA")[1,1] - terra::global(trawl_depth_proportion_interpolated, fun = "isNA")[1,1], " NA values using `terra::focal()`."))
  
  # clear some memory
  rm(trawl_depth_proportion)
  gc()
  
  # fill in remaining NA values with the global mean of the non-NA cell values
  global_trawl_proportion_avg <- terra::global(trawl_depth_proportion_interpolated, fun = "mean", na.rm = TRUE)
  trawl_depth_proportion_interpolated[is.na(trawl_depth_proportion_interpolated)] <- global_trawl_proportion_avg[1,1]
  
  # document how many NA values remain after applying the local average to remaining NA cells (should be 0)
  print(paste0("There are now ", terra::global(trawl_depth_proportion_interpolated, fun = "isNA"), " NA values present after filling the remaining NA's with the global average."))
  
  print(paste0("Resampling trawling depth proportion data for ", j))
  
  # resample the interpolated trawling proportion data to match the resolution of the GFW fishing effort data
  trawl_depth_proportion_resampled <- terra::resample(x = trawl_depth_proportion_interpolated, 
                                                      y = fish_effort_trawl, # use the GFW data as the sample geometry
                                                      method = "near") # nearest neighbor
  
  # document how many NA values were produced by resampling (NA values are produced for some years, but not all)
  print(paste0("Now there are ", terra::global(trawl_depth_proportion_resampled, fun = "isNA")[1,1], " NA values remaining, need to fill NA's one final time."))
  
  # get rid of the rest of any NA values produced by resample:
  trawl_depth_proportion_resampled[is.na(trawl_depth_proportion_resampled)] <- global_trawl_proportion_avg[1,1]
  
  # document how many NA values remain (should be 0)
  print(paste0("Now there are ", terra::global(trawl_depth_proportion_resampled, fun = "isNA")[1,1], " NA values remaining."))
  
  # save the adjusted bottom trawling proportion raster in the folder of the corresponding year of GFW data, & save the most recent year of proportion data (2017) for 2017-2020 GFW data
  if (j == 2017) {
    # save to folders for 2017-2020
    print(paste0("Saving ", j, " file to ", j, " folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2018 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2018/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2019 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2019/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2020 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2020/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
  } else {
    # for trawling proportion data for years 2012-2016, save file to just that year's folder
    print(paste0("Saving ", j, " file to ", j, " folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
  }  
}

# clear some memory
rm(trawl_depth_proportion_resampled)
gc()

toc() 
```

#### Apply the adjusted fishing catch rasters (which have been resampled and interpolated) to subset the GFW trawling data for just bottom trawling

First, we extend the trawling depth proportion raster to the larger extent of the trawling fishing effort raster and vice versa to ensure all dimensions of both extents are maximized and match one another. We then multiply the proportion raster by the trawling raster to reflect the amount of trawling that is attributed to bottom trawling. We exclude mid-water trawling because it does not damage the seafloor.

```{r}
tic() # chunks takes 13 minutes

years_all <- as.factor(2012:2020)

for (i in years_all){
  
  print(paste0("Checking if extents match for ", i, " trawling and fishing catch rasters."))
  # first read in both rasters within the annual folders and save the list of 2 as an object
  rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i), pattern = ".tif", full = TRUE)
  fish_effort_raster <- terra::rast(rasters[1])
  trawl_depth_proportion_raster <- terra::rast(rasters[2])
  print(ext(fish_effort_raster)) # peek at the raster extents
  print(ext(trawl_depth_proportion_raster))
  extent_comparison <- terra::compareGeom(fish_effort_raster, trawl_depth_proportion_raster, stopOnError = FALSE) # set stopOnError = FALSE in order to produce "FALSE" from function if extents don't match
  print(paste0("Raster extent comparison results: ", extent_comparison))

  if (extent_comparison == FALSE){ 
      print("Extending trawling proportion raster to match the extent of the trawling fishing effort raster and vice versa.")
      trawl_depth_proportion_extended <- terra::extend(trawl_depth_proportion_raster, fish_effort_raster)
      fish_effort_extended <- terra::extend(fish_effort_raster, trawl_depth_proportion_extended) # extend both rasters to the extent of the other in all dimensions
    
      # check the extents now
      print(ext(fish_effort_extended))
      print(ext(trawl_depth_proportion_extended))
      extent_comparison <- terra::compareGeom(fish_effort_extended, trawl_depth_proportion_extended, stopOnError = FALSE)
      print(paste0("Raster extent comparison results: ", extent_comparison)) # should be TRUE
      
      # check how many NA values are present in the extended trawl proportion raster
      print(paste0("After extending, there are ", terra::global(trawl_depth_proportion_extended, fun = "isNA")[1,1], " NA values in the trawl proportion raster."))
      # before multiplying the rasters, we need to fill in the NA values in the trawl_depth_proportion_extended raster with 1, so that we do not lose any of the trawling data that lacks corresponding catch data
      # get rid of the rest of any NA values produced by resample:
      trawl_depth_proportion_extended[is.na(trawl_depth_proportion_extended)] <- 1
      # document how many NA values remain (should be 0)
      print(paste0("Now there are ", terra::global(trawl_depth_proportion_extended, fun = "isNA")[1,1], " NA values remaining."))
    
      # clear some memory
      rm(rasters)
      gc()
      
      # re-write the extended rasters to a nested folder called "extended" within the trawling annual folders so we don't need to overwrite the original raster files
      terra::writeRaster(trawl_depth_proportion_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/trawl_depth_proportion_extended_", i, ".tif"), overwrite = TRUE)
      print("Output raster for trawl proportion data with extent to match fishing effort data.")
      terra::writeRaster(fish_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/fish_effort_extended_", i, ".tif"), overwrite = TRUE)
      print("Output raster for fishing effort raster with extent to match trawl proportion data.")
      
      # since the rasters have been rewritten as new filenames, need to redefine the object that represents them
      rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/"), pattern = ".tif", full = TRUE)
      print("Redefined rasters list object.")
  }

  # stack the resampled interpolated raster of bottom trawling with all trawling data for that year from GFW, matching the data by year for 2012-2017 and then using 2017 for the rest of the years of GFW trawling data for which we do not have more recent trawling proportion data
  trawl_stack <- terra::rast(rasters)
  
  print(paste0("Correcting trawling effort data for ", i))
  
  # multiply the rasters in order to only maintain the proportion of trawling fishing effort that is attributed to bottom trawling rather than midwater trawling
  terra::lapp(trawl_stack, fun = function(x,y){return(x*y)}, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_trawl_corrected_", i, ".tif"), overwrite = TRUE)
  
  print(paste0("Saved corrected trawling data for ", i, " to ", paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_trawl_corrected_", i, ".tif")))

  # clear some memory
  rm(trawl_stack)
  gc()
}
  
toc()
```

## Combine the corrected trawling data with the dredging data by summing the fishing effort by coordinate

Similarly to the last step, we extend the annual dredging rasters to the extent of the corrected trawling raster and vice versa. Since the dredging rasters generally have smaller extents than the trawling fishing effort rasters, many `NA` values are produced by extending it. The presence of `NA` values requires us to sum using the argument `na.rm = TRUE` in order to avoid nullifying a large proportion of the trawling fishing effort data points in the calculation. Since both the dredging data and the trawling data were pulled from the Global Fishing Watch database, these files already have the same resolution and were assigned the same CRS earlier in the data processing.

```{r}
tic() # chunk took 13 minutes
years_all <-  as.factor(2012:2020)

for (i in years_all){
  
  print(paste0("Checking if extents match for ", i, " trawling and dredging rasters."))
  # first read in both rasters within the annual folders and save the list of 2 as an object
  rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i), pattern = ".tif", full = TRUE)
  dredge_effort_raster <- terra::rast(rasters[1])
  trawl_effort_raster <- terra::rast(rasters[2])
  print(ext(dredge_effort_raster)) # smaller extent so extend this raster first
  print(ext(trawl_effort_raster))
  extent_comparison <- terra::compareGeom(dredge_effort_raster, trawl_effort_raster, stopOnError = FALSE)
  print(paste0("Raster extent comparison results: ", extent_comparison))

  if (extent_comparison == FALSE){ 
      print("Extending trawling proportion raster to match the extent of the dredging effort raster and vice versa.")
      dredge_effort_extended <- terra::extend(dredge_effort_raster, trawl_effort_raster)
      trawl_effort_extended <- terra::extend(trawl_effort_raster, dredge_effort_extended) # extend both rasters to the extent of the other in all dimensions
    
      # check the extents now
      print(ext(dredge_effort_extended))
      print(ext(trawl_effort_extended))
      extent_comparison <- terra::compareGeom(dredge_effort_extended, trawl_effort_extended, stopOnError = FALSE)
      print(paste0("Raster extent comparison results: ", extent_comparison)) # should be TRUE
    
      # clear some memory
      rm(rasters)
      gc()
      
      # re-write the extended rasters to a nested folder called "extended" to avoid overwriting the original files 
      terra::writeRaster(trawl_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/fish_effort_dredge_", i, "_extended.tif"), overwrite = TRUE)
      print("Re-wrote trawling effort file.")
      terra::writeRaster(dredge_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/fish_effort_trawl_", i, "_extended.tif"), overwrite = TRUE)
      print("Re-wrote dredging effort file.")
      
      # since the rasters have been rewritten, need to redefine the object that represents them
      rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/"), pattern = ".tif", full = TRUE)
      print("Redefined rasters list object.")
  }

  # stack the corrected trawling data with the dredging data from GFW, matching the data by year
  effort_stack <- terra::rast(rasters)
  
  print(paste0("Correcting trawling effort data for ", i))

  terra::app(effort_stack, fun = "sum", na.rm = TRUE, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"), overwrite = TRUE)

  # clear some memory
  rm(effort_stack)
  gc()
  
  print(paste0("Saved corrected fishing effort data for ", i, " to ", paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif")))
}
  
toc() 
```

### Plot a raster to visualize fishing effort

```{r}
# load the EEZ spatial data to layer on plot with fishing effort points for spatial reference
regions_shape()
# subset to just EEZ's
regions_eez_land <- regions %>%
  filter(rgn_type %in% c("eez", "land")) %>% 
  st_transform(crs = "EPSG:4326")

View(head(nigeria))

nigeria <- regions_eez_land %>% 
  filter(rgn_id == 196)

# read in the annual file for combined trawling and dredging fishing effort
fishing_effort_2020 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2020.tif"))
plot(fishing_effort_2020, col = "red") 

# Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
plot(nigeria$geometry, col = "grey96", axes = FALSE, main = "Nigeria Trawling & Dredging Fishing Effort 2020", legend = FALSE)
plot(fishing_effort_2020, axes = FALSE, add = TRUE)
```

### Calculate fishing effort for each OHI region

```{r}
# tic() 
# 
# # Read in OHI regions polygons - several polygons (rows) for each region because one polygon exists for each region & land type combo (eez, land, etc.)
# ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
# # Make the region coordinate reference system the same as the trawling & dredging raster
# ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326") 
# 
# for (i in years_all) {
#   print(paste("Processing ", i, " fishing effort."))
#   
#   # read in the annual fishing raster for all regions that represents corrected trawling plus dredging data 
#   fishing_raster <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"))
#   
#   # extract the fishing effort that falls within the polygons of OHI regions
#   # calculate a weighted sum of the fishing effort across all area of the polygon that the coordinates fall within
#   extracted <- exactextractr::exact_extract(fishing_raster, ohi_regions_wgs, fun = "weighted_sum", weights = "area")
# 
#   # assign names to be the cell terrain type (eez, land, etc.), and the OHI region ID to each value in the vector so we can then convert to a tidy dataframe
#   names(extracted) <- paste(ohi_regions_wgs$type_w_ant, ohi_regions_wgs$rgn_ant_id, sep = " ")
#   
#   # create a df of all these cell values and their labels for terrain type and region id as their own columns
#   fishing_regional <- plyr::ldply(extracted, rbind) %>%
#   # rename the dataframe's default variable names for clarity
#   rename(fishing_effort = "1",
#          "rgn_type rgn_id" = ".id") %>% 
#   # split the conjoined columns for region id and land type into individual columns
#   tidyr::separate(col = "rgn_type rgn_id", 
#                   into = c("rgn_type", "rgn_id"),
#                   sep = " ")
#   # in the next steps we summarize by just rgn_id, so this is the last object that allows for a data check (for example, if fishing effort hours are associated with a land type polygon)
#   
#   # The following code keeps the raster information when the raster lands partially on the land polygons because we do not group by terrain type, just rgn_id
#   fishing_regional_summary <- fishing_regional %>%
#   dplyr::group_by(rgn_id) %>% # drop the land type variable
#   dplyr::summarize(effort = sum(fishing_effort, na.rm = TRUE)) %>% 
#   dplyr::ungroup() %>%
#   dplyr::arrange(as.numeric(rgn_id)) %>% 
#   dplyr::mutate(year = i)
# 
#   # convert this raster info into a csv that represents the trawling and dredging fishing effort for each OHI region for year i 
#   write.csv(fishing_regional_summary, file = paste0(here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"), row.names = FALSE)
#   print(paste0("Saved ", i, " fishing effort csv to ", here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"))
# }
# 
# # data check for points that fall on land, the check can only be executed for 1 isolated year outside the loop:
# # check rows that have land with fishing effort hours
# #land_fishing <- fishing_regional %>% 
# #  filter(rgn_type %in% c("land", "eez-inland", "land-ccamlr", "land-disputed", "land-noeez") & fishing_effort > 0)
# 
# # plot the fishing effort points on the terrain type polygons so we can try to visualize why fishing on land is occurring
# #plot(ohi_regions_wgs$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2012: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
# #plot(fishing_raster, axes = FALSE, col = "red", add = TRUE)
# # turns out it is just from slight error in the coordinates, the points that are documented on land are not far inland, just at the coasts
# 
# toc()
```

```{r}
# troublehsoot code for attributing fishing hours to coordinates within regional polygons
# Read in OHI regions polygons - several polygons (rows) for each region because one polygon exists for each region & land type combo (eez, land, etc.)
# ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
# # Make the region coordinate reference system the same as the trawling & dredging raster
# ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326") 
# # determine class of polygons
# class(ohi_regions_wgs) # sf dataframe
# class(ohi_regions_wgs$geometry) # sfc_MULTIPOLYGON
# View(head(ohi_regions_wgs))
# # convert polygons sf object to a spatVector
# polygon_vector <- terra::vect(ohi_regions_wgs$geometry)
# class(polygon_vector) # "SpatVector"
# 
# # read in the annual fishing raster for all regions that represents corrected trawling plus dredging data 
# fishing_spatrast_2014 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2014.tif"))
# 
# # extract spatrast from polygons
# #extracted <- terra::extract(fishing_spatrast_2014, polygon_vector)
# extracted_with_weights <- terra::extract(fishing_spatrast_2014, polygon_vector, weights = TRUE, progress = "text") # using eaxct = TRUE crashed R, even tho it is similar to weights = TRUE, did not do sum within this function because not sure if it will do na.rm = TRUE
# class(extracted_with_weights)
# View(head(extracted_with_weights))
# # create copy of this in case I mess something up
# extracted_with_weigths_copy <- data.frame(extracted_with_weights)
# 
# # make new copy of copy since I messed something up
# extarcted_with_weights_copy_2 <- data.frame(extracted_with_weigths_copy)
# 
# unique(extarcted_with_weights_copy_2$ID)
# 
# # make new column that multiplies the weights by the cell value
# extracted_with_weights <- extracted_with_weigths_copy %>% 
#   mutate(weighted_value = weight*sum) %>% 
#   group_by(ID) %>% 
#   summarize(polygon_sum = sum(weighted_value, na.rm = TRUE)) # tried keeping na's just to see if rows turns out to what we need
# 
# # why are there 519 values instead of 526?
# unique(extracted_with_weights$ID)
# 
# # check which polygon indices are missing
# # first create a vector that ranges from 1-526 and then compare them
# all_values <- 1:526
# 
# setdiff(all_values, extracted_with_weights$ID) # 281 283 297 299 328 330 332
# 
# # find what regions are associated with those index values
# ohi_regions_wgs[281,]
# ohi_regions_wgs[283,]
# ohi_regions_wgs[297,]
# ohi_regions_wgs[299,]
# ohi_regions_wgs[330,]
# ohi_regions_wgs[332,]
# 
# # check if this is ALL the antarctica polygons? if so, then i think we can just use the col without the ant id instead to match names
# antarctica_polys <- ohi_regions_wgs %>% 
#   filter(rgn_name == "Antarctica")
# # it is not all the antarctica regions, so we can just remove those rows from the vector we use to assign names
# remove_indices <- c(281, 283, 297, 299, 328, 330, 332)
# ohi_regions_wgs_adjusted <- ohi_regions_wgs[-remove_indices,]
# 
# type <- ohi_regions_wgs_adjusted$type_w_ant
# rgn_id <- ohi_regions_wgs_adjusted$rgn_ant_id
# 
# # add the type and region id as new columns
# extracted_w_info <- extracted_with_weights %>% 
#   cbind(type, rgn_id) 
  

# type_vector <- ohi_regions_wgs_adjusted$type_w_ant
# type_w_ant_adjusted <- type_vector[-remove_indices]
# 
# rgn_id_vector <- ohi_regions_wgs_adjusted$rgn_ant_id
# rgn_id_adjusted <- rgn_id_vector[-remove_indices]

#unique(extracted$)
# assign names to be the cell terrain type (eez, land, etc.), and the OHI region ID to each value in the vector so we can then convert to a tidy dataframe
#names(extracted_with_weights) <- paste(ohi_regions_wgs_adjusted$type_w_ant, ohi_regions_wgs_adjusted$rgn_ant_id, sep = " ")

# create a df of all these cell values and their labels for terrain type and region id as their own columns
#fishing_regional <- plyr::ldply(extracted_with_weights, rbind) %>%
  # rename the dataframe's default variable names for clarity
  #rename(fishing_effort = "1",
  #       "rgn_type rgn_id" = ".id") %>% 
  # split the conjoined columns for region id and land type into individual columns
 # tidyr::separate(col = "rgn_type rgn_id", 
 #                 into = c("rgn_type", "rgn_id"),
 #                 sep = " ")

# The following code keeps the raster information when the raster lands partially on the land polygons because we do not group by terrain type, just rgn_id
# fishing_regional_summary <- extracted_w_info %>%
#  # dplyr::mutate(effort = value*weight) %>% # might need to change value here 
#   dplyr::group_by(rgn_id) %>% # drop the land type variable
#   dplyr::summarize(effort = sum(polygon_sum, na.rm = TRUE)) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(as.numeric(rgn_id)) %>%
#   dplyr::mutate(year = 2014)

# filter for ABW polygons only
# aruba_poly <- ohi_regions_wgs %>% filter(rgn_id == 250)
# uk_poly <- ohi_regions_wgs %>% filter(rgn_id == 180)
# # read in the annual file for combined trawling and dredging fishing effort for all regions
# fishing_effort_2014 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2014.tif"))
# # plot all regions' fishing on ABW polygons:
# plot(uk_poly$geometry, col = "grey96", axes = FALSE, main = "Aruba Trawling & Dredging Fishing Effort 2014")
# plot(fishing_effort_2014, axes = FALSE, add = TRUE, col = "red")


#spatraster_2014 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2014.tif"))

# view 2014 raster as df
#fishing_df_2014 <- as.data.frame(fishing_raster_2014, xy = TRUE, na.rm = FALSE)
#View(head(fishing_df_2014_no_na))

#fishing_df_2014_no_na <- fishing_df_2014 %>% filter(sum != "NaN")

# extract the fishing effort that falls within the polygons of OHI regions
# calculate a weighted sum of the fishing effort across all area of the polygon that the coordinates fall within
#extracted_raster <- raster::extract(raster_2014, ohi_regions_wgs, weights = TRUE, normalizeWeights = FALSE, progress = 'text')
#ohi_regions_vec <- terra::vect(ohi_regions_wgs)
#ohi_regions_vec <- terra::vect(ohi_regions_wgs)
# reorder dataframe by the ordered id column rgn_ant_id
#regions_ordered <- sort(ohi_regions_wgs$rgn_ant_id)

#ohi_regions_vec <- terra::vect(ohi_regions_wgs$geometry)

#extracted_spatraster <- terra::extract(spatraster_2014, ohi_regions_vec, "sum", na.rm = TRUE, progress = "text")
#unique(extarcted_spatraster$sum)
# what is the difference between ohi_regions_wgs$rgn_ant_id and ohi_regions_wgs$rgn_id?
# with_ant <- ohi_regions_wgs$rgn_ant_id
# no_ant <- ohi_regions_wgs$rgn_id
# library(useful)
# compare.list(with_ant, no_ant)
# with_ant[274:283]
# no_ant[274:283]
```

```{r}
# convert for loop to use terra::extract() instead of exact_extractr()
# tic() 
# 
# # Read in OHI regions polygons - several polygons (rows) for each region because one polygon exists for each region & land type combo (eez, land, etc.)
# ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
# # Make the region coordinate reference system the same as the trawling & dredging raster
# ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326") 
# 
# polygon_vector <- terra::vect(ohi_regions_wgs$geometry)
# 
# remove_indices <- c(281, 283, 297, 299, 328, 330, 332)
# ohi_regions_wgs_adjusted <- ohi_regions_wgs[-remove_indices,]
# 
# type <- ohi_regions_wgs_adjusted$type_w_ant
# rgn_id <- ohi_regions_wgs_adjusted$rgn_ant_id
# 
# years_all <-  as.factor(2012:2020)
# 
# for (i in years_all) {
#   print(paste("Processing ", i, " fishing effort."))
#   
#   # read in the annual fishing raster for all regions that represents corrected trawling plus dredging data 
#   fishing_raster <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"))
#   
#   # extract the fishing effort that falls within the polygons of OHI regions
#   # calculate a weighted sum of the fishing effort across all area of the polygon that the coordinates fall within
#   print("Extracting fishing effort from polygons.")
#   extracted_with_weights <- terra::extract(fishing_raster, polygon_vector, weights = TRUE)
# 
#   # make new column that multiplies the weights by the cell value
#   extracted_with_weights <- extracted_with_weights %>% 
#     mutate(weighted_value = weight*sum) %>% 
#     group_by(ID) %>% 
#     summarize(polygon_sum = sum(weighted_value, na.rm = TRUE)) # tried keeping na's just to see if rows turns out to what we need
#   
#   # add the type and region id as new columns
#   extracted_w_info <- extracted_with_weights %>% 
#     cbind(type, rgn_id) 
#   
#   # The following code keeps the raster information when the raster lands partially on the land polygons because we do not group by terrain type, just rgn_id
#   # The following code keeps the raster information when the raster lands partially on the land polygons because we do not group by terrain type, just rgn_id
# fishing_regional_summary <- extracted_w_info %>%
#  # dplyr::mutate(effort = value*weight) %>% # might need to change value here 
#   dplyr::group_by(rgn_id) %>% # drop the land type variable
#   dplyr::summarize(effort = sum(polygon_sum, na.rm = TRUE)) %>%
#   dplyr::ungroup() %>%
#   dplyr::arrange(as.numeric(rgn_id)) %>%
#   dplyr::mutate(year = i)
# 
#   # convert this raster info into a csv that represents the trawling and dredging fishing effort for each OHI region for year i 
#   write.csv(fishing_regional_summary, file = paste0(here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"), row.names = FALSE)
#   print(paste0("Saved ", i, " fishing effort csv to ", here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"))
# }

# data check for points that fall on land, the check can only be executed for 1 isolated year outside the loop:
# check rows that have land with fishing effort hours
#land_fishing <- fishing_regional %>% 
#  filter(rgn_type %in% c("land", "eez-inland", "land-ccamlr", "land-disputed", "land-noeez") & fishing_effort > 0)

# plot the fishing effort points on the terrain type polygons so we can try to visualize why fishing on land is occurring
#plot(ohi_regions_wgs$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2012: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
#plot(fishing_raster, axes = FALSE, col = "red", add = TRUE)
# turns out it is just from slight error in the coordinates, the points that are documented on land are not far inland, just at the coasts

#toc()
```

```{r}
# use exact_extractr but dont take the sum yet, just get out a list of df's, one for each polyogn, that contains cell values and the coverage fraction, then multiply those values, and sum to produce sum for each polygon, then make that into a dataframe with polygon number as a col and other col is that sum value, then row bind all those df's for the final annual df for all polygons
# try 2014 first

# Read in OHI regions polygons - several polygons (rows) for each region because one polygon exists for each region & land type combo (eez, land, etc.)
# ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
# # Make the region coordinate reference system the same as the trawling & dredging raster
# ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326") 
# 
# # read in the annual fishing raster for all regions that represents corrected trawling plus dredging data 
# fishing_raster_2014 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2014.tif"))
# 
# # check if there are NA values in the final fishing rasters in general
# # fishing_effort_2014_df <- as.data.frame(fishing_raster_2014, xy = TRUE, na.rm = FALSE)
# # View(head(fishing_effort_2014_df))
# # sort(unique(ohi_regions_wgs$rgn_name))
# # unique(ohi_regions_wgs$rgn_type)
# 
# # filter out antarctica and disputed and all land types besides eez and land 
# ohi_regions_filtered <- ohi_regions_wgs %>% 
#   filter(rgn_type %in% c("eez", "land"),
#          !rgn_name %in% c("Antarctica", "DISPUTED")) 
# 
# # extract the fishing effort that falls within the polygons of OHI regions
# # calculate a weighted sum of the fishing effort across all area of the polygon that the coordinates fall within
# extracted <- exactextractr::exact_extract(fishing_raster_2014, ohi_regions_filtered)
# 
# # sum effort by polygon and append polygon dataframes into 1
# all_polygon_effort <- data.frame()
# 
# for (i in seq_along(extracted)) {
# 
#   print(paste("Processing polygon ", i))
# 
#   df <- extracted[[i]]
# 
#   weighted_effort_polygon <- df %>%
#     mutate(effort = value*coverage_fraction,
#            polygon_id = i) %>%
#     select(polygon_id, effort) %>%
#     group_by(polygon_id) %>%
#     summarize(effort_sum = sum(effort, na.rm = TRUE))
# 
#   all_polygon_effort <- rbind(all_polygon_effort, weighted_effort_polygon)
# 
#   print(paste("Appended effort for polygon ", i, " to dataframe all_polygon_effort."))
# }
# 
# # bind regional id's to the polygon id's
# rgn_id <- ohi_regions_filtered$rgn_id
# regional_polygon_effort <- cbind(all_polygon_effort, rgn_id)
# 
# regional_polygon_effort <- regional_polygon_effort %>% 
#   select(-polygon_id) %>% 
#   mutate(year = 2014)

```

```{r}
tic() 

# Read in OHI regions polygons - several polygons (rows) for each region because one polygon exists for each region & land type combo (eez, land, etc.)
ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
# Make the region coordinate reference system the same as the trawling & dredging raster
ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326") 

# filter out antarctica and disputed and all land types besides eez and land 
ohi_regions_filtered <- ohi_regions_wgs %>% 
  filter(rgn_type %in% c("eez", "land"),
         !rgn_name %in% c("Antarctica", "DISPUTED"))

years_all <-  as.factor(2012:2020)

for (i in years_all) {
  
  print(paste("Processing ", i, " fishing effort."))
  
  # read in the annual fishing raster for all regions that represents corrected trawling plus dredging data 
  fishing_raster <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"))
  
  # extract the fishing effort that falls within the polygons of OHI regions
  extracted <- exactextractr::exact_extract(fishing_raster, ohi_regions_filtered)

  # sum effort by polygon and append polygon dataframes into 1
  all_polygon_effort <- data.frame()

  for (j in seq_along(extracted)) {
    
    df <- extracted[[j]]
    
    weighted_effort_polygon <- df %>%
      # weight the fishing effort within each cell by the coverage of the cell within the polygon 
      mutate(effort = value*coverage_fraction,
             polygon_id = j) %>%
      select(polygon_id, effort) %>%
      group_by(polygon_id) %>%
      summarize(effort_sum = sum(effort, na.rm = TRUE))
    # bind the polygon-specific sum to the master dataframe of polygon sums
    all_polygon_effort <- rbind(all_polygon_effort, weighted_effort_polygon)
    
  }

  # bind regional id's to the polygon id's
  rgn_id <- ohi_regions_filtered$rgn_id
  regional_polygon_effort <- cbind(all_polygon_effort, rgn_id)

  regional_polygon_effort <- regional_polygon_effort %>%
    group_by(rgn_id) %>%
    # combine each land and eez polygon for the same region, because some of the fishing points landed just on the border of these polygons and are associated with land when they should be eez
    summarize(effort_sum = sum(effort_sum, na.rm = TRUE)) %>% 
    mutate(year = i)

  # convert this raster info into a csv that represents the trawling and dredging fishing effort for each OHI region for year i 
  write.csv(regional_polygon_effort, file = paste0(here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"), row.names = FALSE)
  print(paste0("Saved ", i, " fishing effort csv to ", here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv/fishing_effort_", i, ".csv"))
}

toc() # 21 min
```

## Standarize catches per soft-bottom area

```{r, eval=FALSE}
area <-  read.csv(paste0(here(), "/globalprep/hab_prs_hd_subtidal_soft_bottom/v2016/output/habitat_extent_softbottom.csv"))

# Read in csv for each year and combine them in one data frame
files <- list.files(here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_annual_csv"), pattern = ".csv", full = TRUE)

annual_list <- lapply(files, function(x){read.csv(file = x)})
# combine all annual dataframes into one with a year variable present
fishing_all <- bind_rows(annual_list) 

data_density <- fishing_all %>%
  left_join(area, by = "rgn_id") %>%
  filter(rgn_id <= 250) %>%
  mutate(density = effort_sum/km2)

# annual_data_density <- data_density %>% 
#   group_by(year) %>% 
#   summarize(total_effort_km2 = sum(density, na.rm = TRUE))
# 
# annual_density_plot <- ggplot(annual_data_density, aes(year, total_effort_km2)) +
#   geom_point() +
#   labs(title = "Demersal Fishing Effort per Softbottom Habitat Area\nAll regions 2012-2020",
#              x = "Year",
#              y = "Fishing Effort\n(hours per km2 soft bottom habitat)") +
#   theme(plot.title = element_text(hjust=0.5))
# annual_density_plot
# looks very skewed, likely due to outliers

# get summary stats for the data density before winsorizing and summing by year
summary(data_density$density)
# Min.   1st Qu.    Median      Mean   3rd Qu.      Max.      NA's 
#    0.0000    0.0000    0.0005    4.4239    0.2861 2503.8848       135 

# Winsorize the data density 
data_density <- data_density %>%
  # if the data is above the 99th quantile, replace it with the 99th quantile
  mutate(quantDensity = if_else(
    density > quantile(x = density, probs = c(0.99), na.rm = TRUE),
    quantile(x = density, probs = c(0.99), na.rm = TRUE),
    density))

annual_data_density_adjusted <- data_density %>% 
  group_by(year) %>% 
  summarize(quantDensity = sum(quantDensity, na.rm = TRUE))

ggplot(annual_data_density_adjusted, aes(year, quantDensity)) + 
  geom_line() +
  labs(title = "Demersal Fishing Effort per km2 of Softbottom Habitat 2012-2020\nAdjusted for 99th Quantile",
       x = "Year",
       y = "Fishing Effort\n(Hours per km2 of soft bottom habitat)")

write.csv(annual_data_density_adjusted, file = here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/reference_point_quant_99.csv"), row.names = FALSE)

# compare to the same years from the 2020 data (99th quantile)

assessment_2020 <- read.csv(here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2020/int/reference_point_quant_99.csv")) %>% 
  filter(year %in% c(2012:2020))

assessment_2022 <- annual_data_density_adjusted %>% 
  filter(year %in% c(2012:2017))

ggplot() +
  geom_line(data = assessment_2020, aes(x = year, y = quantDensity)) +
  geom_text(aes(2013, 115, label = "2022 Assessment\nGFW data", vjust = 1)) +
  geom_line(data = assessment_2022, aes(x = year, y = quantDensity)) +
  geom_text(aes(2014, 80, label = "2020 Assessment\nWatson catch data", vjust = 1)) +
  labs(x = "Year",
       y = "Fishing Effort\n(catch or fishing hours standardized by km2 of soft bottom habitat)",
       title = "Comparison od 2020 and 2022 Assessment:\nDemersal Fishing Effort per km2 of Softbottom Habitat 2012-2017\nAdjusted for 99th Quantile")

# rescale the density data
## Reference point = 99th quantile across all years
ref_point <- read.csv(here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/reference_point_quant_99.csv"))
ref_point_median <- median(ref_point$quantDensity) # 2022 assessment: 154.01
                                                   
# ref_point (max) 2018 = 7052.172 (year 2015)
# ref_point (max) 2019 = 45.76575 (year 2015)
# ref_point (max) 2020 = 4842.57385 (year 2015)
# ref_point (95th quantile) 2020 = 19.62075 (year 2015)
# ref_point (99th quantile max) 2022 = 225.32314 (year 2020)

## Checking for distribution of data. In the past density data happens to be very skewed
# 2020 assessment scaled the data desnsity (not using any quantile reference point data, the pure data density, that is not summarized by year)
hist(data_density$density) # Yes, very skewed data

## Rescale the density:
data_rescale <- data_density %>%
  mutate(density_rescaled_median = density/ref_point_median) %>%
  mutate(density_rescaled_median_capped = ifelse(density_rescaled_median > 1,
                                                   1,
                                                   density_rescaled_median))
# contains all 220 regions
hist(data_rescale$density_rescaled_median_capped) # heavy skew still

# check which regions have NA values
data_rescale_na <- data_rescale %>% 
  filter(is.na(density_rescaled_median_capped)) 
length(unique(data_rescale_na$rgn_id)) # 15 regions have NA values for fishing effort, this is just because those regions do not have soft bottom habitat in the spatial file

# replace those na values with 0, since there is no potential for those regions to destroy soft bottom habitat within their EEZ's because it is not there to beign with
data_rescale <- data_rescale %>% 
  mutate(density_rescaled_median_capped = ifelse(is.na(density_rescaled_median_capped), 0, density_rescaled_median_capped))

# check it worked:
data_rescale_na <- data_rescale %>% 
  filter(is.na(density_rescaled_median_capped)) 
length(unique(data_rescale_na$rgn_id))
```

## Calculate and save trend and health for softbottom habitat and subtidal habitat

Save Trend and Health of Habitat Types

```{r}
condition_pressure <- data_rescale %>%
  mutate(habitat = "soft_bottom") %>%
  mutate(pressure = density_rescaled_median_capped) %>% 
  mutate(health = 1 - density_rescaled_median_capped) %>% 
  dplyr::select(rgn_id, year, health, pressure, habitat)

# check that all regions are present
length(unique(condition_pressure$rgn_id)) # 220

hist(condition_pressure$pressure)
hist(condition_pressure$health)

## Get habitat trends
stop_year <- max(condition_pressure$year)

trend <- data.frame()

for (status_year in (stop_year-4):stop_year){ 
  trend_years <- status_year:(status_year - 4)
  first_trend_year <- min(trend_years)
  
  trend_new <- condition_pressure %>%
    filter(year %in% trend_years) %>%
    group_by(rgn_id) %>%
    do(mdl = lm(health ~ year, data=.),
       adjust_trend = .$health[.$year == first_trend_year]) %>%
    summarize(rgn_id = rgn_id,
              trend = round(coef(mdl)['year']/adjust_trend * 5, 4)) %>%
    ungroup() %>%
    mutate(trend = ifelse(trend > 1, 1, trend)) %>%
    mutate(trend = ifelse(trend < (-1), (-1), trend))
  
  trend_new <- trend_new %>%
    dplyr::mutate(habitat = "soft_bottom") %>%
    dplyr::mutate(year = status_year) %>%
    dplyr::select(rgn_id, year, habitat, trend)
  
  trend <- rbind(trend, trend_new)
}
  
write.csv(trend, file = here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/output/habitat_trend_softbottom.csv"), row.names = FALSE)

# calculate soft bottom habitat health
health <- condition_pressure %>%
    dplyr::select(rgn_id, year, habitat, health)
  
write.csv(health, file = here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/output/habitat_health_softbottom.csv"), row.names = FALSE)
  
pressure <- condition_pressure %>%
  dplyr::select(rgn_id, year, pressure_score = pressure) 
  
write.csv(pressure, file = here("globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/output/hd_sb_subtidal.csv"))
```




- add eval = false to each chunk, delete old commented out code, fix yaml, knit
- trend
- gf docs
- name final docs: habitat_health_softbottom.csv & habitat_trend_softbottom.csv
















































## Exploration: Check that regions that are infamous for trawling are recorded as trawling as much as we would expect:

```{r}
# usa <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "USA") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))
# 
# chn <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "CHN") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))
# 
# nzl <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "NZL") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))

# see which countries have the most trawling
# rgn_trawl <- fish_effort_all %>% 
#   group_by(eez_admin_rgn) %>% 
#   summarise(total_fishing_hours = sum(apparent_fishing_hours))
# 
# rgn_trawl_max <- rgn_trawl %>% 
#   slice_max(total_fishing_hours, n = 20, with_ties = FALSE) %>%
#   arrange(desc(total_fishing_hours)) 
# 
# # visualize distribution of all regions trawling
# trawling_all <- ggplot(data = rgn_trawl_max, aes(x = eez_admin_rgn, y = total_fishing_hours)) +
#   geom_point() +
#   geom_point(size = 3) +
#   labs(title = "Top 20 Trawling Regions, 2012-2020",
#        subtitle = "GFW Data",
#        x = "Region Code",
#        y = "Total Trawling Fishing Effort Hours") +
#   theme_minimal()
# 
# trawling_all
```

## Data Wrangling: all gear types

```{r}
# fish_effort_allgear <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions"), pattern = ".csv", full = TRUE) %>%
#   lapply(data.table::fread) %>%
#   bind_rows()
# 
# head(fish_effort_allgear)
# 
# unique(fish_effort_allgear$eez_admin_rgn)
# 
# # see which countries have the most fishing in general - all geartypes
# rgn_fe <- fish_effort_allgear %>% 
#   group_by(eez_admin_rgn) %>% 
#   summarise(total_fishing_hours = sum(apparent_fishing_hours))
# 
# rgn_fe_max <- rgn_fe %>% 
#   slice_max(total_fishing_hours, n = 20, with_ties = FALSE) %>%
#   arrange(desc(total_fishing_hours)) 
# 
# # visualize distribution of all regions trawling
# trawling_allgear <- ggplot(data = rgn_fe_max, aes(x = eez_admin_rgn, y = total_fishing_hours)) +
#   geom_point() +
#   geom_point(size = 3) +
#   labs(title = "Top 20 Fishing Regions (all geartypes), 2012-2020",
#        subtitle = "GFW Data",
#        x = "Region Code",
#        y = "Total Fishing Effort Hours") +
#   theme_minimal()
# 
# trawling_allgear
```


## Group apparent trawling fishing effort by latitude, longitude, and year for raster analysis

```{r}
# group all countries and years by lat, long, and year (this df does not have the EEZ admin country anymore, since we don't need that at the moment)
fish_effort_annual <- fish_effort_trawl %>% # change to the relevant trawl data after filtering for that with catch data later
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years, so we will do that in next step
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 

head(fish_effort_annual)

# convert year column from integer to factor in preparation for next steps:
fish_effort_annual$year <- as.factor(fish_effort_annual$year)

# save dataframe before rasterizing:
#write_csv(, ".csv")
```

## Check one annual raster of trawler fishing effort: 2015 as a test before doing all years in loop

```{r}
# try with just 2015 first before loop:
# rasterize the data:

fish_effort_2015 <- fish_effort_trawl %>%
  dplyr::filter(year == "2015") %>%
  dplyr::select(-year)

# set spatial coordinates of a dataframe
# sp::coordinates(fish_effort_2015) <- ~x+y # errors if use "x" + "y"
# # assign EPSG 4326 as the CRS
# raster::rasterFromXYZ(fish_effort_2015, crs = "+init=epsg:4326", digits = 6) 
# proj4string(fish_effort_2015) = CRS("+init=epsg:4326") # code from Gage's footprint project
# #raster::rasterFromXYZ(fish_effort_2015, crs = "EPSG:4326", digits = 6)
# crs(fish_effort_2015)
# plot(fish_effort_2015, col = "red") # plot in red to visualize points easier

# check if using diff raster function makes output diff?
#raster::writeRaster(fish_effort_2015, filename = paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/test.tif'), overwrite = TRUE)

# convert the dataframe to a SpatRaster
fish_effort_2015_rast <- terra::rast(fish_effort_2015)

plot(fish_effort_2015_rast, col = "red") # looks like points but needs to be a df to intersect with EEZ polygons

# convert to points vector (a geometry column?)
#fish_effort_2015_points <- terra::as.points(fish_effort_2015_rast)

#plot(fish_effort_2015_points, col = "red") # looks like blobs where the fishing points were! I suppose the difference between coordinates and points is that points just appear larger when plotted

# load the EEZ spatial data to layer on plot with fishing effort points for reference
regions_shape()

# check out df subset to not overwhelm Mazu by opening regions df in its entirety
View(head(regions))

# filter the regions dataframe for just EEZ polygons
regions_eez <- regions %>%
  filter(rgn_type == "eez") %>% 
  st_transform(crs = 4326)

# try plotting just geometry
plot(regions_eez$geometry, col = "red")

# try to plot points on polygons
#plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "fishing effort 2015", legend = FALSE)		
#plot(fish_effort_2015_points, axes = FALSE, col = "red", add = TRUE) # points are large! not great for viz

# Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2015: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
plot(fish_effort_2015_rast, axes = FALSE, col = "red", add = TRUE)
```

### Visualize trawling & dredging apparent fishing effort data on map of EEZ's as a time series and save the rasters as annual files.

PLot each year and save the raster as a .tif.

- would like to add checks to this, progress bar, more cores for parallelization

```{r}
tic()

#colors <- c("yellow", "red", "blue", "green", "orange", "cyan4") # 6 colors because loop does i +1, so yellow will never be plotted

years = as.factor(2012:2020) # make it factor in order to match the class of year column in the fish_effort_annual dataframe

for(i in years){
  annual_raster <- fish_effort_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast()

    plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawling and Dredging Fishing Effort in ", i, " for AUS, USA, THA, BIH, GBR, ITA", legend = FALSE))
    plot(annual_raster, axes = FALSE, col = "red", add = TRUE, legend = FALSE)
    
    # save annual raster file that encompasses trawling fishing and dredging effort for all countries
    #raster::writeRaster(annual_raster, filename = paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_', i ,'.tif'), overwrite = TRUE)
}

toc()
```

### Plot all years on same map - after all, none of the coordinates were repeated exactly on different years

```{r}
# for(i in years){
#  annual_raster <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_', i ,'.tif'))
# 
#     plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawing and Dredging Fishing Effort in 2012-2020 for AUS, USA, THA, BIH, GBR, ITA", legend = FALSE))
#     plot(annual_raster, axes = FALSE, col = "red", add = TRUE, legend = FALSE)
# }
# 
# toc()
```


```{r}
# fish_effort_2015 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2015.tif'))
# #fish_effort_2015_df <- raster::as.data.frame(fish_effort_2015, xy = TRUE, na.rm = FALSE)
# #tail(fish_effort_2015_df)
# fish_effort_2016 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2016.tif'))
# fish_effort_2017 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2017.tif'))
# fish_effort_2018 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2018.tif'))
```

```{r}
# must use dataframes rather than .tif files for ggplot
# ggplot() +
#   geom_raster(data = eez_df, aes(x = x, y = y, fill = 'eez')) +
#   # add eez csv for plotting:
#   geom_raster(data = fish_effort_2015_df, aes(x = x, y = y)) +
#   scale_fill_viridis_c() +
#   theme_void() +
#   theme(legend.position = "bottom") +
#   coord_equal()
```

## Plot faceted rasters for apparent fishing hours with overlaid map of EEZ's

```{r}
# eez_boundaries <- file.path()
# 
# ggplot() +
#   geom_raster(data = fish_effort_all, aes(x = lon, y = lat, fill = 'apparent_fishing_hours')) +
#   geom_sf(data = eez_boundaries, fill = NA) +
#   scale_fill_viridis_c() +
#   theme_void() +
#   theme(legend.position = "bottom") +
#   coord_equal()
```

## Create dataframe of summarized spatialized fishing effort while maintaining the administrative country for each EEZ

This step enables us to calculate scores for each region on the fishing that occurs in their EEZ?

```{r}
# recall dataframe from earlier with region variable still present: fish_effort_all
# fish_effort_regional <- fish_effort_all %>% 
#   group_by(lat, lon, year, eez_admin_rgn) %>% # only grouped by lat, lon, and year according to output, bc no 2 countries fished in the exact same coordinate in these years 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE))
# 
# year = 2015:2018
# 
# foreach(r = rgns_eez_subset$eez_iso3) %do% {
#   foreach(yr = year) %do% {
#     fish_effort <- fish_effort_all %>%
#       dplyr::filter(year == yr) %>% 
#       dplyr::group_by(lat, lon)
#   }
# }
```

```{r}
# already created files of fishing effort separated by country, start by reading those in, group by year, sum hours
# year = 2015:2018
# foreach(r = rgns_eez_subset$eez_iso3) %do% {
#   regional_fishing_effort <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", r, "_effort_15_18.csv")) %>% 
#     group_by(year) %>% 
#     summarize()
  #}
```


Old Notes:
- consider open issue about needing to subset the id list in order to plug it all into get_event()?



















