---
title: "GFW_API"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Global Fishing Watch API](https://github.com/GlobalFishingWatch/gfwr)

```{r}
library(gfwr)
library(here)
library(raster)
library(tidyverse)
library(foreach)
library(doParallel)
library(tictoc)
library(readr)
library(dplyr)
library(doSNOW)

# Save API token information to an object every time you need to extract the token and pass it to `gfwr` functions:
key <- gfw_auth()

source('http://ohi-science.org/ohiprep_v2022/workflow/R/common.R')

#options(scipen = 999)

# raster template
#r_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
```

# Quick run through of methods that will be applied to every country and year
## Use a combination of the Vessels & Events APIs to extract fishing events from all trawlers between the dates of interest

```{r}
# extract unique trawler ID's for fishing vessels
# when we eventually execute this for the final script, we might need/want to edit the next function to loop through this function for each country, so subsetting the query for example "flag = 'USA' AND geartype = 'trawlers'"
trawlers <- get_vessel_info(query = "geartype = 'trawlers'",
                            search_type = "advanced",
                            dataset = "fishing_vessel") # omit carrier vessels and support vessels

# combine all unique trawler ID's into a comma separated list that we will use for the next query
trawler_ids <- paste0(trawlers$id[1:100], collapse = ",")
# change this to unique(trawler$id)?
# query all fishing events for just trawlers 
fishing_events_trawler <- get_event(event_type = "fishing",
                                    vessel = trawler_ids,
                                    include_region = TRUE,
                                    start_date = "2012-01-01",
                                    end_date = "2013-01-01",
                                    key = key)

# seems that we NEED to subset the list of trawler$id's in order for the next function get_event() to work. consider opening issue about that if it is not fixed?
# subsetting 1:100 worked but 1:1000 took ages

# check that trawlers$id values are the vessel ID's and not the fishing event ID's
#id_check <- trawlers %>% 
#  filter(id == "a6be75fec-cab1-83bc-6e10-17687e4814e1") # that is the case!
```

## get fishing hours from diff in time stamps

```{r}
# wrangle data for total fishing hours per vessel
fishing_events_trawler <- fishing_events_trawler %>%
  # calculate time for each fishing event
  mutate(fishing_event_time = end - start) %>% 
  # change name of column to be able to have unique colnames in next step when we unnest th elist column with the vessel ID
  rename(fishing_event_id = id) %>%
  # unlist column that contains the vessel id so we can sum total hours by vessel
  unnest_wider(vessel) %>% 
  select("fishing_event_id", lat, lon, vessel_id = id, fishing_event_time) %>% 
  group_by(vessel_id) %>% 
  # sum total fishing hours by vessel!
  summarize(total_vessel_fishing_time = sum(fishing_event_time))

```

## Start over process to repeat these steps for each country individually by year, then create a dataframe of all total sums with column names (rgn_id, year, total_fishing_hours)

```{r}
# load all OHI regions df:
region_data()

#rgn = # convert OHI region to GFW 3-letter region code
  
foreach(r = rgns_all$rgn_id)

trawlers <- get_vessel_info(query = paste0("flag = ", rgn, " AND geartype = 'trawlers'"),
                            search_type = "advanced",
                            dataset = "fishing_vessel") # omit carrier vessels and support vessels

# combine all unique trawler ID's into a comma separated list that we will use for the next query
trawler_ids <- paste0(trawlers$id[1:100], collapse = ",")

year <- 2012:2020

# query all fishing events for just trawlers 
fishing_events_trawler <- get_event(event_type = "fishing",
                                    vessel = trawler_ids,
                                    include_region = TRUE,
                                    start_date = paste0(year, "-01-01"),
                                    end_date = paste(year, "-12-31"),
                                    key = key)
```

```{r}
usa_fishing <- get_vessel_info(query = "flag = 'USA'",
                               search_type = "advanced",
                               dataset = "fishing_vessel")

usa_ids <- paste0(usa_fishing$id[100], collapse = ',')
usa_ids

usa_fishing$id

fishing_events_trawler <- get_event(event_type = "fishing",
                                    vessel = usa_ids,
                                    include_region = TRUE,
                                    start_date = "2017-01-01",
                                    end_date = "2017-02-01",
                                    key = key)

fishing_events_trawler$time_diff <- fishing_events_trawler$end - fishing_events_trawler$start
```

## Switch to using the Map Visualization API

```{r}
# try one country's eez first
# retrieve the eez code for usa
code_eez <- get_region_id(region_name = 'CIV', region_source = 'eez', key = key)

fishing_hours_2017 <- gfwr::get_raster(spatial_resolution = 'low',
                 temporal_resolution = 'yearly',
                 group_by = 'flag',
                 date_range = '2017-01-01,2017-12-31',
                 region = code_eez$id, # this runs bc CIV has only 1 EEZ, if you are applying this function to a country with many, need to index which EEZ from the df code_eez or type in manually
                 region_source = 'eez',
                 key = key)

# this df represents all countries' fishing hours in that EEZ, but rows separate the lat and lon, so group by country to get the total per country by year
# in order to expand out to all eez's, i wonder if i can make a list with c() for the region argument 

# try with all USA:
code_eez <- get_region_id(region_name = 'USA', region_source = 'eez', key = key)

# convert the vector of eez ID's to a comma separated character string?
usa_eez_ids <- paste0(code_eez$id, collapse = ',')
usa_eez_ids

usa_eez_fishing_hours_2017 <- gfwr::get_raster(spatial_resolution = 'low',
                 temporal_resolution = 'yearly',
                 group_by = 'flag',
                 date_range = '2017-01-01,2017-12-31',
                 region = usa_eez_ids,
                 region_source = 'eez',
                 key = key)

# did not work, maybe need to iterate through all eez id values in column for each country, and can expand range to 2012-2020 in one query

foreach(r = code_eez$id, )


```

```{r}

```







Notes:
- can we assume that the time difference between stamps is ALL fishing hours? there is also a variable called hours or something in the files downloaded from GFW manually, perhaps look again on website for this difference in variables or contact GFW to inquire
- consider open issue about needing to subset the id list in order to plug it all into get_event()?



