---
title: "OHI 2022 - Soft bottom pressure data prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

This habitat pressure layer is created from apparent fishing effort data from Global Fishing Watch. The apparent fishing effort contains trawling and dredging data within EEZ's because these types of fishing damage the seabed. GFW apparent fishing effort data does not distinguish between bottom and mid-water trawling, so the trawling data is corrected (subset) to just represent the trawling that is likely to be bottom trawling based on catch data from Watson & Tidd (2018). For this catch data, the total tonnes are calculated for each OHI region and then standardized by area (km^2) soft-bottom habitat in each region.

# Updates from previous assessment

New data source: [Global Fishing Watch API](https://github.com/GlobalFishingWatch/gfwr)
Processing this new data requires a completely different data prep. Some methods were based on the `OHI-Science/food_systems` approach [here.](https://github.com/OHI-Science/food_systems/tree/master/fisheries/marine/disturbance). Methods were upscaled appropriately and switched from `raster` to `terra`.

***

# Data Sources 

## Global Fishing Watch Apparent Fishing Effort

**Reference**:
1. Global Fishing Watch. [2022]. www.globalfishingwatch.org
2. [`gfwr` API](https://github.com/GlobalFishingWatch/gfwr)

**Downloaded**: July 21, 2022

**Description**: API to extract apparent fishing effort within global EEZ's, labeled with geartype and date.

**Native data resolution**: 0.01 degree

**Time range**: 2012 - 2020

**Format**:  API

## Catch data:

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://www.sciencedirect.com/science/article/pii/S0308597X18300605?via%3Dihub)

**Downloaded**: July 29, 2022 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0)

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**: 0.5 degree

**Time range**: 2012 - 2017

**Format**:  csv format

***
  
# Methods 
1. Pull apparent fishing effort data for all EEZ regions for 2012-2020 from the GFW API, combine into one dataframe, and filter for trawling and dredging geartypes
2. Correct trawling data by subsetting for only bottom trawling using catch data from Watson & Tidd (2018) and combine with dreding data
3. Layer the fishing effort data with OHI regional polygons and standardize by soft-bottom habitat of each region calculate fishing effort for each OHI region **(insert transformation approach here, like ln(x + 1) transformation if density data are extremely skewed)**
4. Rescale the transformed data y dividing the maximum value across all years. **also consider using a second rescaling using the median values across years**

This script provides pressure scores, health indicators, and trend values for soft-bottom habitats.  

***

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gfwr)
library(terra)
library(raster)
library(tidyverse)
library(foreach)
library(doParallel)
library(tictoc)
library(readr)
library(sf)
library(exactextractr)

# Save API token information to an object every time you need to extract the token and pass it to `gfwr` functions
# jcohen's key is saved to jcohen's .Renviron file, which is read by this function
key <- gfw_auth()
source('http://ohi-science.org/ohiprep_v2022/workflow/R/common.R')
options(scipen = 999)
```

## Iterate through all ISO3 regions' EEZ codes to extract all apparent fishing effort from 2012-2020 

```{r}
tic()

# load all rgn ISO codes
region_data()

# test on subset of countries: choose large countries for which we know there is at least some fishing data
#rgns_subset <- rgns_eez %>%
#  filter(eez_iso3 %in% c("AUS", "USA", "THA", "BIH", "GBR", "ITA"))
# removed CHN from list because cannot process one of the two EEZ: 8486 (first eez in list of 2 from API)

# convert regional codes into characters first:
regions <- rgns_eez %>% 
  # remove China from list because there is so much fishinf effort over the entire time period that the API cannot handle it & crashes
  filter(eez_iso3 != 'CHN') %>% 
  unique(eez_iso3)
# after the loop extracts data for all other countries, we will manually extarct and save a csv for CHN only 

cl <- 3
registerDoParallel(cl)

# iterate through all EEZ codes for all regions to extract apparent fishing hours:
for(i in regions) {
  
  # create dataframe that contains the column `id` that is list of all EEZ codes for one region
  eez_code_df <- get_region_id(region_name = i, region_source = 'eez', key = key)
  # convert that column into a numeric list of EEZ codes to feed into the next loop:
  eez_codes <- eez_code_df$id
  
  print(paste0("Processing apparent fishing hours for ", i, " EEZ code ", eez_codes))
  
  for(j in eez_codes) { 
    fishing_hours <- gfwr::get_raster(spatial_resolution = 'high', # high = 0.01 degree resolution which we think is close to 30 m resolution
                                                temporal_resolution = 'yearly',
                                                group_by = 'flagAndGearType', # maybe change to just geartype
                                                date_range = '2012-01-01,2020-12-31', 
                                                region = j, 
                                                region_source = 'eez',
                                                key = key) %>%
      # rename columns for clarity:
      rename(year = "Time Range",
             apparent_fishing_hours = "Apparent Fishing hours",
             y = Lat,
             x = Lon,
             geartype = Geartype) %>%
      # keep track of the administrative country for each EEZ, even after we combine all data into one dataframe: 
      mutate(eez_admin_rgn = i) %>% 
      select(year, apparent_fishing_hours, y, x, eez_admin_rgn, geartype) # note: we do not care about which region did the actual fishing, just about which country controls that EEZ, which is why we do not maintain the fishing region in the data moving forward
    
    # specify column types before saving the csv so we can correctly concatenate the rows later
    fishing_hours$year <- as.numeric(fishing_hours$year)
    fishing_hours$apparent_fishing_hours <- as.numeric(fishing_hours$apparent_fishing_hours)
    fishing_hours$y <- as.numeric(fishing_hours$y)
    fishing_hours$x <- as.numeric(fishing_hours$x)
    fishing_hours$eez_admin_rgn <- as.character(fishing_hours$eez_admin_rgn)
    fishing_hours$geartype <- as.character(fishing_hours$geartype)
    
    print(paste0("Extracted all apparent fishing hours for ", i, " EEZ code ", j))
    
    write_csv(fishing_hours, paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", i, "_", j, "_effort.csv")) 
  }
}

stopCluster(cl)

toc()
```

### Loop through China's apparent fishing effort separately due to the large quantity of data over all years 

```{r}
chn_code_eez <- get_region_id(region_name = "CHN", region_source = 'eez', key = key)

# check how many eez codes are present, should be 2:

chn_eez1 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2012-01-02,2020-12-30',
                             region = chn_code_eez$id[1],
                             region_source = 'eez',
                             key = key)

write_csv(chn_eez1, path = paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/CHN_", insert_code1, "_effort.csv"))

chn_eez2 <- gfwr::get_raster(spatial_resolution = 'high',
                             temporal_resolution = 'yearly',
                             group_by = 'flagAndGearType',
                             date_range = '2012-01-02,2020-12-30',
                             region = chn_code_eez$id[2],
                             region_source = 'eez',
                             key = key)

write_csv(chn_eez2, path = paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/CHN_", insert_code2, "_effort.csv"))
```

## Exploration: Check what some of the country-specific dataframes contains:

```{r}
# rus <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions/RUS_effort_2012_2020.csv"))
# chn <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions/CHN_effort_2012_2020.csv"))
# #spain:
# esp <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions/ESP_effort_2012_2020.csv"))
# bih <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions/BIH_effort_2012_2020.csv"))
```

## Dataset cleaning: remove files with 0 rows in order to combine all files row-wise into one object in next steps

Document which files (EEZ regions) had no fishing detected by GFW AIS data in 2012-2020 in order to notice trends over years.

```{r}
fish_effort_files <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/"), pattern = ".csv", full = TRUE)

# first check the length to see how many files we start with
num_files_start <- length(fish_effort_files)
print(paste0("Starting with ", num_files_start, " files."))

# delete files that do not have any rows (because no fishing has been recorded in that EEZ)
for (i in seq_along(fish_effort_files)) {
  # read each file and check number of rows
  filename <- fish_effort_files[i]
  print(paste0("Counting rows (fishing effort observations) for ", substr(filename, -29, -21)))
  # save the numbe of rows for that file to an object
  rows <- nrow(data.table::fread(filename))
  print(paste0(rows, " rows in this file."))
  # if there are 0 rows, delete the file
  if (rows == "0") {
      unlink(filename) 
    }
}

# redefine variable after deleting some files
fish_effort_files <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/"), pattern = ".csv", full = TRUE)

# check length again to see how many files we end with
  num_files_end <- length(fish_effort_files)
  print(paste0("Ending with ", num_files_end, " files because ", (num_files_start - num_files_end), " files were deleted for containing no data."))
# 2022 assessment: __ files were removed because had 0 rows, indicating no fishing activity at all within those countries' EEZs in 2012-2020 (from GFW AIS detections):
```

## Data Wrangling: Combine regional csv's into one dataframe and filter by geartype

We do this because our goal is to summarize fishing effort spatially to produce annual rasters of fishing effort. Later, we will use the fishing coordinates and OHI region polygons to attribute fishing activity to certain OHI regions.

```{r}
# concatenate csv's into one list for trawling data, separate from dredging so we can correct the trawling data and then sum the corrected trawling data with the dredging data later on
fish_effort_trawl <- fish_effort_files %>%
  lapply(data.table::fread) %>% # read in each file that represents data for 1 EEZ for 1 country
  bind_rows() %>% # combine all files into one dataframe
  filter(geartype == "trawlers") %>% # trawl fishing damages the seafloor
  dplyr::select(-geartype) # now that all observations are trawlers, we can drop this variable

# concatenate csv's into one list for dredging data, we will later sum this with the corrected trawling data
fish_effort_dredge <- fish_effort_files %>%
  lapply(data.table::fread) %>% # read in each file that represents data for 1 EEZ for 1 country
  bind_rows() %>% # combine all files into one dataframe
  filter(geartype == "dredge_fishing") %>% # dredge fishing damages the seafloor
  dplyr::select(-geartype) # now that all observations are from dredging, we can drop this variable

# take a look at the data
head(fish_effort_dredge) # still contains admin rgn
```

## Group by coordinate and year for apparent fishing effort for both geartypes of interest

We do not group by the administrative region for each EEZ because the coordinates of the fishing activity will allow us to attribute fishing activity to OHI regions. 

```{r}
# trawling
fish_effort_trawl_annual <- fish_effort_trawl %>% 
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 
# convert year column from integer to factor in preparation for next steps:
fish_effort_trawl_annual$year <- as.factor(fish_effort_trawl_annual$year)


# dredging
fish_effort_dredge_annual <- fish_effort_dredge %>% # all dredge data will be used, we will not be subsetting it like we will for the trawling data
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 
# convert year column from integer to factor in preparation for next steps:
fish_effort_dredge_annual$year <- as.factor(fish_effort_dredge_annual$year)


# take a look
head(fish_effort_dredge_annual) # no longer contains EEZ admin rgn
```

## Check one annual raster of trawler fishing effort: 2015 as a test before doing all years in loop

```{r}
# try with just 2015 trawling
# rasterize the data:

# fish_effort_trawl_2015 <- fish_effort_trawl_annual %>%
#   dplyr::filter(year == "2015") %>%
#   dplyr::select(-year)
# 
# # convert the dataframe to a SpatRaster
# # sometimes running this for the first time fails but just run again and it works
# fish_effort_trawl_2015_rast <- terra::rast(fish_effort_trawl_2015, type = "xyz", crs = "EPSG:4326", digits = 6, extent = NULL)
# 
# plot(fish_effort_trawl_2015_rast, col = "red") 
# 
# # compare to depth proportion trawling raster
# 
# 
# # convert to points vector (a geometry column?)
# #fish_effort_2015_points <- terra::as.points(fish_effort_2015_rast)
# #plot(fish_effort_2015_points, col = "red") # looks like blobs where the fishing points were! I suppose the difference between coordinates and points is that points just appear larger when plotted
# 
# # load the EEZ spatial data to layer on plot with fishing effort points for reference
# regions_shape()
# 
# # check out df subset to not overwhelm Mazu by opening regions df in its entirety
# View(head(regions))
# 
# # filter the regions dataframe for just EEZ polygons
# regions_eez <- regions %>%
#   filter(rgn_type == "eez") %>% 
#   st_transform(crs = 4326)
# 
# # try plotting just geometry
# plot(regions_eez$geometry, col = "red")
# 
# # try to plot points on polygons
# #plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "fishing effort 2015", legend = FALSE)		
# #plot(fish_effort_2015_points, axes = FALSE, col = "red", add = TRUE) # points are large! not great for viz
# 
# # Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
# plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2015: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
# plot(fish_effort_trawl_2015_rast, axes = FALSE, col = "red", add = TRUE)
```

### Save the trawling and dredging data as annual raster files before we can adjust them for mid-water versus bottom trawling.

```{r}
tic()
# this loop sometimes randomly errors but just run again and it should work (might error when Mazu's memory is almost full...)
years = as.factor(2012:2020) # make it factor in order to match the class of year column in the fish_effort_annual dataframe

for(i in years){
  # trawling data:
  trawl_annual_raster <- fish_effort_trawl_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast(type = "xyz", crs = "EPSG:4326", digits = 6, extent = NULL)
  
  # save annual raster file for trawling:
  terra::writeRaster(trawl_annual_raster, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i,"/fish_effort_trawl_", i, ".tif"), overwrite = TRUE)
}

for(i in years){
  # dredging data:
  dredge_annual_raster <- fish_effort_dredge_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast(type = "xyz", crs = "EPSG:4326", digits = 6, extent = NULL)

  # save annual raster file for dredging:
  terra::writeRaster(dredge_annual_raster, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_dredge_", i , ".tif"), overwrite = TRUE) # put in folder that will be the destination of the trawling data as well, after it is corrected, so the dredging and trawling by year can be summed easily in a stack
}

toc() # 9 minutes
```

### Correct Trawling Data: Subset trawling fishing effort by catch data for the corresponding year to distinguish between mid-water trawling and bottom trawling

We use fisheries catch data from the paper by Watson et al. to distinguish between the two types of trawling, because the GFW data groups all trawling together. We only maintain bottom trawling because that is the type that destroys soft bottom habitat. We multiply a raster that represents a proxy for the proportion of mid-water trawling to bottom trawling by the Global Fishing Watch raster for that respective year. We then sum that data with all the dredging data in the next steps. We only have catch data (the trawling correction data) for 2012-2017, so we use 2017 data for 2018-2020 GFW data.

### Try 2012 alone before iterating in loop

```{r}
# fish_effort_files <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/"), pattern = ".csv", full = TRUE)
# first read in the annual rasters that represent the proportion of mid-water trawling versus bottom trawling, cell values range 0-1
# these files were created on July 29, 2022 from Watson et al. data by Gage
# value of 0 = all mid-water trawling
# value of 1 = all bottom trawling
# try interpolating only 1 trawling proportion raster before doing all years in a loop
# trawl_depth_proportion_2012 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2012.tif'))
# # plot 2012 data to get an idea of what it looks like
# # pretty interesting that the EEZ outlines can be distinguished in the trawling locality
# terra::plot(trawl_depth_proportion_2012)
# 
# # check how many NA values we start with
# value <- terra::global(trawl_depth_proportion_2012, fun = "isNA") # 191,377
# value[1,1]
# #terra::global(trawl_depth_proportion_2012, fun = "isNA")[1,1]
# # read in years 2013-2017
# # trawl_depth_proportion_2013 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2013.tif'))
# # trawl_depth_proportion_2014 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2014.tif'))
# # trawl_depth_proportion_2015 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2015.tif'))
# # trawl_depth_proportion_2016 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2016.tif'))
# # trawl_depth_proportion_2017 <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2017.tif'))
# 
# # try to take the mean gage did, with raster, to see what output is supposed to look like 
# # trawl_depth_proportion_2012_raster <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2012.tif'))
# # avg_trawl_correction_raster <- cellStats(trawl_depth_proportion_2012_raster, "mean", na.rm=TRUE)
# # plot(avg_trawl_correction_raster) # 1 value not an actual raster 
# 
# #terra::app() is not what we want
# #terra::mean() is not what we want 
# # maybe global is what we want?
# # maybe terra::approximate() is what we want?
# #trawl_depth_proportion_2012_interpolated <- terra::global(trawl_depth_proportion_2012, fun = "mean")
# # maybe terra::focal is what we want ? best option so far, would be even more accurate than taking mean across entire layer because takes a local avg with window
# trawl_depth_proportion_interpolated_2012 <- terra::focal(x = trawl_depth_proportion_2012, 
#                                                          w = 9, 
#                                                          fun = "mean", 
#                                                          na.policy = "only", 
#                                                          na.rm = T)
# 
# terra::plot(trawl_depth_proportion_interpolated_2012)
# 
# # check how many NA values were interpolated
# terra::global(trawl_depth_proportion_interpolated_2012, fun = "isNA") # 102,832
# 
# # fill in the rest of the NA values with global average
# global_trawl_proportion_avg_2012 <- terra::global(trawl_depth_proportion_interpolated_2012, fun = "mean", na.rm = TRUE)
# trawl_depth_proportion_interpolated_2012[is.na(trawl_depth_proportion_interpolated_2012)] <- global_trawl_proportion_avg_2012[1,1]
# # 0 NA values left
# # THIS STEP ABOVE IS WHAT FILLS IN THE CONTINENTS WITH VALUES, SO THE CONTINENTS MUST BE NA RATHER THAN 0, SO NEED TO CONVERT THE CONTINENTS TO 0 BEFORE DOING ANY INTERPOLATION ?? OR MAYET THIS DOESNT MATTER BECAUSE THESE LAND POINTS WILL NOT MATCH UP WITH THE TRAWLING POINTS ANYWAY
# plot(trawl_depth_proportion_interpolated_2012)
# # na.only = "only" argument means we interpolate for only the NA values of the raster 
# # w = 9 means the moving window is 9 cells by 9 cells (a square) around the focal point we are interpolating
# # code help from: https://stackoverflow.com/questions/71801889/how-to-fill-in-missing-na-values-in-raster-with-terra-package
# 
# # check the number of na values in that raster after interpolation
# #terra::global(trawl_depth_proportion_2012_interpolated, fun = "isNA") # 102832, the number of na values decreased by 88545
# # check the number of not na values
# #terra::global(trawl_depth_proportion_2012_interpolated, fun = "notNA") # 156368, the number of not na values increased by 88454
# 
# # now that the trawl proportion data has been interpolated, we need to adjust its spatial resolution to match that of the fishing effort data from GFW for the corresponding year 
# # first read in the GFW raster for the year of interest
# fish_effort_trawl_2012 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2012/fish_effort_trawl_2012.tif"))
# # resample the trawling proportion data to match the resolution and extent of the GFW fishing effort data (use the GFW as the sample geometry)
# trawl_depth_proportion_resampled_2012 <- terra::resample(x = trawl_depth_proportion_interpolated_2012, y = fish_effort_trawl_2012, method = "near")
# # question: should we resample before we interpolate perhaps? - note to ask Gage & Mel
# terra::plot(trawl_depth_proportion_resampled_2012)
# terra::global(trawl_depth_proportion_resampled_2012, fun = "isNA") # 12837
# # get rid of the rest of the NA values produced by resample:
# trawl_depth_proportion_resampled_2012[is.na(trawl_depth_proportion_resampled_2012)] <- global_trawl_proportion_avg_2012[1,1]
# 
# # check extents of each - they match! 
# ext(terra::rast(fish_effort_trawl_2012))
# ext(terra::rast(trawl_depth_proportion_resampled_2012))
# 
# terra::writeRaster(trawl_depth_proportion_resampled_2012, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2012/trawl_depth_proportion_resampled_2012.tif"))
# 
# # stack the resampled interpolated raster of bottom trawling with all trawling data from gfw, matching the data by year for 2012-2017 and then using 2017 for the rest of the years of gfw trawling data for which we do not have watson data 
# trawl_stack_2012 <- terra::rast(list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2012"), pattern = ".tif", full = TRUE))
# 
# # multiply the rasters in order to only maintain the proportion of trawling fishing effort that is attributed to bottom trawling rather than midwater trawling
# trawling_corrected_2012 <- terra::lapp(trawl_stack_2012, fun = function(x,y){return(x*y)}, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/trawling_corrected/trawling_corrected_2012.tif"), overwrite = TRUE)
# 
# terra::plot(trawling_corrected_2012, col = "red")
# this plot represents the subsetted trawling for 2012
```

### Iterate through 2012-2017 data years to subset trawling data for just bottom trawling

```{r}
tic()

# use this subset because we only have trawling proportion data for these years from Watson et al.
years_subset = as.factor(2012:2017)

# for years with trawling depth proportion data:
for (j in years_subset){
  
  print(paste0("Processing trawling fishing effort for ", j))
  # read in the GFW apparent fishing effort raster for trawling
  # we need this read in to resample the Watson catch data to the same resolution
  fish_effort_trawl <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/fish_effort_trawl_", j, ".tif"))
  
  # read in the annual file that represents the proportion of bottom trawling to midwater trawling, based on catch data, from Watson et al.
  trawl_depth_proportion <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_", j, ".tif"))
  
  print(paste0("Interpolating trawling depth proportion data for ", j))
  # interpolate many of the missing values in the trawling depth proportion raster by taking the local average of cells around it (using a window of 9 cells) that represent the proportion of bottom trawling to midwater trawling that occurred at that coordinate. This provides a more comprehensive raster which we will later multiply by the GFW trawling raster, cell by cell
  trawl_depth_proportion_interpolated <- terra::focal(x = trawl_depth_proportion, 
                                                      w = 5, # sets a 5 cell window around NA pixel
                                                      fun = "mean", 
                                                      na.policy = "only", # only interpolate NA values
                                                      na.rm = T,
                                                      overwrite = TRUE)
  
  # document how many NA values were filled by `terra::focal()`
  print(paste0("Started with ", terra::global(trawl_depth_proportion, fun = "isNA")[1,1], " NA values, and ended with ", terra::global(trawl_depth_proportion_interpolated, fun = "isNA")[1,1], " so filled ", terra::global(trawl_depth_proportion, fun = "isNA")[1,1] - terra::global(trawl_depth_proportion_interpolated, fun = "isNA")[1,1], " NA values using `terra::focal()`."))
  
  # clear some memory
  rm(trawl_depth_proportion)
  gc()
  
  # fill in remaining NA values with the global mean of the non-NA cell values
  global_trawl_proportion_avg <- terra::global(trawl_depth_proportion_interpolated, fun = "mean", na.rm = TRUE)
  trawl_depth_proportion_interpolated[is.na(trawl_depth_proportion_interpolated)] <- global_trawl_proportion_avg[1,1]
  
  # document how many NA values remain after applying the local average to remaining NA cells (should be 0)
  print(paste0("There are now ", terra::global(trawl_depth_proportion_interpolated, fun = "isNA"), " NA values present after filling the remaining NA's with the global average."))
  
  print(paste0("Resampling trawling depth proportion data for ", j))
  
  # resample the interpolated trawling proportion data to match the resolution of the GFW fishing effort data
  trawl_depth_proportion_resampled <- terra::resample(x = trawl_depth_proportion_interpolated, 
                                                      y = fish_effort_trawl, # use the GFW data as the sample geometry
                                                      method = "near") # nearest neighbor
  
  # document how many NA values were produced by resampling (NA values are produced for some years, but not all)
  print(paste0("Now there are ", terra::global(trawl_depth_proportion_resampled, fun = "isNA")[1,1], " NA values remaining, need to fill NA's one final time."))
  
  # get rid of the rest of any NA values produced by resample:
  trawl_depth_proportion_resampled[is.na(trawl_depth_proportion_resampled)] <- global_trawl_proportion_avg[1,1]
  
  # document how many NA values remain (should be 0)
  print(paste0("Now there are ", terra::global(trawl_depth_proportion_resampled, fun = "isNA")[1,1], " NA values remaining."))
  
  # save the adjusted bottom trawling proportion raster in the folder of the corresponding year of GFW data, & save the most recent year of proportion data (2017) for 2017-2020 GFW data
  if (j == 2017) {
    # save to folders for 2017-2020
    print(paste0("Saving ", j, " file to ", j, " folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2018 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2018/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2019 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2019/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
    print(paste0("Saving ", j, " file to 2020 folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2020/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
  } else {
    # for trawling proportion data for years 2012-2016, save file to just that year's folder
    print(paste0("Saving ", j, " file to ", j, " folder."))
    terra::writeRaster(x = trawl_depth_proportion_resampled, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", j, "/trawl_depth_proportion_resampled_", j, ".tif"), overwrite = TRUE)
  }  
}

# clear some memory
rm(trawl_depth_proportion_resampled)
gc()

toc() # 14 minutes
```

```{r}
# scratch code
# check the origin and resolution of the rasters btw data sources
# read in the GFW apparent fishing effort raster for trawling, we need this read in to resample the Watson catch data to the same resolution
# fish_effort_trawl <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2012/fish_effort_trawl_2012.tif"))
# # read in the annual file that represents the proportion of bottom trawling to midwater trawling, based on catch data, from Watson 
# trawl_depth_proportion <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2012.tif"))
# 
# fish_effort_trawl # res 0.01, 0.01, ext -179.985, 180.005, -55.305, 73.065  (xmin, xmax, ymin, ymax)
# trawl_depth_proportion # res 0.5, 0.5, ext -180, 180, -90, 90  (xmin, xmax, ymin, ymax)
# 
# origin(fish_effort_trawl) 
# fish_effort_trawl_disagg <- terra::disagg(fish_effort_trawl, fact = 50, method = "near")
# 
# 
# 
# origin(trawl_depth_proportion)
# 
# terra::global(fish_effort_trawl, fun = "isNA")

# 2017
# test that raster that was saved looks like we would expect compared to unprocessed raster from Watson
# watson_unprocessed_2017 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2017.tif"))
# 
# plot(watson_unprocessed_2017)
# 
# terra::global(watson_unprocessed_2017, fun = "isNA") # 201,000
# 
# watson_processed_2017 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2017/trawl_depth_proportion_resampled_2017.tif"))
# 
# plot(watson_processed_2017)
# 
# terra::global(watson_processed_2017, fun = "isNA") # 156,880,905
# 
# # 2012
# # test that raster that was saved looks like we would expect compared to unprocessed raster from Watson
# watson_unprocessed_2012 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/trawling_correction/bottom_trawl_props/bottom_trawl_prop_2012.tif"))
# 
# plot(watson_unprocessed_2012)
# 
# terra::global(watson_unprocessed_2012, fun = "isNA") # 191377
# 
# watson_processed_2012 <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2012/trawl_depth_proportion_resampled_2012.tif"))
# 
# plot(watson_processed_2012)
# 
# terra::global(watson_processed_2012, fun = "isNA") # 135,657,275
```

#### Apply the adjusted fishing catch rasters (which have been resampled and interpolated) to subset the GFW trawling data for just bottom trawling

First, we extend the trawling depth proportion raster to the larger extent of the trawling fishing effort raster and vice versa to ensure all dimensions of both extents are maximized and match one another. We then multiply the proportion raster by the trawling raster to reflect the amount of trawling that is attributed to bottom trawling. Mid-water trawling does not damage the seafloor.

```{r}
tic()

years_all <- as.factor(2012:2020)

for (i in years_all){
  
  print(paste0("Checking if extents match for ", i, " trawling and fishing catch rasters."))
  # first read in both rasters within the annual folders and save the list of 2 as an object
  rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i), pattern = ".tif", full = TRUE)
  fish_effort_raster <- terra::rast(rasters[1])
  trawl_depth_proportion_raster <- terra::rast(rasters[2])
  print(ext(fish_effort_raster)) # peek at the raster extents
  print(ext(trawl_depth_proportion_raster))
  extent_comparison <- terra::compareGeom(fish_effort_raster, trawl_depth_proportion_raster, stopOnError = FALSE) # set stopOnError = FALSE in order to produce "FALSE" from function if extents don't match
  print(paste0("Raster extent comparison results: ", extent_comparison))

  if (extent_comparison == FALSE){ 
      print("Extending trawling proportion raster to match the extent of the trawling fishing effort raster and vice versa.")
      trawl_depth_proportion_extended <- terra::extend(trawl_depth_proportion_raster, fish_effort_raster)
      fish_effort_extended <- terra::extend(fish_effort_raster, trawl_depth_proportion_extended) # extend both rasters to the extent of the other in all dimensions
    
      # check the extents now
      print(ext(fish_effort_extended))
      print(ext(trawl_depth_proportion_extended))
      extent_comparison <- terra::compareGeom(fish_effort_extended, trawl_depth_proportion_extended, stopOnError = FALSE)
      print(paste0("Raster extent comparison results: ", extent_comparison)) # should be TRUE
    
      # clear some memory
      rm(rasters)
      gc()
      
      # re-write the extended rasters to a nested folder called "extended" within the trawling annual folders so we don't need to overwrite the original raster files
      terra::writeRaster(trawl_depth_proportion_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/trawl_depth_proportion_extended_", i, ".tif"), overwrite = TRUE)
      print("Output raster for trawl proportion data with extent to match fishing effort data.")
      terra::writeRaster(fish_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/fish_effort_extended_", i, ".tif"), overwrite = TRUE)
      print("Output raster for fishing effort raster with extent to match trawl proportion data.")
      
      # since the rasters have been rewritten as new filenames, need to redefine the object that represents them
      rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/", i, "/extended/"), pattern = ".tif", full = TRUE)
      print("Redefined rasters list object.")
  }

  # stack the resampled interpolated raster of bottom trawling with all trawling data for that year from GFW, matching the data by year for 2012-2017 and then using 2017 for the rest of the years of GFW trawling data for which we do not have more recent trawling proportion data
  trawl_stack <- terra::rast(rasters)
  
  print(paste0("Correcting trawling effort data for ", i))
  
  # multiply the rasters in order to only maintain the proportion of trawling fishing effort that is attributed to bottom trawling rather than midwater trawling
  terra::lapp(trawl_stack, fun = function(x,y){return(x*y)}, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_trawl_corrected_", i, ".tif"), overwrite = TRUE)
  
  print(paste0("Saved corrected trawling data for ", i, " to ", paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/fish_effort_trawl_corrected_", i, ".tif")))

  # clear some memory
  rm(trawl_stack)
  gc()
}
  
toc() # 12 min
```

## Combine the corrected trawling data with the dredging data by summing the fishing effort by coordinate

Similarly to the last step, we extend the annual dredging rasters to the extent of the corrected trawling raster and vice versa. Since the dredging rasters generally have smaller extents than the trawling fishing effort rasters, many NA values are produced by extending it. The presence of NA values requires us to sum using the argument na.rm = TRUE in order to avoid nullifying a large proportion of the trawling fishing effort data points in the calculation. Since both the dredging data and the trawling data were pulled from the Global Fising Watch database, these files already have the same resolution and were assigned the same CRS earlier in the data processing.

```{r}
tic()
years_all <-  as.factor(2012:2020)

for (i in years_all){
  
  print(paste0("Checking if extents match for ", i, " trawling and dredging rasters."))
  # first read in both rasters within the annual folders and save the list of 2 as an object
  rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i), pattern = ".tif", full = TRUE)
  dredge_effort_raster <- terra::rast(rasters[1])
  trawl_effort_raster <- terra::rast(rasters[2])
  print(ext(dredge_effort_raster)) # smaller extent so extend this raster first
  print(ext(trawl_effort_raster))
  extent_comparison <- terra::compareGeom(dredge_effort_raster, trawl_effort_raster, stopOnError = FALSE)
  print(paste0("Raster extent comparison results: ", extent_comparison))

  if (extent_comparison == FALSE){ 
      print("Extending trawling proportion raster to match the extent of the dredging effort raster and vice versa.")
      dredge_effort_extended <- terra::extend(dredge_effort_raster, trawl_effort_raster)
      trawl_effort_extended <- terra::extend(trawl_effort_raster, dredge_effort_extended) # extend both rasters to the extent of the other in all dimensions
    
      # check the extents now
      print(ext(dredge_effort_extended))
      print(ext(trawl_effort_extended))
      extent_comparison <- terra::compareGeom(dredge_effort_extended, trawl_effort_extended, stopOnError = FALSE)
      print(paste0("Raster extent comparison results: ", extent_comparison)) # should be TRUE
    
      # clear some memory
      rm(rasters)
      gc()
      
      # re-write the extended rasters to a nested folder called "extended" to avoid overwriting the original files 
      terra::writeRaster(trawl_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/fish_effort_dredge_", i, "_extended.tif"), overwrite = TRUE)
      print("Re-wrote trawling effort file.")
      terra::writeRaster(dredge_effort_extended, paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/fish_effort_trawl_", i, "_extended.tif"), overwrite = TRUE)
      print("Re-wrote dredging effort file.")
      
      # since the rasters have been rewritten, need to redefine the object that represents them
      rasters <- list.files(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/", i, "/extended/"), pattern = ".tif", full = TRUE)
      print("Redefined rasters list object.")
  }

  # stack the corrected trawling data with the dredging data from GFW, matching the data by year
  effort_stack <- terra::rast(rasters)
  
  print(paste0("Correcting trawling effort data for ", i))

  terra::app(effort_stack, fun = "sum", na.rm = TRUE, filename = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"), overwrite = TRUE)

  # clear some memory
  rm(effort_stack)
  gc()
  
  print(paste0("Saved corrected fishing effort data for ", i, " to ", paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif")))
}
  
toc() # 15 minutes
```

### Plot the annual rasters to visualize fishing effort over time

```{r}
# load the EEZ spatial data to layer on plot with fishing effort points for reference
regions_shape()

# define function to be used in loop:
#map <- function(i){
#    plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawling & Dredging Fishing Effort ", i, ": AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
#    plot(fishing_effort_annual, axes = FALSE, col = "red", add = TRUE)
#    }

# Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
#plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2015: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
#plot(fish_effort_trawl_2015_rast, axes = FALSE, col = "red", add = TRUE)

# look into the original trawling_data, before correction with the catch proportion data
#trawling_og <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/trawling/2013/fish_effort_trawl_2013.tif"))
#plot(trawling_og, col = "red")

# look into the corrected catch data raster
#trawling_corrected <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/dredging_trawling_sum/2013/fish_effort_trawl_corrected_2013.tif"))
#plot(trawling_corrected, col = "red")

for (i in years_all){
  # read in the annual file for combined trawling and dredging fishing effort
  fishing_effort_annual <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_", i, ".tif"))
  plot(fishing_effort_annual, col = "red") 
  
  # Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
  plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawling & Dredging Fishing Effort ", i, ": AUS, USA, THA, BIH, GBR, ITA", legend = FALSE))
  plot(fishing_effort_annual, axes = FALSE, col = "red", add = TRUE)
  }

```

```{r}
# delete:

# convert to points vector (a geometry column?)
#fish_effort_2015_points <- terra::as.points(fish_effort_2015_rast)
#plot(fish_effort_2015_points, col = "red") # looks like blobs where the fishing points were! I suppose the difference between coordinates and points is that points just appear larger when plotted

# load the EEZ spatial data to layer on plot with fishing effort points for reference
regions_shape()

# check out df subset to not overwhelm Mazu by opening regions df in its entirety
View(head(regions))

# filter the regions dataframe for just EEZ polygons
regions_eez <- regions %>%
  filter(rgn_type == "eez") %>% 
  st_transform(crs = 4326)

# try plotting just geometry
plot(regions_eez$geometry, col = "red")

# try to plot points on polygons
#plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "fishing effort 2015", legend = FALSE)		
#plot(fish_effort_2015_points, axes = FALSE, col = "red", add = TRUE) # points are large! not great for viz

# Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2015: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
plot(fish_effort_trawl_2015_rast, axes = FALSE, col = "red", add = TRUE)
```


Next steps:
- iterate over all countries (see if there are more countries that have too much fishing for the API)
- adjust to quantile if necessary
- add eval = false to each chunk, delete old commented out code, fix yaml, knit
- trend

### Calculate fishing effort for each OHI region

```{r}
# Read in the combined trawling and dredging raster that will serve as a CRS reference, this could be any year of data because they all have the same CRS 
#fishing_raster <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2012.tif"))

# Read in OHI regions polygon
ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update") # geometry type = multipolygon

# Make the region coordinate reference system the same as the trawling & dredging raster
ohi_regions_wgs <-  st_transform(x = ohi_regions, crs = "EPSG:4326")
# check:
#crs(ohi_regions_wgs)
class(ohi_regions_wgs) # sf dataframe
View(head(ohi_regions_wgs))
unique(ohi_regions_wgs$type_w_ant)

# convert the dataframe into a spatVector so we can extract the fishing points from the OHI region polygons 
#ohi_regions_vec <- terra::vect(ohi_regions_wgs)
# convert the daatframe into a --?


# 2012 only before doing in loop
print(paste("Processing ", i))

fishing_raster <- terra::rast(paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/final_rasters/dredging_trawling_combined_2012.tif"))

# extract the fishing effort that falls within the polygons of OHI regions
#extracted <- terra::extract(fishing_raster, ohi_regions_vec, weights = TRUE)

# cannot extract the spatRaster from the `sf dataframe` object ohi_regions_wgs with the argument weights = area because error is returned:
# Error in weights == "area" : comparison (1) is possible only for atomic and list types


# check the geom in ohi_regions_wgs is valid before running extact_extract
#ohi_regions_valid <- sf::st_is_valid(ohi_regions_wgs)
#ohi_regions_valid # all valid 
# checks if geometries are valid for an sf object (returns true or false for each polygon row)

class(ohi_regions_wgs)
extracted <- exactextractr::exact_extract(fishing_raster, ohi_regions_wgs, fun = "weighted_sum", weights = "area")

# assign names to be the cell terrain type (eez or land), and the OHI region ID to each value in the vector so we can differentiate between those we need and those we don't
names(extracted) <- paste(ohi_regions_wgs$type_w_ant, ohi_regions_wgs$rgn_ant_id, sep = " ") 
extracted
# create a df of all these cell values and 
fishing_regional <- plyr::ldply(extracted, rbind) %>% 
  rename(fishing_effort = "1",
         "rgn_type rgn_id" = ".id") %>% 
  tidyr::separate(col = "rgn_type rgn_id", 
                  into = c("rgn_type", "rgn_id"),
                  sep = " ")

unique(fishing_regional$rgn_type) # [1] "eez"           "land"          "eez-inland"    "eez-ccamlr"    "land-ccamlr"   "eez-disputed"  "land-disputed" "land-noeez"   # [9] "fao"

# The following code keeps the raster information when the raster lands partially on the land polygons
fishing_regional_summary <- fishing_regional %>%
  #tidyr::separate(.id, c("rgn_type", "rgn_id"), sep = "_") %>%
  #dplyr::mutate(effort = value*weight) %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::summarize(effort = sum(fishing_effort, na.rm = TRUE)) %>% # why are there over 220 rgns? maybe old regions that arent part of modern ohi analysis?
  dplyr::ungroup() %>%
  dplyr::arrange(as.numeric(rgn_id)) #%>%
 # dplyr::mutate(year = effort_year)

# Guernsey (rgn 228) sure does fish a lot?

# convert this raster info into a csv that represents the trawling and dredging fishing effort for each OHI region for year i 
write.csv(fishing_regional_summary, file = paste0(dir_M, "/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fishing_effort_annual/fishing_annual_csv/fishing_effort_", i, ".csv"), row.names = FALSE)





 
## Gage's code to extract values for each OHI region

## Loop to extract values for each OHI region
foreach(sb_catch = catch_rasts) %dopar% { 
   # sb_catch = catch_rasts[12] if you just want to run one year's raster without looping through all of them

cat(sb_catch)

#catch_year <- str_extract(basename(sb_catch), "(\\d)+") # Extract the corresponding year for each file

catch <- raster(sb_catch)

data <- raster::extract(catch, ohi_regions_wgs, weights = TRUE, normalizeWeights = FALSE, progress = 'text') 

names(data) <- paste(ohi_regions$type_w_ant, ohi_regions$rgn_ant_id, sep="_") 
sb_catch_rgn   <- plyr::ldply(data, rbind)

# The following code keeps the raster information when the raster lands partially on the land polygons
sb_catch_rgn <- sb_catch_rgn %>%
  tidyr::separate(.id, c("rgn_type", "rgn_id"), sep="_") %>%
  dplyr::mutate(tonnes = value*weight) %>%
  dplyr::group_by(rgn_type, rgn_id) %>%
  dplyr::summarize(tonnes = sum(tonnes, na.rm = TRUE)) %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(as.numeric(as.character(rgn_id))) %>%
  dplyr::mutate(year = catch_year)

write.csv(sb_catch_rgn, file = file.path(here(), "globalprep/hab_prs_hd_subtidal_soft_bottom/v2020/int", sprintf("sb_catch_rgn_%s.csv", catch_year)), row.names=FALSE)

}


```










































## Exploration: Check that regions that are infamous for trawling are recorded as trawling as much as we would expect:

```{r}
# usa <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "USA") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))
# 
# chn <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "CHN") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))
# 
# nzl <- fish_effort_all %>% 
#   filter(eez_admin_rgn == "NZL") %>% 
#   group_by(year) %>% 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours))

# see which countries have the most trawling
# rgn_trawl <- fish_effort_all %>% 
#   group_by(eez_admin_rgn) %>% 
#   summarise(total_fishing_hours = sum(apparent_fishing_hours))
# 
# rgn_trawl_max <- rgn_trawl %>% 
#   slice_max(total_fishing_hours, n = 20, with_ties = FALSE) %>%
#   arrange(desc(total_fishing_hours)) 
# 
# # visualize distribution of all regions trawling
# trawling_all <- ggplot(data = rgn_trawl_max, aes(x = eez_admin_rgn, y = total_fishing_hours)) +
#   geom_point() +
#   geom_point(size = 3) +
#   labs(title = "Top 20 Trawling Regions, 2012-2020",
#        subtitle = "GFW Data",
#        x = "Region Code",
#        y = "Total Trawling Fishing Effort Hours") +
#   theme_minimal()
# 
# trawling_all
```

## Data Wrangling: all gear types

```{r}
# fish_effort_allgear <- list.files(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/all_regions"), pattern = ".csv", full = TRUE) %>%
#   lapply(data.table::fread) %>%
#   bind_rows()
# 
# head(fish_effort_allgear)
# 
# unique(fish_effort_allgear$eez_admin_rgn)
# 
# # see which countries have the most fishing in general - all geartypes
# rgn_fe <- fish_effort_allgear %>% 
#   group_by(eez_admin_rgn) %>% 
#   summarise(total_fishing_hours = sum(apparent_fishing_hours))
# 
# rgn_fe_max <- rgn_fe %>% 
#   slice_max(total_fishing_hours, n = 20, with_ties = FALSE) %>%
#   arrange(desc(total_fishing_hours)) 
# 
# # visualize distribution of all regions trawling
# trawling_allgear <- ggplot(data = rgn_fe_max, aes(x = eez_admin_rgn, y = total_fishing_hours)) +
#   geom_point() +
#   geom_point(size = 3) +
#   labs(title = "Top 20 Fishing Regions (all geartypes), 2012-2020",
#        subtitle = "GFW Data",
#        x = "Region Code",
#        y = "Total Fishing Effort Hours") +
#   theme_minimal()
# 
# trawling_allgear
```


## Group apparent trawling fishing effort by latitude, longitude, and year for raster analysis

```{r}
# group all countries and years by lat, long, and year (this df does not have the EEZ admin country anymore, since we don't need that at the moment)
fish_effort_annual <- fish_effort_trawl %>% # change to the relevant trawl data after filtering for that with catch data later
  group_by(x, y, year) %>% # 2022 assessment: did not group by year (according to output) because no exact coordinate was repeated one multiple years, so we will do that in next step
  summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) 

head(fish_effort_annual)

# convert year column from integer to factor in preparation for next steps:
fish_effort_annual$year <- as.factor(fish_effort_annual$year)

# save dataframe before rasterizing:
#write_csv(, ".csv")
```

## Check one annual raster of trawler fishing effort: 2015 as a test before doing all years in loop

```{r}
# try with just 2015 first before loop:
# rasterize the data:

fish_effort_2015 <- fish_effort_trawl %>%
  dplyr::filter(year == "2015") %>%
  dplyr::select(-year)

# set spatial coordinates of a dataframe
# sp::coordinates(fish_effort_2015) <- ~x+y # errors if use "x" + "y"
# # assign EPSG 4326 as the CRS
# raster::rasterFromXYZ(fish_effort_2015, crs = "+init=epsg:4326", digits = 6) 
# proj4string(fish_effort_2015) = CRS("+init=epsg:4326") # code from Gage's footprint project
# #raster::rasterFromXYZ(fish_effort_2015, crs = "EPSG:4326", digits = 6)
# crs(fish_effort_2015)
# plot(fish_effort_2015, col = "red") # plot in red to visualize points easier

# check if using diff raster function makes output diff?
#raster::writeRaster(fish_effort_2015, filename = paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/test.tif'), overwrite = TRUE)

# convert the dataframe to a SpatRaster
fish_effort_2015_rast <- terra::rast(fish_effort_2015)

plot(fish_effort_2015_rast, col = "red") # looks like points but needs to be a df to intersect with EEZ polygons

# convert to points vector (a geometry column?)
#fish_effort_2015_points <- terra::as.points(fish_effort_2015_rast)

#plot(fish_effort_2015_points, col = "red") # looks like blobs where the fishing points were! I suppose the difference between coordinates and points is that points just appear larger when plotted

# load the EEZ spatial data to layer on plot with fishing effort points for reference
regions_shape()

# check out df subset to not overwhelm Mazu by opening regions df in its entirety
View(head(regions))

# filter the regions dataframe for just EEZ polygons
regions_eez <- regions %>%
  filter(rgn_type == "eez") %>% 
  st_transform(crs = 4326)

# try plotting just geometry
plot(regions_eez$geometry, col = "red")

# try to plot points on polygons
#plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "fishing effort 2015", legend = FALSE)		
#plot(fish_effort_2015_points, axes = FALSE, col = "red", add = TRUE) # points are large! not great for viz

# Plot the raster of fishing effort points (without making them terra points) on top of the EEZ polygons
plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = "Fishing Effort 2015: AUS, USA, THA, BIH, GBR, ITA", legend = FALSE)
plot(fish_effort_2015_rast, axes = FALSE, col = "red", add = TRUE)
```

### Visualize trawling & dredging apparent fishing effort data on map of EEZ's as a time series and save the rasters as annual files.

PLot each year and save the raster as a .tif.

- would like to add checks to this, progress bar, more cores for parallelization

```{r}
tic()

#colors <- c("yellow", "red", "blue", "green", "orange", "cyan4") # 6 colors because loop does i +1, so yellow will never be plotted

years = as.factor(2012:2020) # make it factor in order to match the class of year column in the fish_effort_annual dataframe

for(i in years){
  annual_raster <- fish_effort_annual %>%
    dplyr::filter(year == i) %>%
    dplyr::select(-year) %>%
    terra::rast()

    plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawling and Dredging Fishing Effort in ", i, " for AUS, USA, THA, BIH, GBR, ITA", legend = FALSE))
    plot(annual_raster, axes = FALSE, col = "red", add = TRUE, legend = FALSE)
    
    # save annual raster file that encompasses trawling fishing and dredging effort for all countries
    #raster::writeRaster(annual_raster, filename = paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_', i ,'.tif'), overwrite = TRUE)
}

toc()
```

### Plot all years on same map - after all, none of the coordinates were repeated exactly on different years

```{r}
# for(i in years){
#  annual_raster <- terra::rast(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_', i ,'.tif'))
# 
#     plot(regions_eez$geometry, col = "grey96", axes = FALSE, main = paste0("Trawing and Dredging Fishing Effort in 2012-2020 for AUS, USA, THA, BIH, GBR, ITA", legend = FALSE))
#     plot(annual_raster, axes = FALSE, col = "red", add = TRUE, legend = FALSE)
# }
# 
# toc()
```


```{r}
# fish_effort_2015 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2015.tif'))
# #fish_effort_2015_df <- raster::as.data.frame(fish_effort_2015, xy = TRUE, na.rm = FALSE)
# #tail(fish_effort_2015_df)
# fish_effort_2016 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2016.tif'))
# fish_effort_2017 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2017.tif'))
# fish_effort_2018 <- raster::raster(paste0(dir_M, '/git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2022/int/fish_effort_annual/fish_effort_2018.tif'))
```

```{r}
# must use dataframes rather than .tif files for ggplot
# ggplot() +
#   geom_raster(data = eez_df, aes(x = x, y = y, fill = 'eez')) +
#   # add eez csv for plotting:
#   geom_raster(data = fish_effort_2015_df, aes(x = x, y = y)) +
#   scale_fill_viridis_c() +
#   theme_void() +
#   theme(legend.position = "bottom") +
#   coord_equal()
```

## Plot faceted rasters for apparent fishing hours with overlaid map of EEZ's

```{r}
# eez_boundaries <- file.path()
# 
# ggplot() +
#   geom_raster(data = fish_effort_all, aes(x = lon, y = lat, fill = 'apparent_fishing_hours')) +
#   geom_sf(data = eez_boundaries, fill = NA) +
#   scale_fill_viridis_c() +
#   theme_void() +
#   theme(legend.position = "bottom") +
#   coord_equal()
```

## Create dataframe of summarized spatialized fishing effort while maintaining the administrative country for each EEZ

This step enables us to calculate scores for each region on the fishing that occurs in their EEZ?

```{r}
# recall dataframe from earlier with region variable still present: fish_effort_all
# fish_effort_regional <- fish_effort_all %>% 
#   group_by(lat, lon, year, eez_admin_rgn) %>% # only grouped by lat, lon, and year according to output, bc no 2 countries fished in the exact same coordinate in these years 
#   summarize(total_fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE))
# 
# year = 2015:2018
# 
# foreach(r = rgns_eez_subset$eez_iso3) %do% {
#   foreach(yr = year) %do% {
#     fish_effort <- fish_effort_all %>%
#       dplyr::filter(year == yr) %>% 
#       dplyr::group_by(lat, lon)
#   }
# }
```

```{r}
# already created files of fishing effort separated by country, start by reading those in, group by year, sum hours
# year = 2015:2018
# foreach(r = rgns_eez_subset$eez_iso3) %do% {
#   regional_fishing_effort <- read.csv(paste0(dir_M, "/git-annex/globalprep/_raw_data/global_fishing_watch/d2022/annual_mapping_api/", r, "_effort_15_18.csv")) %>% 
#     group_by(year) %>% 
#     summarize()
  #}
```


Old Notes:
- consider open issue about needing to subset the id list in order to plug it all into get_event()?



















